---
title: "Data download - errdap"
author: "Gage Clawson (IMAS)"
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output: 
  pdf_document: 
    number_sections: yes
    toc: true
    toc_depth: 4
editor_options: 
  chunk_output_type: console
---

# Summary 

Download ERDDAP sea surface temperature and chlorophyll data

```{r echo = FALSE}
# This chunk sets up default settings for all chunks below
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,fig.width = 7.5,fig.height = 5,dev = 'png',dpi=300)
```

```{r include=FALSE}
# Load all necessary packages
library(tidyverse)
library(sf)
library(stars)
library(rerddap)
library(rnaturalearth)
library(glue)
library(scico)
library(lubridate)
library(furrr)
library(here)
library(terra)
library(ncdf4)
library(strex)
source(here::here("prep/02_prep_spatial_layers_model/_functions_data_wrangling_erddap.R"))
source(here("R/dir.R"))

# Set the data directory. This specifies the folder in our drive where the data can be found. 

data_directory <- rdsi_raw_dir

# Get high-res ocean data
ocean <- ne_download(scale = 50, type = 'ocean', category = 'physical',returnclass = "sf") %>%
  dplyr::select(geometry)
```

# Defining our global grid

```{r}
pixel_size <- 0.5

# Start with global ocean
starting_shape <- ocean
```

Only need to run once: 

```{r eval = params$download_data}
# Make a grid using desired pixel size
data_grid <- starting_shape %>%
  make_grid_custom(pixel_size = pixel_size)

data_grid %>%
  # Add geometry wkt column for saving this as a csv
  mutate(geometry_wkt = st_as_text(geometry)) %>% 
  # Make geometry point, grab lon/lat in lower left-hand corner
  # We'll use this later for joining to GFW data
  st_cast("POINT") %>%
  dplyr::mutate(lon = sf::st_coordinates(.)[,1],
                lat = sf::st_coordinates(.)[,2])%>%
  st_set_geometry(NULL)  %>% 
  group_by(pixel_id,geometry_wkt) %>% 
  summarize(lon = min(lon),
            lat = min(lat))%>%
  ungroup()  %>%
  as_tibble() %>%
  write_csv(here::here("data/model_features/global_grid.csv"))
```

We define a global grid using pixels that are `r pixel_size` degree latitude by `r pixel_size` degree longitude. We will then aggregate all of our model feature data to these `r pixel_size`x`r pixel_size` degree pixels. We include only pixels that overlap with the ocean. Each pixel is assigned a static `pixel_id`.

```{r}
data_grid <- data.table::fread(here::here("data/model_features/global_grid.csv")) %>%
  st_as_sf(wkt = "geometry_wkt",
           crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

# test <- data_grid %>%
#   st_drop_geometry() %>%
#   dplyr::select(lon, lat, pixel_id) %>%
#   rast(., type = "xyz") # 0.5 by 0.5.. cool it worked.
# plot(test)

```

\pagebreak

# Environmental model features

## SST and SST anomaly data 

We use 0.25 degree data from [SST, Daily Optimum Interpolation (OI), AVHRR Only, Version 2.1, Final, Global, 0.25°, 1981-present, Lon+/-180](https://coastwatch.pfeg.noaa.gov/erddap/info/ncdcOisst21Agg_LonPM180/index.html). These data download ~25x as fast and are much faster to spatially aggregate.

Preferably we could use a data set that goes back to 1950.. but perhaps we just gapfill back to 1950 with climate projections instead of real satellite data like this? 

```{r eval = params$download_data}

# Download all global SST and SST anomaly data
download_errdap_data_wrapper(dataset_name = "ncdcOisst21Agg_LonPM180",
                             variables = c("sst","anom"),
                             temporal_resolution = "day",
                             date_start = "2024-01-01",
                             date_end = "2024-12-31",
                             data_grid = data_grid,
                             missing_dates = missing_dates_2,
                             run_parallel = TRUE)
```


```{r eval = params$aggregate_data}
# Are there missing dates from our downloads? If so, set the missing_dates variable to missing_dates_2 and rerun! 
missing_dates <- determine_missing_dates(dataset_name = "ncdcOisst21Agg_LonPM180",
                                         temporal_resolution = "day",
                                         date_start = "1981-09-02",
                                         date_end = "2024-12-31")

missing_dates_2 <- missing_dates[[1]]$date
```

Unfortunately, this SST data does not appear to be available to download through rerrdap as of July 9, 2025 (thanks Trump...). Here is code I will run to update the data for the remaining years we need to download (just 2024): https://www.ncei.noaa.gov/data/sea-surface-temperature-optimum-interpolation/v2.1/access/avhrr/

First, download the data:

```{r}
# R script to download NOAA OISST data for 2024 (January to December)
# Downloads daily files from NCEI AVHRR dataset

library(lubridate)

# Base URL for OISST AVHRR data
base_url <- "https://www.ncei.noaa.gov/data/sea-surface-temperature-optimum-interpolation/v2.1/access/avhrr/"


# Function to download a single file
download_oisst_file <- function(date, base_url, data_dir = "data") {
 
  # date <- as.Date("2024-01-01")
  
# Format date as YYYYMMDD
date_str <- format(date, format = "%Y%m%d")

# Format year-month for folder structure (YYYYMM)
year_month <- format(date, "%Y%m")
  
  # Construct filename
  filename <- paste0("oisst-avhrr-v02r01.", date_str, ".nc")
  
  # Construct full URL with monthly folder structure
  file_url <- paste0(base_url, year_month, "/", filename)
  
  # Local file path
  local_path <- file.path(data_directory, "erddap/ncdcOisst21Agg_LonPM180_raw", filename)
  
  # Check if file already exists
  if (file.exists(local_path)) {
    cat(sprintf("File already exists: %s\n", filename))
    return(TRUE)
  }
  
  # Attempt to download
  cat(sprintf("Downloading from %s/: %s\n", year_month, filename))
  
  tryCatch({
    download.file(file_url, local_path, mode = "wb", quiet = FALSE)
    
  }, error = function(e) {
    cat(sprintf("Error downloading %s: %s\n", filename, e$message))
    if (file.exists(local_path)) file.remove(local_path)
    return(FALSE)
  })
}

# Generate date sequence for 2024
start_date <- as.Date("2024-01-01")
end_date <- as.Date("2024-12-31")
dates <- seq(start_date, end_date, by = "day")

cat(sprintf("Date range: %s to %s\n", start_date, end_date))

# Download files
successful_downloads <- 0
failed_downloads <- 0
start_time <- Sys.time()

for (i in seq_along(dates)) {
#  i = 2
  date <- dates[i]
  
  # Progress indicator
  if (i %% 10 == 0 || i == 1) {
    cat(sprintf("Progress: %d/%d (%.1f%%)\n", i, length(dates), (i/length(dates))*100))
  }
  
  # Download file
  success <- download_oisst_file(date, base_url)
  
  if (success) {
    successful_downloads <- successful_downloads + 1
  } else {
    failed_downloads <- failed_downloads + 1
  }
  
  # Small delay to be respectful to the server
  Sys.sleep(0.5)
}

end_time <- Sys.time()
duration <- end_time - start_time


# List downloaded files
downloaded_files <- list.files(file.path(data_directory, "erddap/ncdcOisst21Agg_LonPM180_raw"), pattern = "oisst-avhrr-v02r01\\.2024.*\\.nc$", full.names = FALSE)
cat(sprintf("Files in data directory: %d\n", length(downloaded_files)))
# should be 366? 

```


```{r}
# we need to prep the netcdf data to be a csv with zlev, latitude, longitude, sst, and anom. 
files_to_run <- list.files(file.path(data_directory, "erddap/ncdcOisst21Agg_LonPM180_raw"), full.names = TRUE)

for(i in 1:length(files_to_run)){
# Open the NetCDF file

#   i = 2
  file_path <- files_to_run[[i]]
  nc_data <- nc_open(file_path)

# Extract coordinate variables
lon <- ncvar_get(nc_data, "lon")
lat <- ncvar_get(nc_data, "lat")
zlev <- ncvar_get(nc_data, "zlev")
time <- ncvar_get(nc_data, "time")

# Convert longitude from 0-360 to -180-180
lon[lon > 180] <- lon[lon > 180] - 360

# Extract data variables
sst_data <- ncvar_get(nc_data, "sst")
anom_data <- ncvar_get(nc_data, "anom")

# Close the NetCDF file
nc_close(nc_data)

# Create a data frame with all combinations of coordinates
# Note: NetCDF arrays are in [lon, lat] order
cat("\nCreating coordinate grid...\n")

# Create coordinate grids using expand.grid for proper pairing
coord_grid <- expand.grid(longitude = lon, latitude = lat)

# Flatten the data arrays (they are 2D: [lon, lat])
# as.vector reads column-wise, which matches expand.grid order
sst_flat <- as.vector(sst_data)
anom_flat <- as.vector(anom_data)

# Create the data frame
df <- data.frame(
  zlev = zlev[1],  # Single z-level value for all points
  latitude = coord_grid$latitude,
  longitude = coord_grid$longitude,
  sst = sst_flat,
  anom = anom_flat
)

# Write to CSV
file_date <- as.Date(str_after_last(str_before_last(basename(file_path), "\\.nc"), "\\."), format = "%Y%m%d")
output_file <- glue("ncdcOisst21Agg_LonPM180_{file_date}.csv")
write.csv(df, file.path(data_directory, "erddap/ncdcOisst21Agg_LonPM180", output_file), row.names = FALSE)

}

```


\pagebreak

# Chlorophyll

We get monthly chlorophyll data from [Chlorophyll-a, Aqua MODIS, NPP, L3SMI, Global, 4km, Science Quality, 2003- present (Monthly Composite)](https://coastwatch.pfeg.noaa.gov/erddap/griddap/erdMH1chlamday.html). Units are mg m-3 .


```{r eval = params$download_data}
# Download all global  chlorp data

missing_dates_2 <- c()

download_errdap_data_wrapper(dataset_name = "erdMH1chlamday",
                             variables = c("chlorophyll"),
                             temporal_resolution = "month",
                             date_start = "2003-01-16",
                             date_end = "2022-05-16",
                             data_grid = data_grid,
                             missing_dates = missing_dates_2,
                             run_parallel = TRUE)
```

```{r eval = params$aggregate_data}
# Are there missing dates from our downloads?
missing_dates <- determine_missing_dates(dataset_name = "erdMH1chlamday",
                                         temporal_resolution = "month",
                                         date_start = "2003-01-16",
                                         date_end = "2022-05-16")

missing_dates_2 <- missing_dates[[1]]$date

```

Similar to the SST data, this CHL data does not appear to be available to download through rerrdap as of July 9, 2025 (thanks Trump...). Here is code I will run to update the data for the remaining years we need to download (just 2024):

First, download the data manually from the NASA Ocean Color website, for L3 browser: https://oceandata.sci.gsfc.nasa.gov/l3/
You can give it a date range to download. For us, we will download monthly files from June 2022 to December 2024. Our old data stops in May 2022. You must have an account with Earthdata to do this. 

Once downloaded, we will read in the netcdf files and save as a csv, in the same format as what we did for when we could download directly from ERRDAP...

```{r}
# we need to prep the netcdf data to be a csv with zlev, latitude, longitude, sst, and anom. 

# we need to prep the netcdf data to be a csv with zlev, latitude, longitude, sst, and anom. 
files_to_run <- list.files(file.path(data_directory, "erddap/erdMH1chlamday_raw"), full.names = TRUE)

for(i in 2:length(files_to_run)){
# Open the NetCDF file

#   i = 1
  file_path <- files_to_run[[i]]
  
# Open the NetCDF file
nc_data <- ncdf4::nc_open(file_path)

# Extract coordinate variables
lon <- ncvar_get(nc_data, "lon")
lat <- ncvar_get(nc_data, "lat")

# Extract data variables
chl_data <- ncvar_get(nc_data, "chlor_a")

# Close the NetCDF file
nc_close(nc_data)

# Create a data frame with all combinations of coordinates
# Note: NetCDF arrays are in [lon, lat] order
cat("\nCreating coordinate grid...\n")

# Create coordinate grids using expand.grid for proper pairing
coord_grid <- expand.grid(longitude = lon, latitude = lat)

# Flatten the data arrays (they are 2D: [lon, lat])
# as.vector reads column-wise, which matches expand.grid order
chl_flat <- as.vector(chl_data)

# Create the data frame
df <- data.frame(
  latitude = coord_grid$latitude,
  longitude = coord_grid$longitude,
  chlorophyll = chl_flat) %>%
  filter(!is.na(chlorophyll))

# Write to CSV
#AQUA_MODIS.20241201_20241231.L3m.MO.CHL.chlor_a.4km.nc
file_date <- as.Date(str_after_last(str_before_nth(basename(file_path), "_", 2), "\\."), format = "%Y%m%d")
output_file <- glue("erdMH1chlamday_{file_date}.csv")
write.csv(df, file.path(data_directory, "erddap/erdMH1chlamday", output_file), row.names = FALSE)
}

```

