---
title: "Data wrangling - errdap"
author: "Gage Clawson (IMAS)"
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output: 
  pdf_document: 
    number_sections: yes
    toc: true
    toc_depth: 4
editor_options: 
  chunk_output_type: console
---

# Summary 

Spatially and temporally aggregate errdap data to 0.5 and 1 degree grids

```{r echo = FALSE}
# This chunk sets up default settings for all chunks below
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,fig.width = 7.5,fig.height = 5,dev = 'png',dpi=300)
```

```{r include=FALSE}
# Load all necessary packages
library(tidyverse)
library(sf)
library(stars)
library(rerddap)
library(rnaturalearth)
library(glue)
library(scico)
library(lubridate)
library(furrr)
library(here)
library(terra)
library(ncdf4)
library(strex)
source(here::here("prep/02_prep_spatial_layers_model/_functions_data_wrangling_erddap.R"))
source(here("R/dir.R"))

# Set the data directory. This specifies the folder in our drive where the data can be found. 

data_directory <- rdsi_raw_dir

# Get high-res ocean data
ocean <- ne_download(scale = 50, type = 'ocean', category = 'physical',returnclass = "sf") %>%
  dplyr::select(geometry)
```

# Defining our global grid

We'll start by aggregating to a 0.5 by 0.5 degree grid

```{r}
pixel_size <- 0.5

# Start with global ocean
starting_shape <- ocean
```

Only need to run once: 

```{r eval = params$download_data}
# Make a grid using desired pixel size
data_grid <- starting_shape %>%
  make_grid_custom(pixel_size = pixel_size)

data_grid %>%
  # Add geometry wkt column for saving this as a csv
  mutate(geometry_wkt = st_as_text(geometry)) %>% 
  # Make geometry point, grab lon/lat in lower left-hand corner
  # We'll use this later for joining to GFW data
  st_cast("POINT") %>%
  dplyr::mutate(lon = sf::st_coordinates(.)[,1],
                lat = sf::st_coordinates(.)[,2])%>%
  st_set_geometry(NULL)  %>% 
  group_by(pixel_id,geometry_wkt) %>% 
  summarize(lon = min(lon),
            lat = min(lat))%>%
  ungroup()  %>%
  as_tibble() %>%
  write_csv(here::here("data/model_features/global_grid.csv"))
```

We define a global grid using pixels that are `r pixel_size` degree latitude by `r pixel_size` degree longitude. We will then aggregate all of our model feature data to these `r pixel_size`x`r pixel_size` degree pixels. We include only pixels that overlap with the ocean. Each pixel is assigned a static `pixel_id`.

```{r}
data_grid <- data.table::fread(here::here("data/model_features/global_grid.csv")) %>%
  st_as_sf(wkt = "geometry_wkt",
           crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

# test <- data_grid %>%
#   st_drop_geometry() %>%
#   dplyr::select(lon, lat, pixel_id) %>%
#   rast(., type = "xyz") # 0.5 by 0.5.. cool it worked.
# plot(test)

```


## SST and SST anomaly data 

We use 0.25 degree data from [SST, Daily Optimum Interpolation (OI), AVHRR Only, Version 2.1, Final, Global, 0.25°, 1981-present, Lon+/-180](https://coastwatch.pfeg.noaa.gov/erddap/info/ncdcOisst21Agg_LonPM180/index.html). These data download ~25x as fast and are much faster to spatially aggregate.

```{r eval = params$aggregate_data}
# Spatially aggregate SST and SST anomaly data
spatially_aggregate_errdap_data_wrapper(dataset_name = "ncdcOisst21Agg_LonPM180",
                                        spatial_aggregation = data_grid,
                                        years = 2014:2024,  # really only need these years
                                        run_parallel = TRUE)
```

```{r eval = params$aggregate_data}
# Spatially aggregate SST and SST anomaly data

temporally_aggregate_errdap_data_wrapper(dataset_name = "ncdcOisst21Agg_LonPM180",
                                         temporal_aggregation = "year",
                                         run_parallel = TRUE,
                                         years_per_chunk = 5)

```


# Chlorophyll

We get monthly chlorophyll data from [Chlorophyll-a, Aqua MODIS, NPP, L3SMI, Global, 4km, Science Quality, 2003- present (Monthly Composite)](https://coastwatch.pfeg.noaa.gov/erddap/griddap/erdMH1chlamday.html). Units are mg m-3 .


```{r eval = params$aggregate_data}
# Spatially aggregate chl data
spatially_aggregate_errdap_data_wrapper(dataset_name = "erdMH1chlamday",
                                        spatial_aggregation = data_grid,
                                        years = 2014:2024, # only need these dates 
                                        run_parallel = TRUE)
```

```{r eval = params$aggregate_data}
# Spatially aggregate SST and SST anomaly data


temporally_aggregate_errdap_data_wrapper(dataset_name = "erdMH1chlamday",
                                         temporal_aggregation = "year",
                                         run_parallel = TRUE,
                                         years_per_chunk = 5)

```


# Now we prep the same data to a 1 degree by 1 degree grid

## Defining our global grid

```{r}
pixel_size <- 1

# Create polygon rectangle of globe
# This will serve as basis of grid
global_polygon_sf <-
  tibble(lon = c(-180,-180,180,180,-180),
       lat = c(-90,90,90,-90,-90))%>%
  as.matrix() %>%
  list(.) %>%
  st_polygon() %>%
  st_sfc(crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs") %>%
  st_as_sf()

# Start with global ocean
starting_shape <- ocean

data_grid <- data.table::fread(here::here("data/model_features/deg_1_x_1/global_grid.csv")) %>%
  st_as_sf(wkt = "geometry_wkt",
           crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs") %>%
              mutate(pixel_area_m2 = sf::st_area(geometry_wkt) %>%
                 units::drop_units())

```

Prep 2015-2024 SST, CHL, data, and save

```{r}
## start with prepping 1 by 1 sst data for 2015 only. Then we will appead it onto the existing data downloaded from the paper
# Spatially aggregate SST and SST anomaly data
spatially_aggregate_errdap_data_wrapper(dataset_name = "ncdcOisst21Agg_LonPM180",
                                        spatial_aggregation = data_grid,
                                        years = 2016:2024, # choose years you want to run
                                        run_parallel = TRUE)


# temporally aggregate SST and SST anomaly data

temporally_aggregate_errdap_data_wrapper(dataset_name = "ncdcOisst21Agg_LonPM180",
                                         temporal_aggregation = "year",
                                         run_parallel = TRUE,
                                         years_per_chunk = 5)

## now lets merge them all together
files <- list.files(here("data/model_features/deg_1_x_1/ncdcOisst21Agg_LonPM180/"), full.names = TRUE)

all_sst <- lapply(files, read.csv) %>%
  bind_rows() 

write.csv(all_sst, here("data/model_features/deg_1_x_1/errdap_sst.csv"), row.names = FALSE)


# Spatially aggregate chl data for 2015
spatially_aggregate_errdap_data_wrapper(dataset_name = "erdMH1chlamday",
                                        spatial_aggregation = data_grid,
                                        years = 2016:2024, # choose years you want to run
                                        run_parallel = TRUE)

temporally_aggregate_errdap_data_wrapper(dataset_name = "erdMH1chlamday",
                                         temporal_aggregation = "year",
                                         run_parallel = TRUE,
                                         years_per_chunk = 5)
## now lets merge them all together
files <- list.files(here("data/model_features/deg_1_x_1/erdMH1chlamday/"), full.names = TRUE)

all_chl <- lapply(files, read.csv) %>%
  bind_rows() 

write.csv(all_chl, here("data/model_features/deg_1_x_1/errdap_chl.csv"), row.names = FALSE)


```

