---
title: "Data wrangling - errdap"
author: "Gage Clawson (IMAS)"
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output: 
  pdf_document: 
    number_sections: yes
    toc: true
    toc_depth: 4
editor_options: 
  chunk_output_type: console
---

# Summary 

Spatially and temporally aggregate errdap data to 0.5 and 1 degree grids

```{r echo = FALSE}
# This chunk sets up default settings for all chunks below
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,fig.width = 7.5,fig.height = 5,dev = 'png',dpi=300)
```

```{r include=FALSE}
# Load all necessary packages
library(tidyverse)
library(sf)
library(stars)
library(rerddap)
library(rnaturalearth)
library(glue)
library(scico)
library(lubridate)
library(furrr)
library(here)
library(terra)
library(ncdf4)
library(strex)
library(FNN)

source(here::here("prep/02_prep_spatial_layers_model/_functions_data_wrangling_erddap.R"))
source(here("R/dir.R"))

# Set the data directory. This specifies the folder in our drive where the data can be found. 

data_directory <- rdsi_raw_dir

# Get high-res ocean data
ocean <- ne_download(scale = 50, type = 'ocean', category = 'physical',returnclass = "sf") %>%
  dplyr::select(geometry)
```

# Defining our global grid

We'll start by aggregating to a 0.5 by 0.5 degree grid

```{r}
pixel_size <- 0.5

# Start with global ocean
starting_shape <- ocean
```

Only need to run once: 

```{r eval = params$download_data}
# Make a grid using desired pixel size
data_grid <- starting_shape %>%
  make_grid_custom(pixel_size = pixel_size)

data_grid %>%
  # Add geometry wkt column for saving this as a csv
  mutate(geometry_wkt = st_as_text(geometry)) %>% 
  # Make geometry point, grab lon/lat in lower left-hand corner
  # We'll use this later for joining to GFW data
  st_cast("POINT") %>%
  dplyr::mutate(lon = sf::st_coordinates(.)[,1],
                lat = sf::st_coordinates(.)[,2])%>%
  st_set_geometry(NULL)  %>% 
  group_by(pixel_id,geometry_wkt) %>% 
  summarize(lon = min(lon),
            lat = min(lat))%>%
  ungroup()  %>%
  as_tibble() %>%
  write_csv(here::here("data/model_features/global_grid.csv"))
```

We define a global grid using pixels that are `r pixel_size` degree latitude by `r pixel_size` degree longitude. We will then aggregate all of our model feature data to these `r pixel_size`x`r pixel_size` degree pixels. We include only pixels that overlap with the ocean. Each pixel is assigned a static `pixel_id`.

```{r}
data_grid <- data.table::fread(here::here("data/model_features/global_grid.csv")) %>%
  st_as_sf(wkt = "geometry_wkt",
           crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

# test <- data_grid %>%
#   st_drop_geometry() %>%
#   dplyr::select(lon, lat, pixel_id) %>%
#   rast(., type = "xyz") # 0.5 by 0.5.. cool it worked.
# plot(test)

```


## SST and SST anomaly data 

We use 0.25 degree data from [SST, Daily Optimum Interpolation (OI), AVHRR Only, Version 2.1, Final, Global, 0.25°, 1981-present, Lon+/-180](https://coastwatch.pfeg.noaa.gov/erddap/info/ncdcOisst21Agg_LonPM180/index.html). These data download ~25x as fast and are much faster to spatially aggregate.

```{r eval = params$aggregate_data}
# Spatially aggregate SST and SST anomaly data
spatially_aggregate_errdap_data_wrapper(dataset_name = "ncdcOisst21Agg_LonPM180",
                                        spatial_aggregation = data_grid,
                                        years = 2014:2024,  # really only need these years
                                        run_parallel = TRUE)
```

```{r eval = params$aggregate_data}
# Spatially aggregate SST and SST anomaly data

temporally_aggregate_errdap_data_wrapper(dataset_name = "ncdcOisst21Agg_LonPM180",
                                         temporal_aggregation = "year",
                                         run_parallel = TRUE,
                                         years_per_chunk = 5)

```

There are cells which are missing SST information. For these, we should fill in with nearest neighbor approach 

```{r}
  sst_data_raw <- lapply(list.files(here("data/model_features/ncdcOisst21Agg_LonPM180"), full.names = TRUE), read.csv) %>%
        bind_rows()

sst_data <- sst_data_raw %>%
  dplyr::select(-anom_sst_c_mean, -anom_sst_c_sd) %>%
  filter(year %in% c(2014:2024)) %>%
    mutate(sst_c_mean = ifelse(sst_c_mean == 0 & is.na(sst_c_sd), NA, sst_c_mean)) %>%
  left_join(data_grid %>% st_drop_geometry())


missing_data <- sst_data %>%
  filter(is.na(sst_c_mean)|is.na(sst_c_sd)) %>%
  mutate(sst_c_mean = ifelse(sst_c_mean == 0, NA, sst_c_mean))

sst_missing <- sst_data %>% filter(is.na(sst_c_mean) | is.na(sst_c_sd))

sst_complete <- sst_data %>% filter(!is.na(sst_c_mean) & !is.na(sst_c_sd))

nrow(sst_missing) + nrow(sst_complete) == nrow(sst_data) # TRUE


# 3. Function to fill in missing values for each year
fill_missing_sst <- function(missing_df, complete_df) {
  filled <- list()
  
  for (yr in unique(missing_df$year)) {
    miss_year <- missing_df %>% filter(year == yr)
    comp_year <- complete_df %>% filter(year == yr)
    
    if (nrow(miss_year) == 0 || nrow(comp_year) == 0) next
    
    # Coordinates for matching
    miss_coords <- miss_year %>% select(lon, lat) %>% as.matrix()
    comp_coords <- comp_year %>% select(lon, lat) %>% as.matrix()
    
    # Nearest neighbor match
    nn <- get.knnx(comp_coords, miss_coords, k = 1)
    
    # Add values from nearest neighbor
    miss_year$sst_c_mean <- comp_year$sst_c_mean[nn$nn.index[,1]]
    miss_year$sst_c_sd <- comp_year$sst_c_sd[nn$nn.index[,1]]
    
    filled[[as.character(yr)]] <- miss_year
  }
  
  bind_rows(filled)
}

# 4. Apply filling function
sst_filled <- fill_missing_sst(sst_missing, sst_complete)

# 5. Combine filled + complete
sst_data_imputed <- bind_rows(sst_complete, sst_filled) %>%
  dplyr::select(-lon, -lat)

# check
sum(is.na(sst_data_imputed$sst_c_mean))  # Should be 0

nrow(sst_data_imputed) == nrow(sst_data) # TRUE

## save 

write.csv(sst_data_imputed, here("data/model_features/errdap_sst.csv"), row.names = FALSE) # only 85 mb 

```



# Chlorophyll

We get monthly chlorophyll data from [Chlorophyll-a, Aqua MODIS, NPP, L3SMI, Global, 4km, Science Quality, 2003- present (Monthly Composite)](https://coastwatch.pfeg.noaa.gov/erddap/griddap/erdMH1chlamday.html). Units are mg m-3 .


```{r eval = params$aggregate_data}
# Spatially aggregate chl data
spatially_aggregate_errdap_data_wrapper(dataset_name = "erdMH1chlamday",
                                        spatial_aggregation = data_grid,
                                        years = 2014:2024, # only need these dates 
                                        run_parallel = TRUE)
```

```{r eval = params$aggregate_data}
# Spatially aggregate SST and SST anomaly data


temporally_aggregate_errdap_data_wrapper(dataset_name = "erdMH1chlamday",
                                         temporal_aggregation = "year",
                                         run_parallel = TRUE,
                                         years_per_chunk = 5)

```

Now fill in any missing chl data

```{r}
chl_data_raw <- lapply(list.files(here("data/model_features/erdMH1chlamday"), full.names = TRUE), read.csv) %>%
        bind_rows()

chl_data <- chl_data_raw %>%
  filter(year %in% c(2014:2024)) %>%
  mutate(chl_mg_per_m3_mean = ifelse(chl_mg_per_m3_mean == 0 & is.na(chl_mg_per_m3_sd), NA, chl_mg_per_m3_mean)) %>%
  mutate(chl_mg_per_m3_sd = ifelse(!is.na(chl_mg_per_m3_mean) & is.na(chl_mg_per_m3_sd), 0, chl_mg_per_m3_sd)) # if sd is NA but there is a mean, make it 0, meaning there was no variation.

all_pixels <- data_grid %>% st_drop_geometry() %>% select(pixel_id)
all_years <- chl_data %>% distinct(year)
pixel_year_grid <- crossing(all_pixels, all_years)

chl_data_full <- pixel_year_grid %>%
  left_join(chl_data, by = c("pixel_id", "year")) %>%
  left_join(data_grid %>% st_drop_geometry())

missing_data <- chl_data_full %>%
  filter(is.na(chl_mg_per_m3_mean)|is.na(chl_mg_per_m3_sd))


chl_missing <- chl_data_full %>% filter(is.na(chl_mg_per_m3_mean) | is.na(chl_mg_per_m3_sd))

chl_complete <- chl_data_full %>% filter(!is.na(chl_mg_per_m3_mean) & !is.na(chl_mg_per_m3_sd))

nrow(chl_missing) + nrow(chl_complete) == nrow(chl_data_full) # TRUE


# 3. Function to fill in missing values for each year
fill_missing_chl <- function(missing_df, complete_df) {
  filled <- list()
  
  for (yr in unique(missing_df$year)) {
    miss_year <- missing_df %>% filter(year == yr)
    comp_year <- complete_df %>% filter(year == yr)
    
    if (nrow(miss_year) == 0 || nrow(comp_year) == 0) next
    
    # Coordinates for matching
    miss_coords <- miss_year %>% select(lon, lat) %>% as.matrix()
    comp_coords <- comp_year %>% select(lon, lat) %>% as.matrix()
    
    # Nearest neighbor match
    nn <- get.knnx(comp_coords, miss_coords, k = 1)
    
    # Add values from nearest neighbor
    miss_year$chl_mg_per_m3_mean <- comp_year$chl_mg_per_m3_mean[nn$nn.index[,1]]
    miss_year$chl_mg_per_m3_sd <- comp_year$chl_mg_per_m3_sd[nn$nn.index[,1]]
    
    filled[[as.character(yr)]] <- miss_year
  }
  
  bind_rows(filled)
}

# 4. Apply filling function
chl_filled <- fill_missing_chl(chl_missing, chl_complete)

# 5. Combine filled + complete
chl_data_imputed <- bind_rows(chl_complete, chl_filled) %>%
  dplyr::select(-lon, -lat)

# Optional check
sum(is.na(chl_data_imputed$chl_mg_per_m3_mean))  # Should be 0

nrow(chl_data_imputed) == nrow(chl_data_full) # TRUE

# ## save 
write.csv(chl_data_imputed, here("data/model_features/errdap_chl.csv"), row.names = FALSE) # 88 mb

```

# Now we prep the same data to a 1 degree by 1 degree grid

## Defining our global grid

```{r}
pixel_size <- 1

# Create polygon rectangle of globe
# This will serve as basis of grid
global_polygon_sf <-
  tibble(lon = c(-180,-180,180,180,-180),
       lat = c(-90,90,90,-90,-90))%>%
  as.matrix() %>%
  list(.) %>%
  st_polygon() %>%
  st_sfc(crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs") %>%
  st_as_sf()

# Start with global ocean
starting_shape <- ocean

data_grid <- data.table::fread(here::here("data/model_features/deg_1_x_1/global_grid.csv")) %>%
  st_as_sf(wkt = "geometry_wkt",
           crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs") %>%
              mutate(pixel_area_m2 = sf::st_area(geometry_wkt) %>%
                 units::drop_units())

```

Prep 2015-2024 SST, CHL, data, and save

```{r}
## start with prepping 1 by 1 sst data for 2015 only. Then we will appead it onto the existing data downloaded from the paper
# Spatially aggregate SST and SST anomaly data
spatially_aggregate_errdap_data_wrapper(dataset_name = "ncdcOisst21Agg_LonPM180",
                                        spatial_aggregation = data_grid,
                                        years = 2016:2024, # choose years you want to run
                                        run_parallel = TRUE)


# temporally aggregate SST and SST anomaly data

temporally_aggregate_errdap_data_wrapper(dataset_name = "ncdcOisst21Agg_LonPM180",
                                         temporal_aggregation = "year",
                                         run_parallel = TRUE,
                                         years_per_chunk = 5)

```

There are cells which are missing SST information. For these, we should fill in with nearest neighbor approach 

```{r}
  sst_data_raw <- lapply(list.files(here("data/model_features/deg_1_x_1/ncdcOisst21Agg_LonPM180"), full.names = TRUE), read.csv) %>%
        bind_rows()

sst_data <- sst_data_raw %>%
  dplyr::select(-anom_sst_c_mean, -anom_sst_c_sd) %>%
  filter(year %in% c(2014:2024)) %>%
    mutate(sst_c_mean = ifelse(sst_c_mean == 0 & is.na(sst_c_sd), NA, sst_c_mean)) %>%
  left_join(data_grid %>% st_drop_geometry())


missing_data <- sst_data %>%
  filter(is.na(sst_c_mean)|is.na(sst_c_sd)) %>%
  mutate(sst_c_mean = ifelse(sst_c_mean == 0, NA, sst_c_mean))

sst_missing <- sst_data %>% filter(is.na(sst_c_mean) | is.na(sst_c_sd))

sst_complete <- sst_data %>% filter(!is.na(sst_c_mean) & !is.na(sst_c_sd))

nrow(sst_missing) + nrow(sst_complete) == nrow(sst_data) # TRUE


# 3. Function to fill in missing values for each year
fill_missing_sst <- function(missing_df, complete_df) {
  filled <- list()
  
  for (yr in unique(missing_df$year)) {
    miss_year <- missing_df %>% filter(year == yr)
    comp_year <- complete_df %>% filter(year == yr)
    
    if (nrow(miss_year) == 0 || nrow(comp_year) == 0) next
    
    # Coordinates for matching
    miss_coords <- miss_year %>% select(lon, lat) %>% as.matrix()
    comp_coords <- comp_year %>% select(lon, lat) %>% as.matrix()
    
    # Nearest neighbor match
    nn <- get.knnx(comp_coords, miss_coords, k = 1)
    
    # Add values from nearest neighbor
    miss_year$sst_c_mean <- comp_year$sst_c_mean[nn$nn.index[,1]]
    miss_year$sst_c_sd <- comp_year$sst_c_sd[nn$nn.index[,1]]
    
    filled[[as.character(yr)]] <- miss_year
  }
  
  bind_rows(filled)
}

# 4. Apply filling function
sst_filled <- fill_missing_sst(sst_missing, sst_complete)

# 5. Combine filled + complete
sst_data_imputed <- bind_rows(sst_complete, sst_filled) %>%
  dplyr::select(-lon, -lat, -pixel_area_m2)

# check
sum(is.na(sst_data_imputed$sst_c_mean))  # Should be 0

nrow(sst_data_imputed) == nrow(sst_data) # TRUE

## save 

write.csv(sst_data_imputed, here("data/model_features/deg_1_x_1/errdap_sst.csv"), row.names = FALSE) # 30 mb 

```




```{r}
# Spatially aggregate chl data
spatially_aggregate_errdap_data_wrapper(dataset_name = "erdMH1chlamday",
                                        spatial_aggregation = data_grid,
                                        years = 2016:2024, # choose years you want to run
                                        run_parallel = TRUE)

temporally_aggregate_errdap_data_wrapper(dataset_name = "erdMH1chlamday",
                                         temporal_aggregation = "year",
                                         run_parallel = TRUE,
                                         years_per_chunk = 5)
# ## now lets merge them all together
# files <- list.files(here("data/model_features/deg_1_x_1/erdMH1chlamday/"), full.names = TRUE)
# 
# all_chl <- lapply(files, read.csv) %>%
#   bind_rows() 
# 
# write.csv(all_chl, here("data/model_features/deg_1_x_1/errdap_chl.csv"), row.names = FALSE)


```


Now fill in any missing chl data

```{r}
chl_data_raw <- lapply(list.files(here("data/model_features/deg_1_x_1/erdMH1chlamday"), full.names = TRUE), read.csv) %>%
        bind_rows()

chl_data <- chl_data_raw %>%
  filter(year %in% c(2014:2024)) %>%
  mutate(chl_mg_per_m3_mean = ifelse(chl_mg_per_m3_mean == 0 & is.na(chl_mg_per_m3_sd), NA, chl_mg_per_m3_mean)) %>%
  mutate(chl_mg_per_m3_sd = ifelse(!is.na(chl_mg_per_m3_mean) & is.na(chl_mg_per_m3_sd), 0, chl_mg_per_m3_sd)) # if sd is NA but there is a mean, make it 0, meaning there was no variation.

all_pixels <- data_grid %>% st_drop_geometry() %>% select(pixel_id)
all_years <- chl_data %>% distinct(year)
pixel_year_grid <- crossing(all_pixels, all_years)

chl_data_full <- pixel_year_grid %>%
  left_join(chl_data, by = c("pixel_id", "year")) %>%
  left_join(data_grid %>% st_drop_geometry())

missing_data <- chl_data_full %>%
  filter(is.na(chl_mg_per_m3_mean)|is.na(chl_mg_per_m3_sd))


chl_missing <- chl_data_full %>% filter(is.na(chl_mg_per_m3_mean) | is.na(chl_mg_per_m3_sd))

chl_complete <- chl_data_full %>% filter(!is.na(chl_mg_per_m3_mean) & !is.na(chl_mg_per_m3_sd))

nrow(chl_missing) + nrow(chl_complete) == nrow(chl_data_full) # TRUE


# 3. Function to fill in missing values for each year
fill_missing_chl <- function(missing_df, complete_df) {
  filled <- list()
  
  for (yr in unique(missing_df$year)) {
    miss_year <- missing_df %>% filter(year == yr)
    comp_year <- complete_df %>% filter(year == yr)
    
    if (nrow(miss_year) == 0 || nrow(comp_year) == 0) next
    
    # Coordinates for matching
    miss_coords <- miss_year %>% select(lon, lat) %>% as.matrix()
    comp_coords <- comp_year %>% select(lon, lat) %>% as.matrix()
    
    # Nearest neighbor match
    nn <- get.knnx(comp_coords, miss_coords, k = 1)
    
    # Add values from nearest neighbor
    miss_year$chl_mg_per_m3_mean <- comp_year$chl_mg_per_m3_mean[nn$nn.index[,1]]
    miss_year$chl_mg_per_m3_sd <- comp_year$chl_mg_per_m3_sd[nn$nn.index[,1]]
    
    filled[[as.character(yr)]] <- miss_year
  }
  
  bind_rows(filled)
}

# 4. Apply filling function
chl_filled <- fill_missing_chl(chl_missing, chl_complete)

# 5. Combine filled + complete
chl_data_imputed <- bind_rows(chl_complete, chl_filled) %>%
  dplyr::select(-lon, -lat, -pixel_area_m2)

# Optional check
sum(is.na(chl_data_imputed$chl_mg_per_m3_mean))  # Should be 0

nrow(chl_data_imputed) == nrow(chl_data_full) # TRUE

# ## save 
write.csv(chl_data_imputed, here("data/model_features/deg_1_x_1/errdap_chl.csv"), row.names = FALSE) # 88 mb

```

