---
title: "Prep country level data indices for modelling"
author: "Gage Clawson (IMAS)"
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output: 
  pdf_document: 
    number_sections: yes
    toc: true
    toc_depth: 4
editor_options: 
  chunk_output_type: console
---

# Summary

Prep country level data indices for modelling
 - Pacific Decadal Oscillation index 
 - El Nino South Oscillation Index
 - World bank development regions
 - Global fishing index data 
 
```{r echo = FALSE}
# This chunk sets up default settings for all chunks below
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,fig.width = 7.5,fig.height = 5,dev = 'png',dpi=300)
```

```{r include=FALSE}
# Load all necessary packages
library(tidyverse)
library(sf)
library(glue)
library(rnaturalearth)
library(countrycode)
library(terra)
library(here)
library(janitor)
library(readxl)

options(scipen = 20)

source(here("R/dir.R"))

```

# Defining our global grid

```{r}
pixel_size <- 0.5

mollweide_projection <- "+proj=moll +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +units=m +no_defs"

# Read in data_grid, generated in data_wrangling.Rmd
# We will do spatial joining in Mollweide, for proper calculations
# Note that data grid still retains lat and lon columns, for joining with non-spatial tibbles
# For Mollweide, always wrap around dateline, then transform, then calculate areas at end

data_grid <- data.table::fread(here::here("data/model_features/global_grid.csv")) %>%
  st_as_sf(wkt = "geometry_wkt",
           crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs") %>%
              mutate(pixel_area_m2 = sf::st_area(geometry_wkt) %>%
                 units::drop_units())
```

Pacific Decadal Oscillation index 

https://www.ncei.noaa.gov/pub/data/cmb/ersst/v5/index/ersst.v5.pdo.dat

```{r}

# Pull and wrangle PDO data from NOAA
# Calculate annual mean and SD for our time period of interest

wrangle_pdo_data <- 
  read.table(url("https://www.ncei.noaa.gov/pub/data/cmb/ersst/v5/index/ersst.v5.pdo.dat"), skip = 2) %>%
      as_tibble() %>%
    rename(year = V1) %>%
    pivot_longer(-year) %>%
    mutate(month = stringr::str_remove_all(name,"V") %>%
             as.numeric() - 1,
           date = lubridate::ymd(glue::glue("{year}-{month}-1")),
           year = lubridate::year(date)) %>%
    dplyr::select(year,pdo_index = value) %>%
    filter(year >= 1950,
           year <= 2024) %>%
    mutate(pdo_index = ifelse(pdo_index == 99.99,NA_real_,pdo_index)) %>%
    group_by(year) %>%
  summarise(pdo_index_mean = mean(pdo_index, na.rm = TRUE),
            pdo_index_sd = sd(pdo_index, na.rm = TRUE)) %>%
  ungroup() #


write.csv(wrangle_pdo_data, here("data/model_features/pdo_index.csv"), row.names = FALSE)

```

El Nino South Oscillation Index
Download from:https://psl.noaa.gov/data/correlation/oni.data

N. Rayner et al., Global analyses of sea surface temperature, sea ice, and night marine air temperature since the late nineteenth century. J. Geophys. Res. Atmos. 108, 1–25 (2003).

```{r}
# Pull and wrangle ENSO data from NOAA
# https://psl.noaa.gov/data/correlation/oni.data
# Calculate annual mean and SD for our time period of interest
wrangle_enso_data <- 
  read.table(url("https://psl.noaa.gov/data/correlation/oni.data"),skip=1,nrows=75) %>%
    as_tibble() %>%
    rename(year = V1) %>%
    pivot_longer(-year) %>%
    mutate(month = stringr::str_remove_all(name,"V") %>%
             as.numeric() - 1,
           date = lubridate::ymd(glue::glue("{year}-{month}-1")),
           year = lubridate::year(date)) %>%
    dplyr::select(year,enso_index = value) %>%
    filter(year >= 1950,
           year <= 2024)%>%
    group_by(year) %>%
    summarize(across(everything(),list(mean = ~mean(.,na.rm=TRUE),
                                       sd = ~sd(.,na.rm=TRUE)))) %>%
    ungroup()

write.csv(wrangle_enso_data, here("data/model_features/enso_index.csv"), row.names = FALSE)

```

# World bank development regions

```{r}

eez_lookup_ids <- read.csv(here("data/model_features/eez/eez_lookup_fix.csv"))

eez_region_features <- eez_lookup_ids %>%
    mutate(eez_region_world_bank_7 = countrycode::countrycode(eez_sovereign,"iso3c","region"))%>%
    distinct(eez_sovereign, eez_region_world_bank_7, eez_id)  %>% 
  mutate(eez_region_world_bank_7 = case_when(
    eez_sovereign == "ESH" ~ "Middle East & North Africa", 
    eez_sovereign == "High seas" ~ "High seas", 
    eez_sovereign == "MYT" ~ "Sub-Saharan Africa",
    eez_sovereign == "REU" ~ "Sub-Saharan Africa",
    eez_sovereign == "RAA" ~ "Europe & Central Asia",
    eez_sovereign == "RAM" ~ "Europe & Central Asia",
    eez_sovereign == "WLF" ~ "East Asia & Pacific",
    TRUE ~ eez_region_world_bank_7
  ))

write.csv(eez_region_features, here("data/model_features/world_bank_regions.csv"), row.names = FALSE)

```

Global fishing index data 

```{r}
eez_lookup_ids <- read.csv(here("data/model_features/eez/eez_lookup.csv"))

# Custom function to get the mode
get_mode <- function(x) {
  ux <- na.omit(unique(x))
  ux[which.max(tabulate(match(x, ux)))]
}

gfi_raw <- readxl::read_xlsx("/home/ubuntu/data_storage/raw_data/Global Fishing Index 2021 - Data download V1.1/Global Fishing Index 2021 Data for Download V1.1.xlsx", sheet = 3, skip = 1)  %>% 
  dplyr::select(flag_fin  = "ISO Code", gov_score = "Governance capacity") %>%
  mutate(gov_score = as.factor(gov_score)) # make it into a categorical variable

eez_gfi <- eez_region_features %>%
  left_join(gfi_raw, by = c("eez_sovereign" = "flag_fin")) %>% # lets gapfill using the mode of the same world bank region
  group_by(eez_region_world_bank_7) %>%
  mutate(most_common_score = get_mode(gov_score)) %>%
  ungroup() %>%
  mutate(gov_score_new = ifelse(is.na(gov_score), most_common_score, gov_score)) %>%
  mutate(gov_score_new = gov_score_new - 1) %>% # for some reason its adding one?
  dplyr::select(flag_fin = eez_sovereign, gov_score = gov_score_new)

write.csv(eez_gfi, here("data/model_features/global_fishing_index_governance.csv"), row.names = FALSE)
```

