---
title: "Allocate country level data to GFW cells using proportions"
output: html_document
date: "2024-09-24"
---

# Setup

Load packages and directories

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(gfwr)
library(terra)
library(tidyverse)
library(here)
library(glue)
library(data.table)
library(janitor)
library(sf)
library(countrycode)
library(raster)
library(qs)

source(here("R/dir.R"))


```

# Summary

We take GFW data and calculate for each gear type, vessel length, flag country, and EEZ, the proportion of catch occurring in that cell out of the total for those categories. 

Then we have total amounts of those categories from Rousseau et al. and we can multiply the total amounts by the proportions to allocate catch in each cell.



# Data Sources 

## Global Fishing Watch

**Reference**:

**Downloaded**: 


# Methods 


```{r}

rousseau_gear_fix <- qs::qread(here("data/int/rousseau_gear_fix.qs"))

```

Allocate national fishing effort to cells by joining by geer type, vessel length, and flag country. We will start with Industrial only, since that is technically what GFW includes. 

 - Join by gear, vessel length, flag country
 - multiply cell level nv_prop, hours_prop, gt_prop by the total amounts contained in the Rousseau data
 - save 


```{r}

for(res in c("lowres", "highres")){
for(s in c("I", "APW", "UP")){
  for(y in 2012:2017){

  #   res = "lowres"
  # y = 2012
  # s =  "I"
  
    
    total_effort <- rousseau_gear_fix %>%
      filter(year == y,
             sector == s) %>%
            rename(p_imas = p, gt_imas = gt, nv_imas = nv)
  
    
    cell_effort_props <- qs::qread(glue(file.path(rdsi_dir, "prep/prorate_allocation/{res}/gfw_effort/all_effort_gear_length_props_{y}.qs")))
    
    ## now join total effort and cell props by gear, vesssel length, and flag country
    
    joined_effort <- cell_effort_props %>%
      dplyr::select(-year) %>%
      right_join(total_effort, by = c("length_category", "gear" = "gear_new", "flag_fin" = "country")) %>%
      rename(gfw_nv = nv) %>%
      mutate(
        # nom_regular_hours = prop_hours_cell*nom_active_hours,
        #      eff_regular_hours = prop_hours_cell*eff_active_hours,
             nom_fishing_hours = prop_fishing_hours_cell*nom_active_hours,
             eff_fishing_hours = prop_fishing_hours_cell*eff_active_hours,
             p = p_imas*prop_engine_power_kw_cell,
             gt = prop_tonnage_gt_cell*gt_imas, 
             nv = prop_nv_cell*nv_imas
             ) 
    
    
    sum(joined_effort$eff_fishing_hours, na.rm = TRUE)/sum(total_effort$eff_active_hours) # 0.4984574 missing ~50% of hours still... so  A LOT doesn't match for 2012. But 2012 is the worst year of coverage for GFW. We'll do some data checking after running everything to see how later years fare. 
  
    
    allocated_effort_save <- joined_effort %>%
      dplyr::select(x, y, flag_fin, flag_country_name = flag_name, gear, length_category, eez_id = iso_sov1, nom_fishing_hours, eff_fishing_hours, p, gt, nv, gfw_fishing_hours = fishing_hours, gfw_p = engine_power_kw_gfw, gfw_gt = tonnage_gt_gfw, gfw_nv, nom_active_hours_imas = nom_active_hours, eff_active_hours_imas = eff_active_hours, p_imas, gt_imas, nv_imas)
    
    
    # test <- allocated_effort_save %>%
    #   filter(is.na(cell_ll_lon)) 
    
    ## ok so this all looks ok to me
    
qs::qsave(allocated_effort_save, glue(file.path(rdsi_dir, "prep/prorate_allocation/{res}/prorate_effort_{s}_{y}.qs")))
    
    
    }
  }
}


```

Data checking

```{r}

## lets check Industrial for all years


for(s in c("I", "APW", "UP")){
  for(y in 2012:2017){


  # y = 2012
  # s =  "I"
  
    total_effort <- rousseau_gear_fix %>%
      filter(year == y,
             sector == s) %>%
            rename(p_imas = p, gt_imas = gt, nv_imas = nv)


    

allocated_effort <- qs::qread(glue(file.path(rdsi_dir, "prep/prorate_allocation/lowres/prorate_effort_{s}_{y}.qs")))
    
    print(sum(allocated_effort$eff_fishing_hours, na.rm = TRUE)/sum(total_effort$eff_active_hours)) # this is how much effort is included in prorate

    
  }
}

## Industrial 
# [1] 0.4984574, 2012
# [1] 0.59654, 2013
# [1] 0.6026659, 2014
# [1] 0.6180964, 2015
# [1] 0.645804, 2016
# [1] 0.9500397, 2017

## ok so 2017 is the best coverage (expected). AIS didn't really have great uptake until 2016, so GFW recommends not putting too much stock into 2012-2015

## Artisanal powered
# [1] 0.0218172, 2012
# [1] 0.1358585, 2013
# [1] 0.1055112, 2014
# [1] 0.07240682, 2015
# [1] 0.1103558, 2016
# [1] 0.1822191, 2017

## Unpowered
# [1] 0.0002325565, 2012
# [1] 0.003034169, 2013
# [1] 0.002224677, 2014
# [1] 0.003099035, 2015
# [1] 0.007503027, 2016
# [1] 0.007939641, 2017

## the artisanal and unpowered are obviously horrible because of no AIS coverage


```

Let's make some plots to explore what this actually looks like. 

```{r}
## 2017
## Plot industrial by gear type and vessel length 


# eez_rgns <- qread(here("data/raw/spatial/eez_rast_csv_0.01.qs"))
# eez_lookup <- read.csv(here("data/raw/spatial/eez_lookup.csv")) %>%
#   clean_names() # made in /R/spatial_files.R

allocated_effort_files <- list.files(file.path(rdsi_dir, "prep/prorate_allocation/lowres/"), pattern = "2017", full.names = TRUE)

allocated_effort_all <- lapply(allocated_effort_files, function(file) {
  # Read the data
  data <- qs::qread(file)
  
  # Extract sector from the filename (assuming the sector is before "_2017" in the filename)
  sector <- gsub(".*prorate_effort_(.*)_2017.*", "\\1", file)
  
  # Add the sector column
  data$sector <- sector
  
  return(data)
}) %>%
  bind_rows()


test <- allocated_effort_all %>%
  group_by(x, y) %>%
  summarise(nom_fishing_hours = sum(nom_fishing_hours, na.rm = TRUE)) %>%
  ungroup() 


test_prorate <- test %>%
  filter(!is.na(x)|!is.na(y)) %>%
  rast(., type = "xyz")

plot(log(test_prorate+1))




raw_gfw_2 <- qread(file.path(rdsi_raw_dir, "global_fishing_watch/apparent_fishing_hours_v2/v3_aggregated/all_effort_2017.qs"))

test_gfw_2 <- raw_gfw_2 %>%
  group_by(cell_ll_lon, cell_ll_lat) %>%
  summarise(fishing_hours = sum(fishing_hours, na.rm = TRUE)) %>%
  ungroup() %>%
  rast(., type = "xyz")

plot(log(test_gfw_2+1))





effort_by_length <- allocated_effort_all %>% 
  filter(!is.na(x) | !is.na(y)) %>%
  group_by(x, y, length_category) %>%
  summarise(nom_fishing_hours = sum(nom_fishing_hours, na.rm = TRUE)) %>%
  ungroup() %>%
  filter(nom_fishing_hours > 0)
  

p <- ggplot(effort_by_length, aes(x = x, y = y, fill = log(nom_fishing_hours +1))) +
  geom_tile() +
  scale_fill_gradientn(colors = rev(terrain.colors(10)), name = "Fishing Hours (logged)") +
  facet_wrap(~ length_category, nrow = 3) +
  coord_fixed() +
  labs(x = "Longitude", y = "Latitude") +
  theme_minimal()

ggsave(plot = p, here("prep/02_allocation_proportions/figs/prorate_by_length.png"), height = 12, width = 12, dpi = 300, bg = "white")

effort_by_gear <- allocated_effort_all %>% 
  filter(!is.na(x) | !is.na(y)) %>%
  group_by(x, y, gear) %>%
  summarise(nom_fishing_hours = sum(nom_fishing_hours, na.rm = TRUE)) %>%
  ungroup() %>%
    filter(nom_fishing_hours > 0)

  

p <- ggplot(effort_by_gear, aes(x = x, y = y, fill = log(nom_fishing_hours+1))) +
  geom_tile() +
  scale_fill_gradientn(colors = rev(terrain.colors(10)), name = "Fishing Hours (logged)") +
  facet_wrap(~ gear, nrow = 5) +
  coord_fixed() +
  labs(x = "Longitude", y = "Latitude") +
  theme_minimal()

ggsave(plot = p, here("prep/02_allocation_proportions/figs/prorate_by_gear.png"), height = 12, width = 12, dpi = 300, bg = "white")

effort_by_sector <- allocated_effort_all %>% 
  filter(!is.na(x) | !is.na(y)) %>%
  group_by(x, y, sector) %>%
  summarise(nom_fishing_hours = sum(nom_fishing_hours, na.rm = TRUE)) %>%
  ungroup() %>%
    filter(nom_fishing_hours > 0)

  

p <- ggplot(effort_by_sector, aes(x = x, y = y, fill = log(nom_fishing_hours+1))) +
  geom_tile() +
  scale_fill_gradientn(colors = rev(terrain.colors(10)), name = "Fishing Hours (logged)") +
  facet_wrap(~ sector, nrow = 5) +
  coord_fixed() +
  labs(x = "Longitude", y = "Latitude") +
  theme_minimal()

ggsave(plot = p, here("prep/02_allocation_proportions/figs/prorate_by_sector.png"), height = 12, width = 12, dpi = 300, bg = "white") # wow. these all look horrible! 


## look at higher res GFW data 

raw_gfw_files <- list.files(file.path(rdsi_raw_dir, "global_fishing_watch/apparent_fishing_hours_mmsi/v3/"), full.names = TRUE, pattern = "2017")
  

raw_gfw_df <- lapply(raw_gfw_files, data.table::fread) %>%
  bind_rows()

sum(raw_gfw_df$apparent_fishing_hours, na.rm = TRUE) # 30210310
sum(raw_gfw_2$fishing_hours, na.rm = TRUE) # 45165727

test_gfw_2 <- raw_gfw_df %>%
  group_by(x, y) %>%
  summarise(fishing_hours = sum(apparent_fishing_hours, na.rm = TRUE)) %>%
  ungroup() %>%
  rast(., type = "xyz")

plot(log(test_gfw_2+1)) 


```

