---
title: "Fit prediction model for fishing effort"
output: html_document
date: "2024-12-11"
---

# Summary 

We use the data compiled in the previous scripts to fit a RF regression model to predict the proportion of fishing effort occurring in cells. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library("qs")
library(betareg)
library(foreach)
library(doParallel)
library(here)
library(mgcv)
library(tictoc)
library(progress)
library(terra)
library(glue)
library(arrow)
library(strex)
library(broom)
library(randomForest)

source(here("R/dir.R"))


```

## Functions to prepare enviro and effort data  

```{r}

elnino <- read.csv(here("data/model_features/enso_index.csv"))

pdo <- read.csv(here("data/model_features/pdo_index.csv"))

world_bank <- read.csv(here("data/model_features/world_bank_regions.csv")) %>%
  filter(ISO_SOV1 != "GIB") # filter out gibralter bc it is duplicating the UK for some reason

gfi_df <- read.csv(here("data/model_features/global_fishing_index_governance.csv"))
  
# Load prepared data
model_data <- readRDS(here("data/model_features/prepared_data_1deg.rds")) %>%
  mutate(year = as.numeric(year)) # i think we need to adjust the prepared data to include cells which have no fishing effort in them. We can do this in the loop below? 

## read in environmental variables

  global_grid <- read.csv(here("data/model_features/deg_1_x_1/global_grid.csv"))
  
  ocean_data <- read.csv(here("data/model_features/deg_1_x_1/errdap_chl.csv")) %>%
  left_join(read.csv(here("data/model_features/deg_1_x_1/errdap_sst.csv")), by = c("pixel_id", "year")) %>% 
  left_join(read.csv(here("data/model_features/deg_1_x_1/remss_wind.csv")), by = c("pixel_id", "year")) %>%
  filter(year %in% c(2015:2024)) 


  spatial_data <- global_grid %>%
    left_join(read.csv(here("data/model_features/deg_1_x_1/gfw_static_spatial_measures.csv")) %>% dplyr::select(-lat, -lon), by = "pixel_id") %>%
   left_join(read.csv(here("data/model_features/deg_1_x_1/mesopelagiczones/mesopelagiczones_fixed.csv")), by = "pixel_id") %>%
   left_join(read.csv(here("data/model_features/deg_1_x_1/eez/eez.csv")), by = "pixel_id") %>%
   left_join(read.csv(here("data/model_features/deg_1_x_1/fao/fao_fixed.csv")), by = "pixel_id") %>%
   left_join(read.csv(here("data/model_features/deg_1_x_1/oceans_fixed.csv")), by = "pixel_id") %>%
    left_join(read.csv(here("data/model_features/deg_1_x_1/seamounts.csv")), by = "pixel_id") %>% 
   crossing(., year = c(2015:2024)) 
  
  
  env_data <- spatial_data %>%
    left_join(ocean_data) %>% # cool.
    dplyr::select(-geometry_wkt) %>%
    left_join(elnino) %>%
    left_join(pdo) %>%
    left_join(world_bank, by = c("eez_id")) %>%
    mutate(eez_region_world_bank_7 = ifelse(eez_sovereign %in% c("High seas", "Land"), "High seas", eez_region_world_bank_7)) %>% 
    left_join(gfi_df, by = c("eez_sovereign" = "flag_fin")) %>% # add in global fishing index data here
    mutate(gov_score = ifelse(eez_id >= 99999 & is.na(gov_score), "high_seas", gov_score)) %>%
        mutate(gov_score = ifelse(eez_id < 99999 & is.na(gov_score), "no_data", gov_score)) %>%
    dplyr::select(-eez_sovereign, -nearest_seamount_id) %>% 
    distinct() 

```

```{r}
hist_fish_data <- qs::qread(here("data/int/rousseau_gear_fix.qs")) %>%
    ## remove artisanal
    filter(sector == "I") %>%
    group_by(flag_fin = country, gear = gear_new, length_category) %>%
    summarize(
      total_fishing_hours = sum(eff_active_hours, na.rm = TRUE),
      .groups = 'drop'
    ) %>%
    dplyr::select(flag_fin, gear, length_category, total_fishing_hours) %>%
    filter(total_fishing_hours > 0)
  
flags <- unique(hist_fish_data$flag_fin) # get the flags we need to run models for # 151 of them

env_grid <- env_data %>% 
  dplyr::select(lon, lat) %>% distinct()

full_model_formula <- formula(
  prop_fishing_hours_cell ~ 
    # Categorical/factor predictors
    gear  + 
    length_category +
    meso_id +  
    eez_id + 
    fao_id + 
    ocean +  # Spatial categorical variables
    eez_region_world_bank_7 + # world bank regions
    gov_score + # global fishing index
    # Continuous predictors
    lon + lat + 
    elevation_m + # depth
    distance_from_port_m + 
    distance_from_shore_m +
    chl_mg_per_m3_mean + 
    chl_mg_per_m3_sd +
    sst_c_mean + 
    sst_c_sd +
    wind_speed_ms_mean + 
    wind_speed_ms_sd +
    enso_index_mean + # el nino data 
    enso_index_sd + # pacific decadal oscillation
    pdo_index_mean +  
    pdo_index_sd + 
    nearest_seamount_distance_m + 
    # Year effect
    year
)

```

## Run models and variable selection for every individual country

 - calculate the full model using all variables available per flag country
 - calculate variable importance metrics and root mean squared error (RMSE)
 - Set a threshold of the 10th quantile of the variable importance (RMSE for regression) and remove any variables with variable importance less than that
 - Rerun model with new variables and calculate model importance metrics and RMSE again
 - If the RMSE doesn't DECREASE at all, we stop the model pruning, if it does, we continue, hoping that the RMSE will improve even more in the next iteration. 
 - If the threshold doesn't remove any variables, we increase the threshold by 1% until a variable is removed, and rerun the process
 - We loop through this until the RMSE does not improve at all (improvement being a decrease in RMSE)
  - NOTE: Always keep year as a variable
 
 NOTE: Lower RMSE is better

NOTE: need to check on flags which have very little data. Some of the flags have only ~5 rows of data (flag == "SVN" for example only has 2 rows), so obviously the predictions for these would be horrible. Maybe we want to exclude these and gapfill with something later on? Maybe add a sample size filter? Check to see if these even ran for the classification. If they didn't, we probably won't include them anyways. 

Save full models first 

```{r}
flags <- unique(hist_fish_data$flag_fin) # get the flags we need to run models for # 167 of them

missing_flags <- setdiff(unique(hist_fish_data$flag_fin), unique(model_data$flag_fin))
length(missing_flags) #  these are flags that are in the rousseau data but not in the gfw data, meaning we can't make predictions on them.

flags <- setdiff(flags, missing_flags)

flags_to_run <- setdiff(flags, "")

for(flag in flags_to_run) {

#  flag = "PRK"
  
model_data_flag <- model_data %>%
  dplyr::select(lon, lat, flag_fin, gear, length_category, year, prop_fishing_hours_cell) %>%
  filter(flag_fin == flag) %>% 
  left_join(env_data) %>%
  na.omit() # removes any NAs

if(nrow(model_data_flag) == 1){
  next()
}

# prop_fishing_hours_cell = the proportion of that flag country's fishing effort represented in that cell (per gear, length, and year groupings; see below)

# test_prop <- model_data_flag %>% 
#   group_by(flag_fin, gear, length_category, year) %>%
#   summarise(prop = sum(prop_fishing_hours_cell)) %>%
#   ungroup() # all should be 1? Some of these aren't because of the weird NAs in the environmental data, and we exclude those cells. I guess it doesn't make too much of a difference as we will rescale the output of the model to 0-1 anyways after predictions. 

set.seed(123)
samp <- sample(nrow(model_data_flag), 0.6 * nrow(model_data_flag))  # do 60/40 split since this data is mostly small

train <- model_data_flag[samp, ]

test <- model_data_flag[-samp, ]

tic()
model <- randomForest(full_model_formula, data = train, type = "regression", proximity = FALSE, ntree = 100, importance = TRUE) 
toc() # half a second for NZL; 45 secs for CHN

# Initialize tracking
var_imp_i <- importance(model)
n_vars <- nrow(var_imp_i)

qs::qsave(model, glue(file.path(rdsi_dir, "prep/random_forest/stage_2_models/pruning/stage_2_rf_train_{flag}_{n_vars}.qs")))
}
```

Now apply variable selection methods 

```{r}
flags_to_run <- setdiff(flags, "")

  for(flag in flags_to_run){
    rmse_values <- c()

model_data_flag <- model_data %>%
  dplyr::select(lon, lat, flag_fin, gear, length_category, year, prop_fishing_hours_cell) %>%
  filter(flag_fin == flag) %>%
  left_join(env_data) %>%
  dplyr::select(-total_fishing_hours, -log_total_fishing_hours) %>%
  na.omit() 

if(nrow(model_data_flag) == 1){
  next()
}

# prop_fishing_hours_cell = the proportion of that flag country's fishing effort represented in that cell (per gear, length, and year groupings; see below)

# test_prop <- model_data_flag %>% 
#   group_by(flag_fin, gear, length_category, year) %>%
#   summarise(prop = sum(prop_fishing_hours_cell)) %>%
#   ungroup() # all should be 1? Some of these aren't because of the weird NAs in the environmental data, and we exclude those cells. I guess it doesn't make too much of a difference as we will rescale the output of the model to 0-1 anyways after predictions. 

set.seed(123)
samp <- sample(nrow(model_data_flag), 0.6 * nrow(model_data_flag))  # do 60/40 split since this data is mostly small

train <- model_data_flag[samp, ]

test <- model_data_flag[-samp, ]


if(!file.exists(glue(file.path(rdsi_dir, "prep/random_forest/stage_2_models/pruning/stage_2_rf_train_{flag}_25.qs")))){
tic()
model <- randomForest(full_model_formula, data = train, type = "regression", proximity = FALSE, ntree = 100, importance = TRUE) 
toc() # half a second for NZL; 45 secs for CHN
}else{
  
  model <- qs::qread(glue(file.path(rdsi_dir, "prep/random_forest/stage_2_models/pruning/stage_2_rf_train_{flag}_25.qs")))
  
}


# Initialize tracking
var_imp_i <- importance(model)
num_vars <- c(nrow(var_imp_i))
threshold_i <- quantile(var_imp_i[, 2], 0.10)  # Use the 10th percentile instead of a fixed multiplier
n_vars <- nrow(var_imp_i)

  if (threshold_i == 0) {
    threshold_i <- 0.0000001
  }

pred_props <- predict(model, newdata = test, type = "response")  # Probability of class 1
rmse <- sqrt(mean((test$prop_fishing_hours_cell - pred_props)^2)) # use RMSE for probabilities instead?

rmse_values <- c(rmse_values, rmse)


# Iteratively remove low-importance variables
iteration <- 1
while (TRUE) {
  
  # Select variables above threshold
    selected_vars <- unique(c("year", names(var_imp_i[, 4][var_imp_i[, 4] > threshold_i])))
  
  # Stop if too few variables remain
  if (length(selected_vars) < 6) break  # Avoid over-pruning. We can change this to any number of variables. Maybe 10 would be better computationally? 
  
          # Increase the threshold iteratively until at least one variable is removed
    while (length(selected_vars) == num_vars[length(num_vars)]) {
        threshold_i <- threshold_i * 1.01  # Increase threshold by 1%
        selected_vars <- unique(c("year", names(var_imp_i[, 4][var_imp_i[, 4] > threshold_i])))
    }

  # Refit model with selected variables
  model_i <- randomForest(prop_fishing_hours_cell ~ ., data = train[, c("prop_fishing_hours_cell", selected_vars)], 
                        type = "regression", proximity = FALSE, 
                        ntree = 100, importance = TRUE)

  
  # Get new variable importance
  var_imp_i <- importance(model_i)

  threshold_i <- quantile(var_imp_i[, 2], 0.1) # calculate new threshold since variable importance metric will change
  
  # Compute new RMSE
  pred_props_i <- predict(model_i, newdata = test, type = "response") 
  rmse_i <- sqrt(mean(test$prop_fishing_hours_cell - pred_props_i)^2)
  
  # Store metrics
  rmse_values <- c(rmse_values, rmse_i)
  num_vars <- c(num_vars, length(selected_vars))
  
  n_vars <- length(selected_vars)
  file_path <-   glue(file.path(rdsi_dir, "prep/random_forest/stage_2_models/pruning/stage_2_rf_train_{flag}_{n_vars}.qs"))
  
  if(!file.exists(file_path)){
    # save model here. Put number of variables (length(selected_vars)) in model save name so we know which one to pick for best predictions. 
  qs::qsave(model_i, file_path)
  }

  # Check RMSE stability: Stop if no improvement
  if (length(rmse_values) > 1) {
    
    rmse_improvement <- (rmse_values[length(rmse_values) - 1] - rmse_values[length(rmse_values)]) / rmse_values[length(rmse_values) - 1]
    
    if (rmse_improvement <= 0) break  # Stop if RMSE stabilizes
    
  }

  iteration <- iteration + 1
}

}


```

Now write code to select the "best" model per the variable selection from above. We will rerun the model with just those variables on the FULL dataset to leverage all of the data from GFW and save

```{r}
# grab the flags that were run in the folder
stage_2_path <- file.path(rdsi_dir, "prep/random_forest/stage_2_models/pruning")
stage_2_files <- list.files(stage_2_path, full.names = TRUE)
stage_2_flags <- unique(sub(".*stage_2_rf_train_([A-Z]{3})_.*", "\\1", stage_2_files)) # 117 flags 


for(flag in stage_2_flags) {

#  flag = "ZAF"
  flag_files <- list.files(stage_2_path, pattern = glue("_{flag}_"))
  # select the model with the next to lowest number of variables. E.g., if the final model run for ZAF is 17, then we want to grab the model with the next to lowest number of variables, which is 19. 
  n_variables <- as.numeric(str_before_first(str_after_last(flag_files, "_"), "\\."))
  
  if(length(n_variables) > 1){
  best_model_n <- as.character(sort(n_variables[2]))
  }else{
    best_model_n <- as.character(n_variables[1])
  }
  
  best_train_model <- qs::qread(file.path(stage_2_path, glue("stage_2_rf_train_{flag}_{best_model_n}.qs")))
  
  rf_formula <- as.formula(deparse(formula(best_train_model)) |> paste(collapse = " "))
  
model_data_flag <- model_data %>%
  dplyr::select(lon, lat, flag_fin, gear, length_category, year, prop_fishing_hours_cell) %>%
  filter(flag_fin == flag) %>% 
  left_join(env_data) %>%
  left_join(hist_fish_data) %>% 
  dplyr::select(-total_fishing_hours, -log_total_fishing_hours) %>%
  na.omit() # this removes the categories which are not in the Rousseau data. For example, Rousseau does not have data for NZL, Lines_Longlines, 24-50m, 2015, but GFW does. 

if(nrow(model_data_flag) == 1){
  next()
} # skip if not enough data, I don't think this is a worry here but just in case.


tic()
model <- randomForest(rf_formula, data = model_data_flag, type = "regression", proximity = FALSE, ntree = 100, importance = TRUE) 
toc() # half a second for NZL; 45 secs for CHN


qs::qsave(model, glue(file.path(rdsi_dir, "prep/random_forest/stage_2_models/stage_2_rf_model_full_data_{flag}_{best_model_n}.qs")))
}


```
