---
title: "Create fishing access dataset"
output: html_document
date: "2024-12-11"
---

# Summary

Here we create a dataset that described for each flag country, the EEZs they are able to fish in for each year. We derive this data from Rousseau et al, who derived it from Watson et al, who used fishing access rights datasets.

The data will be formatted as year \| flag country \| eez (including high seas?) \| fishing (1 or 0)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(countrycode)
library(sf)
library(qs)

source(here("R/dir.R"))
```

The data we need to access is located here: /home/ubuntu/gem/private/users/yannickr/effort_histsoc_1841_2017_EEZ_addFAO.csv

```{r}
effort_df <- read.csv("/home/ubuntu/gem/private/users/yannickr/effort_histsoc_1841_2017_EEZ_addFAO.csv")

colnames(effort_df)

eez_sf <- st_read(file.path(rdsi_raw_dir, "marine_regions/World_EEZ_v12_20231025"), layer = 'eez_v12')

eez_lookup_ids <- eez_sf %>% 
      filter(POL_TYPE !=  "Overlapping claim") %>% 
  st_drop_geometry() %>%
  dplyr::distinct(eez_sovereign = ISO_SOV1, iso_ter = ISO_TER1) %>%
        filter(eez_sovereign != "ATA")  %>% 
  add_row(eez_sovereign = "High seas", iso_ter = "High seas") %>%
  filter(!is.na(iso_ter))

# read in our eez lookup 

eez_lookup <- read.csv(here("data/model_features/deg_1_x_1/eez/eez_lookup.csv"))

## need SAUP country codes they used
saup_codes <- read.csv("https://data.imas.utas.edu.au/attachments/1241a51d-c8c2-4432-aa68-3d2bae142794/SAUPtoCountry.csv")


effort_access_df <- effort_df %>%
  filter(NomActive > 0, 
        Year >= 1950,
        Sector == "Industrial") %>%
  distinct(year = Year,  flag_id = SAUP, eez_country_name) %>% 
  left_join(saup_codes, by = c("flag_id" = "SAUP")) %>%
  dplyr::select(year, flag_fin = Country, eez_country_name) %>%
  mutate(iso_ter = countrycode(eez_country_name, origin = "country.name", destination = "iso3c")) %>%
  mutate(iso_ter = ifelse(eez_country_name == "High Seas", "High seas", iso_ter)) %>%
  left_join(eez_lookup_ids) %>%
    mutate(eez_country_name = ifelse(eez_country_name == "High Seas", "High seas", eez_country_name))  %>%
    mutate(eez_sovereign = case_when(
    eez_country_name == "Dominican Rp" ~ "DOM",
    eez_country_name == "NethAntilles" ~ "NLD",
    eez_country_name %in% c("US Virgin Is", "Amer Samoa") ~ "USA",
    eez_country_name %in% c("Fr Guiana", "Kerguelen Is", "Fr Polynesia") ~ "FRA", 
    eez_country_name %in% c("Channel Is", "Br Virgin Is", "Br Ind Oc Tr") ~ "GBR", 
    eez_country_name == "Serbia Montenegro" ~ "MNE", # choosing montenegro here as serbia is landlocked
    eez_country_name == "West Sahara" ~ "MAR", # this is technically not right? I guess it depends on the year 
    eez_country_name == "Untd Arab Em" ~ "ARE", 
    eez_country_name == "Micronesia" ~ "FSM",
    TRUE ~ eez_sovereign
  )) %>%
    mutate(access = 1) %>%
  mutate(eez_sovereign = case_when(
    eez_country_name == "Falkland Is" ~ "ARG",
    eez_country_name == "Gibraltar" ~ "GBR",
    eez_country_name == "Mayotte" ~ "FRA", 
    eez_country_name %in% c("Hong Kong", "Macau", "Taiwan") ~ "CHN", # contentious? 
    eez_country_name == "Ukraine" ~ "RUS", # this isnt true... 
    TRUE ~ eez_sovereign
  )) %>%
  distinct(year, flag_fin, eez_sovereign, access) %>%
  left_join(eez_lookup) 
  
qs::qsave(effort_access_df, here("data/int/prediction_historical_data/fishing_access.qs"))

```

Create access mask for high seas areas and use FAO areas as access areas for those

```{r}
effort_access_df_highseas <- effort_df %>%
  filter(NomActive > 0, 
        Year >= 1950,
        Sector == "Industrial",
        eez_country_name == "High Seas") %>%
  distinct(year = Year,  flag_id = SAUP, eez_country_name, fao_area) %>% # maybe add gear here? Vessel length too? For as much detail as we can get? 
  left_join(saup_codes, by = c("flag_id" = "SAUP")) %>%
  dplyr::select(year, flag_fin = Country, eez_country_name, fao_area) %>%
    mutate(eez_country_name = ifelse(eez_country_name == "High Seas", "High seas", eez_country_name))  %>%
    mutate(access_fao = 1) %>%
  distinct(year, flag_fin, eez_sovereign = eez_country_name, fao_area, access_fao) %>%
  left_join(eez_lookup) 

qs::qsave(effort_access_df_highseas, here("data/int/prediction_historical_data/high_seas_fao_fishing_access.qs"))


fao_lookup <- read.csv(here("data/model_features/fao/fao_major_ids.csv"))

antartica_access <- effort_access_df_highseas %>%
  distinct(year, flag_fin, eez_sovereign, fao_area, access_fao) %>%
  left_join(fao_lookup, by = c("fao_area" = "fao_id")) %>%
  filter(str_detect(name_en, "Antarctic"))
```

Try with catch data instead?

```{r}
catch_df <- read.csv("/home/ubuntu/gem/private/users/yannickr/catch_histsoc_1869_2017_EEZ_addFAO.csv") %>%
  filter(Sector == "industrial")

catch_access_df <- catch_df %>%
  mutate(catch = Reported + IUU + Discards) %>%
  filter(catch > 1, 
        Year >= 1950,
        SAUP != 999) %>%
  distinct(year = Year, SAUP, FCountryName, eez_country_name) %>%
  left_join(saup_codes, by = c("SAUP")) %>%
  mutate(Country = ifelse(SAUP == 830, "GBR", Country)) %>%
  dplyr::select(year, flag_fin = Country, eez_country_name) %>%
  filter(!is.na(flag_fin)) %>%
  mutate(iso_ter = countrycode(eez_country_name, origin = "country.name", destination = "iso3c")) %>%
  mutate(iso_ter = ifelse(eez_country_name == "High Seas", "High seas", iso_ter)) %>%
  left_join(eez_lookup_ids) %>%
    mutate(eez_country_name = ifelse(eez_country_name == "High Seas", "High seas", eez_country_name))  %>%
    mutate(eez_sovereign = case_when(
    eez_country_name == "Dominican Rp" ~ "DOM",
    eez_country_name == "NethAntilles" ~ "NLD",
    eez_country_name %in% c("US Virgin Is", "Amer Samoa") ~ "USA",
    eez_country_name %in% c("Fr Guiana", "Kerguelen Is", "Fr Polynesia") ~ "FRA", 
    eez_country_name %in% c("Channel Is", "Br Virgin Is", "Br Ind Oc Tr") ~ "GBR", 
    eez_country_name == "Serbia Montenegro" ~ "MNE", # choosing montenegro here as serbia is landlocked
    eez_country_name == "West Sahara" ~ "MAR", # this is technically not right? I guess it depends on the year 
    eez_country_name == "Untd Arab Em" ~ "ARE", 
    eez_country_name == "Micronesia" ~ "FSM",
    TRUE ~ eez_sovereign
  )) %>%
    mutate(access = 1) %>%
  mutate(eez_sovereign = case_when(
    eez_country_name == "Falkland Is" ~ "ARG",
    eez_country_name == "Gibraltar" ~ "GBR",
    eez_country_name == "Mayotte" ~ "FRA", 
    eez_country_name %in% c("Hong Kong", "Macau", "Taiwan") ~ "CHN", # contentious? 
    eez_country_name == "Ukraine" ~ "RUS", # this isnt true... 
    TRUE ~ eez_sovereign
  )) %>%
  distinct(year, flag_fin, eez_sovereign, access) %>%
  left_join(eez_lookup)

```

Fix Ukraine access data: It is not in the Watson data before 1988, so we will use the Rousseau effort data for those years
 - filter for hours of effort > 100 
 - create a master list of countries where Ukraine fishes in from the Watson data from 1990-2000
 - Only include countries in the rousseau data if they are in the watson data 

```{r}

ukraine_watson <- catch_access_df %>%
  filter(year %in% c(1990:2000),
         flag_fin == "UKR") %>%
  distinct(flag_fin, eez_sovereign, access, eez_id)

rousseau_access <- qread(here("data/int/prediction_historical_data/fishing_access.qs"))

ukr_rousseau_access <- effort_df %>% 
    mutate(iso_ter = countrycode(eez_country_name, origin = "country.name", destination = "iso3c")) %>%
  mutate(iso_ter = ifelse(eez_country_name == "High Seas", "High seas", iso_ter)) %>%
  left_join(eez_lookup_ids) %>%
    mutate(eez_country_name = ifelse(eez_country_name == "High Seas", "High seas", eez_country_name))  %>%
    mutate(eez_sovereign = case_when(
    eez_country_name == "Dominican Rp" ~ "DOM",
    eez_country_name == "NethAntilles" ~ "NLD",
    eez_country_name %in% c("US Virgin Is", "Amer Samoa") ~ "USA",
    eez_country_name %in% c("Fr Guiana", "Kerguelen Is", "Fr Polynesia") ~ "FRA", 
    eez_country_name %in% c("Channel Is", "Br Virgin Is", "Br Ind Oc Tr") ~ "GBR", 
    eez_country_name == "Serbia Montenegro" ~ "MNE", # choosing montenegro here as serbia is landlocked
    eez_country_name == "West Sahara" ~ "MAR", # this is technically not right? I guess it depends on the year 
    eez_country_name == "Untd Arab Em" ~ "ARE", 
    eez_country_name == "Micronesia" ~ "FSM",
    TRUE ~ eez_sovereign
  ))  %>%
  filter(NomActive > 0, 
        Year >= 1950,
        Sector == "Industrial") %>%
  rename(year = Year,  flag_id = SAUP) %>% 
  left_join(saup_codes, by = c("flag_id" = "SAUP")) %>%
  group_by(year, Country, eez_sovereign) %>%
  summarise(NomActive = sum(NomActive, na.rm = TRUE)) %>%
  ungroup() %>%
  dplyr::select(year, flag_fin = Country, eez_sovereign, NomActive) %>%
  filter(flag_fin == "UKR",
         NomActive > 100) %>% # maybe filter for >1000 hours when we have to use this? 
  distinct(year, flag_fin, eez_sovereign) %>%
  mutate(access = 1) %>%
  filter(year < 1988) 


ukr_access_1950_1987 <- ukr_rousseau_access %>%
  filter(eez_sovereign %in% unique(ukraine_watson$eez_sovereign)) %>%
  left_join(eez_lookup)


catch_access_fin <- catch_access_df %>%
  rbind(ukr_access_1950_1987)

qs::qsave(catch_access_fin, here("data/int/prediction_historical_data/catch_fishing_access.qs"))

```

```{r}
catch_access_df_highseas <- catch_df %>%
    mutate(catch = Reported + IUU + Discards) %>%
  filter(catch > 0, 
         Year >= 1950, 
         eez_country_name == "High Seas",
         Sector == "industrial") %>%
    left_join(saup_codes, by = c("SAUP")) %>%
    mutate(access_fao = 1,
           eez_country_name = "High seas") %>%
  left_join(fao_lookup, by = c("fao_area" = "fao_id")) %>%
  distinct(year = Year, flag_fin = Country, fao_area, fao_area_name = name_en, eez_country_name, access_fao)

ukraine_watson <- catch_access_df_highseas %>%
  filter(year %in% c(1990:2000),
         flag_fin == "UKR") %>%
  distinct(flag_fin, eez_country_name, fao_area, fao_area_name, access_fao)

# fix ukraine 
ukraine_effort_highseas <- effort_df %>%
  filter(NomActive > 0, 
        Year >= 1950,
        Sector == "Industrial",
        eez_country_name == "High Seas") %>%
  rename(year = Year,  flag_id = SAUP) %>%
  left_join(saup_codes, by = c("flag_id" = "SAUP")) %>%
  dplyr::select(year, flag_fin = Country, eez_country_name, fao_area, NomActive) %>%
    mutate(eez_country_name = ifelse(eez_country_name == "High Seas", "High seas", eez_country_name))  %>%
  group_by(year, flag_fin, eez_country_name, fao_area) %>%
  summarise(NomActive = sum(NomActive, na.rm = TRUE)) %>%
  ungroup() %>%
  dplyr::select(year, flag_fin, eez_country_name, fao_area, NomActive) %>%
  filter(flag_fin == "UKR",
         NomActive > 100) %>% # maybe filter for >1000 hours when we have to use this? 
  distinct(year, flag_fin, eez_country_name, fao_area) %>%
  mutate(access_fao = 1) %>%
  filter(year < 1988) %>%
  filter(fao_area %in% c(unique(ukraine_watson$fao_area))) %>%
  left_join(fao_lookup, by = c("fao_area" = "fao_id")) %>%
  rename(fao_area_name = name_en) %>%
  dplyr::select(-ocean)
  
catch_access_df_highseas_fin <- catch_access_df_highseas %>%
  rbind(ukraine_effort_highseas)

qs::qsave(catch_access_df_highseas_fin, here("data/int/prediction_historical_data/catch_fishing_access_high_seas.qs"))

```
