---
title: "Predict props to cells per flag, year, gear, length"
output: html_document
date: "2024-12-11"
---

# Summary

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(terra)
library(foreach)
library(doParallel)
library(progress)
library(pryr)  # for memory tracking
library(glue)
library(qs)
library(here)
library(randomForest)
library(janitor)
library(arrow)
library(countrycode)
library(sf)
library(ranger)
library(data.table)

source(here("R/dir.R"))
```

## Functions to prepare enviro and effort data  

```{r}

elnino <- read.csv(here("data/model_features/enso_index.csv"))

pdo <- read.csv(here("data/model_features/pdo_index.csv"))

world_bank <- read.csv(here("data/model_features/world_bank_regions.csv"))

gfi_df <- read.csv(here("data/model_features/global_fishing_index_governance.csv"))
  
# Load prepared data
model_data <- readRDS(here("data/model_features/prepared_data_1deg.rds")) %>%
  mutate(year = as.numeric(year)) # i think we need to adjust the prepared data to include cells which have no fishing effort in them. We can do this in the loop below? 

## read in environmental variables

  global_grid <- read.csv(here("data/model_features/deg_1_x_1/global_grid.csv"))
  
  ocean_data <- read.csv(here("data/model_features/deg_1_x_1/errdap_chl.csv")) %>%
  left_join(read.csv(here("data/model_features/deg_1_x_1/errdap_sst.csv")), by = c("pixel_id", "year")) %>% 
  left_join(read.csv(here("data/model_features/deg_1_x_1/remss_wind.csv")), by = c("pixel_id", "year")) %>%
  filter(year %in% c(2015:2024))
  
  ocean_data_historical <- qs::qread(here("data/int/prediction_historical_data/one_degree/chl_yearly_1950_2014.qs")) %>%
    dplyr::select(-pixel_area_m2) %>%
  left_join(qs::qread(here("data/int/prediction_historical_data/one_degree/sst_yearly_1950_2014.qs")) %>%  dplyr::select(-pixel_area_m2), by = c("pixel_id", "year")) %>% 
  left_join(qs::qread(here("data/int/prediction_historical_data/one_degree/wind_yearly_1950_2014.qs")) %>% mutate(year = as.numeric(year)) %>% dplyr::select(-pixel_area_m2), by = c("pixel_id", "year"))

  ocean_data_all <- ocean_data %>%
    rbind(ocean_data_historical)

  spatial_data <- global_grid %>%
    left_join(read.csv(here("data/model_features/deg_1_x_1/gfw_static_spatial_measures.csv")) %>% dplyr::select(-lat, -lon), by = "pixel_id") %>%
   left_join(read.csv(here("data/model_features/deg_1_x_1/mesopelagiczones/mesopelagiczones_fixed.csv")), by = "pixel_id") %>%
   left_join(read.csv(here("data/model_features/deg_1_x_1/eez/eez_fix.csv")), by = "pixel_id") %>%
   left_join(read.csv(here("data/model_features/deg_1_x_1/fao/fao_fixed.csv")), by = "pixel_id") %>%
   left_join(read.csv(here("data/model_features/deg_1_x_1/oceans_fixed.csv")), by = "pixel_id") %>%
    left_join(read.csv(here("data/model_features/deg_1_x_1/seamounts.csv")), by = "pixel_id") %>% 
   crossing(., year = c(1950:2024)) 
  
    gfw_reception_data <- read.csv(here("data/model_features/deg_1_x_1/gfw_reception_quality.csv"))
    
    sea_ice_df <- qs::qread(here("data/model_features/deg_1_x_1/sea_ice_features_lit.qs"))
    
  mpa_data <- qs::qread(here("data/model_features/deg_1_x_1/mpa/mpa_data_all.qs"))
  
  
  env_data <- spatial_data %>%
    left_join(sea_ice_df) %>%
    left_join(ocean_data_all) %>% # cool.
    left_join(mpa_data) %>%
    left_join(elnino) %>%
    left_join(pdo) %>% 
    left_join(gfw_reception_data) %>%
    left_join(world_bank, by = c("eez_id")) %>%
    mutate(eez_region_world_bank_7 = ifelse(eez_id %in% c("High seas", "Land"), "High seas", eez_region_world_bank_7)) %>% 
    left_join(gfi_df, by = c("eez_sovereign" = "flag_fin")) %>% # add in global fishing index data here
    mutate(gov_score = ifelse(eez_id >= 99999 & is.na(gov_score), "high_seas", gov_score)) %>%
        mutate(gov_score = ifelse(eez_id < 99999 & is.na(gov_score), "no_data", gov_score)) %>%
    dplyr::select(-eez_sovereign, -nearest_seamount_id) %>% 
    distinct()  %>%
    dplyr::select(-geometry_wkt)

mollweide_projection <- "+proj=moll +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +units=m +no_defs"

data_grid_area <- global_grid %>%
    st_as_sf(wkt = "geometry_wkt",
           crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")%>% 
  st_wrap_dateline(options = c("WRAPDATELINE=YES", "DATELINEOFFSET=180"), quiet = TRUE) %>% 
  st_transform(mollweide_projection) %>%
  mutate(pixel_area_m2 = st_area(geometry_wkt)%>%
           units::drop_units()) 

```

```{r}

  hist_fish_data <- qs::qread(here("data/int/rousseau_gear_fix.qs")) %>%
    ## remove artisanal
    filter(sector == "I") %>%
    group_by(year, flag_fin = country, gear = gear_new, length_category) %>%
    summarize(
      total_nominal_fishing_hours = sum(nom_active_hours, na.rm = TRUE),
      total_effective_fishing_hours = sum(eff_active_hours, na.rm = TRUE),
      .groups = 'drop'
    ) %>%
    dplyr::select(flag_fin, year, gear, length_category, total_nominal_fishing_hours, total_effective_fishing_hours) %>%
        filter(total_nominal_fishing_hours > 0) %>%
  mutate(flag_country_name = countrycode(flag_fin, origin = "iso3c", destination = "country.name"))

  
flags <- unique(hist_fish_data$flag_fin) # get the flags we need to run models for # 151 of them

```


Read in EEZ id and FAO id lookup tables so we can save with the actual iso3c and FAO id numbers

```{r}

eez_lookup <- read.csv(here("data/model_features/deg_1_x_1/eez/eez_lookup_fix.csv")) %>%
  clean_names() %>%
  mutate(eez_country_name = countrycode(eez_sovereign, origin = "iso3c", destination = "country.name")) %>%
    mutate(eez_country_name = case_when(
    eez_sovereign == "RAA" ~ "Azores", 
    eez_sovereign == "RAM" ~ "Madeira",
    TRUE ~ eez_country_name
  )) %>%
  mutate(eez_country_name = ifelse(is.na(eez_country_name), "High seas", eez_country_name))


fao_lookup <- read.csv(here("data/model_features/deg_1_x_1/fao/fao_major_ids.csv")) %>%
  clean_names() %>%
  dplyr::select(fao_id, "fao_major_fishing_area" = "name_en")

global_grid <- read.csv(here("data/model_features/deg_1_x_1/global_grid.csv")) %>%
  dplyr::select(lon, lat, pixel_id)

access_data <- qs::qread(here("data/int/prediction_historical_data/catch_fishing_access.qs"))
access_data_high_seas <- qs::qread(here("data/int/prediction_historical_data/catch_fishing_access_high_seas.qs"))

test <- access_data %>%
  filter(year == 2017, 
         flag_fin == "BEN")

env_grid <- env_data %>% 
  dplyr::select(lon, lat) %>% distinct()

env_data_fishing_areas <- env_data %>% 
  filter(sea_ice_present == 0) %>%
      left_join(eez_lookup)

test <- env_data_fishing_areas %>%
  distinct(pixel_id, eez_sovereign) %>%
  group_by(eez_sovereign) %>%
  summarise(n_cells = n()) # there are some eezs with only 1 pixel 

env_grid_fishing_areas <- env_data_fishing_areas %>% 
  dplyr::select(lon, lat) %>% distinct()

```
Now let's run the stage 1 and 2 predictions ONCE and save the datasets. We can lower the size of the datasets by filtering for stage 1 p > 0, since we will never use the p = 0 cells. 

 - Takes ~5 hours with 38 cores
 - ONLY NEEDS TO BE RUN ONCE! (unless you've changed the source models) 

```{r}
years <- c(1950:2017)
stage_1_path <- file.path(rdsi_dir, "prep/random_forest/stage_1_models/")
stage_1_files <- list.files(stage_1_path, full.names = TRUE)
stage_1_flags <- unique(sub(".*stage_1_rf_model_full_data_([A-Z]{3})_.*", "\\1", stage_1_files))

stage_2_path <- file.path(rdsi_dir, "prep/random_forest/stage_2_models/")
stage_2_files <- list.files(stage_2_path, full.names = TRUE)
stage_2_flags <- unique(sub(".*stage_2_rf_model_full_data_([A-Z]{3})_.*", "\\1", stage_2_files))

diff_flags_1 <- setdiff(stage_1_flags, stage_2_flags)
diff_flags_2 <- setdiff(stage_2_flags, stage_1_flags)

model_flags <- intersect(stage_1_flags, stage_2_flags) # 147

cl <- makeCluster(38) 
registerDoParallel(cl)

 foreach(flag = model_flags, .packages = c("dplyr", "tidyverse", "randomForest", "qs", "glue", "ranger", "terra")) %dopar% {
    #  flag = "VNM"
   
   #### Read in models 
   # read in stage 2 model! 
# Get the list of matching files
matching_files <- list.files(
  path = file.path(rdsi_dir, "prep/random_forest/stage_2_models/"), 
  pattern = glue::glue("stage_2_rf_model_full_data_{flag}_.*\\.qs$"),
  full.names = TRUE
)

# Check if a matching file exists
if (length(matching_files) > 0) {
  stage2_model <- qs::qread(matching_files[1])  # Read the first matching file
} else {
  stop("No matching file found for flag: ", flag)
}

## read in stage 1 model! 
# Get the list of matching files
matching_files <- list.files(
  path = file.path(rdsi_dir, "prep/random_forest/stage_1_models/"), 
  pattern = glue::glue("stage_1_rf_model_full_data_{flag}_.*\\.qs$"),
  full.names = TRUE
)

# Check if a matching file exists
if (length(matching_files) > 0) {
  stage1_model <- qs::qread(matching_files[1])  # Read the first matching file
}  else {
  stop("No matching file found for flag: ", flag)
}

 for(yr in years) {
  
# yr = 1950

# get combination of categories we need to run through our models
model_data_flag <- hist_fish_data %>%
  dplyr::select(flag_fin, gear, length_category, year, total_nominal_fishing_hours ) %>%
  filter(flag_fin == flag, year == yr)

distinct_cats <- model_data_flag %>%
    distinct(flag_fin, gear, length_category) 

full_data <- tidyr::crossing(env_grid_fishing_areas, distinct_cats) %>% 
  crossing(., year = yr) %>%
  left_join(env_data_fishing_areas, by = c("lon", "lat", "year")) %>%
  na.omit()

if(nrow(full_data) == 0){
    cat("skipping", flag, yr, "check to make sure this is right")
  next()
}


# now make predictions on the full dataset 
stage1_preds <- full_data %>%
  mutate(pred_presence = predict(stage1_model, data = .)$predictions[, "1"])

stage1_preds_filt <- stage1_preds %>%
  filter(pred_presence > 0) # we'll never use these cells, so can filter out now. Significantly lowers the number of rows we will need to save. 


oos_preds <- full_data %>%
  mutate(pred_prop = predict(stage2_model, data = .)$predictions) # these are stage 2 predictions

oos_preds_filt <- oos_preds %>%
  filter(pixel_id %in% c(unique(stage1_preds_filt$pixel_id)))


qs::qsave(stage1_preds_filt, file.path(rdsi_dir, glue("prep/random_forest/stage_1_predictions/yearly/model_preds_{flag}_{yr}.qs"))) 

qs::qsave(oos_preds_filt, file.path(rdsi_dir, glue("prep/random_forest/stage_2_predictions/yearly/model_preds_{flag}_{yr}.qs"))) 

  }

}

stopCluster(cl)


```

First, we need to make an EEZ "accessibility" layer based on predictions for each flag from 2015-2017. This is necessary because we trust the predictions from 2015-2017 the most, as these line up with the AIS data years. Since we eventually make predictions for each flag with p>0 (for domestic fishing), we will need to limit the number of cells in which predictions can be made based on the maximum number that is allowed from 2015-2017. Since fishing is likely to contract farther back in time, it is reasonable to assume that the 2015-2017 number of cells where predictions are made are the "maximum" allowable number of cells for fishing in each flag country. 

 - Problem: There are some countries for which we can not make predictions from 2015-2017, but do make predictions for earlier years (Yemen for example).  - Solution: In these cases, we use the positive predictions from maximum year they have in the data to as our accessibility layer. For example, Yemen has predictions for 2013, so we will use that as the maximum number of cells. 

 
```{r}
stage_1_path <- file.path(rdsi_dir, "prep/random_forest/stage_1_predictions/yearly/")
stage_1_files <- list.files(stage_1_path, full.names = TRUE)
stage_1_flags <- unique(sub(".*model_preds_([A-Z]{3})_.*", "\\1", stage_1_files))

stage_2_path <- file.path(rdsi_dir, "prep/random_forest/stage_2_predictions/yearly/")
stage_2_files <- list.files(stage_2_path, full.names = TRUE)
stage_2_flags <- unique(sub(".*model_preds_([A-Z]{3})_.*", "\\1", stage_2_files))

setdiff(stage_1_flags, stage_2_flags) # 0
setdiff(stage_2_flags, stage_1_flags) # 0 

model_flags <- intersect(stage_1_flags, stage_2_flags) # 147; perfect! 

years <- c(2015:2017)

env_data_loop <- env_data %>%
  left_join(eez_lookup)


cl <- makeCluster(38) 
registerDoParallel(cl)


 foreach(flag = model_flags, .packages = c("dplyr", "tidyverse", "randomForest", "qs", "glue", "ranger")) %dopar% {

 for(yr in years) {
# flag = "VNM"
# yr = 2017

# get combination of categories we need to run through our models
model_data_flag <- hist_fish_data %>%
  dplyr::select(flag_fin, gear, length_category, year, total_nominal_fishing_hours) %>%
  filter(flag_fin == flag, year == yr)

if(nrow(model_data_flag) == 0){
  
  model_data_flag <- hist_fish_data %>%
      dplyr::select(flag_fin, gear, length_category, year, total_nominal_fishing_hours) %>%
      filter(flag_fin == flag)
  
   max_year <- max(model_data_flag$year)
   
     
  model_data_flag <- hist_fish_data %>%
      dplyr::select(flag_fin, gear, length_category, year, total_nominal_fishing_hours) %>%
      filter(flag_fin == flag, year == max_year)
  
  
stage1_preds <- qread(file.path(rdsi_dir, glue("prep/random_forest/stage_1_predictions/yearly/model_preds_{flag}_{max_year}.qs")))

  
}else{
  stage1_preds <- qread(file.path(rdsi_dir, glue("prep/random_forest/stage_1_predictions/yearly/model_preds_{flag}_{yr}.qs")))

}



# distinct EEZ cells only
eez_cells <- stage1_preds %>%
  filter(flag_fin == eez_sovereign) %>%
  distinct(pixel_id, lon, lat, year, flag_fin, eez_sovereign, pred_presence)

# test <- eez_cells %>%
#   distinct(pixel_id)

eez_cells_distinct <- env_data_loop %>%
  filter(eez_sovereign == flag) %>%
  distinct(pixel_id, lon, lat)

# accessible cells in own EEZ
N_accessible <- nrow(eez_cells_distinct)
m <- 0.3

k_min <- ceiling(m * N_accessible) # these are the minimum amount of cells we will allow

# 99th percentile threshold inside own EEZ
# p95_eez <- stage1_preds %>%
#     filter(flag_fin == eez_sovereign) %>%
#    distinct(pixel_id, lon, lat, pred_presence) %>%
#   summarise(p95 = quantile(pred_presence, probs = 0.95, na.rm = TRUE)) %>%
#   pull(p95)

# eez_limit <- min(p95_eez, 0.5) # this is the threshold for our predictions

# cells retained by threshold rule
eez_cells_thresh <- stage1_preds %>%
    filter(flag_fin == eez_sovereign) %>%
 # filter(pred_presence >= eez_limit)
  filter(pred_presence > 0)

# test <- eez_cells_thresh %>%
#   dplyr::select(lon, lat, pixel_id) %>%
#   rast(., type = "xyz")

n_thresh <- nrow(eez_cells_thresh %>% distinct(lon, lat))

# decide whether to use threshold rule or fallback top-K rule
if (n_thresh < k_min) { # if we have less rows from our cutoff with 99th quantile than our minimum, we will select the top k_min cells to keep. This will really only happen for countries with really small EEZs that do not predict well, or countries that just do not predict well/have a lot of AIS
  
  K_ref <- ceiling(0.3 * N_accessible)
  
  selected_eez_cells <- eez_cells %>%
    group_by(pixel_id, lon, lat, year, flag_fin, eez_sovereign) %>%
    summarise(pred_presence = max(pred_presence)) %>%
    ungroup() %>%
    arrange(desc(pred_presence)) %>%
    slice_head(n = K_ref) %>%
    distinct(pixel_id, lon, lat, year, flag_fin, eez_sovereign) %>%
    mutate(keep = 1) %>%
    mutate(limit_type = "30% of EEZ")

# test <- selected_eez_cells %>%
#   dplyr::select(lon, lat, pixel_id) %>%
#   rast(., type = "xyz")
  
  cat(flag, yr, ": using fallback top-K rule with", K_ref, "cells\n")
  
} else {
  
  selected_eez_cells <- eez_cells_thresh %>%
    distinct(pixel_id, lon, lat, year, flag_fin, eez_sovereign) %>%
        mutate(keep = 1) %>%
    mutate(limit_type = "Threshold of p > 0")
  
  cat(flag, yr, ": using threshold rule with", n_thresh, "cells\n")
}


qs::qsave(selected_eez_cells, file.path(rdsi_dir, glue("prep/random_forest/eez_2015_2017_reference_cells_p95/yearly/model_preds_{flag}_{yr}.qs")))

  }

}

stopCluster(cl)

```
 

Now let's grab the year from 2015-2017 which produced the largest number of allowable cells

```{r}
# now loop over each country and bind each data set into one that we will read those in for a mask in the next prediction loop

for(flag in model_flags){
  # flag = "VNM"
  files <- list.files( file.path(rdsi_dir, "prep/random_forest/eez_2015_2017_reference_cells_0/yearly/"), pattern = flag, full.names = TRUE)
  
  if (length(files) == 0) {
  cat("skipping check to make sure this is right")
  next()
  }
  
df <- lapply(files, qread) %>%
  bind_rows() %>%
  dplyr::select(pixel_id, lon, lat, year, flag_fin, eez_sovereign, keep, limit_type)

df_choose <- df %>% 
  group_by(year) %>%
  summarise(n_cells = n(), .groups = "drop")

year_keep <- df_choose %>%
  slice_max(order_by = n_cells, n = 1, with_ties = TRUE) %>%
  slice_max(order_by = year, n = 1) %>%
  pull(year)

df_filtered <- df %>%
  filter(year == year_keep) %>%
  dplyr::distinct(pixel_id, lon, lat, year, flag_fin, eez_sovereign, keep, limit_type)
 

  # test <- df_filtered %>%
  #   dplyr::select(lon, lat, keep) %>%
  #   rast(., type = "xyz")

  
qs::qsave(df_filtered, file.path(rdsi_dir, glue("prep/random_forest/eez_2015_2017_reference_cells_0/access_mask_2015_2017_{flag}.qs"))) 

}

# test a couple to make sure they look OK
test <- qs::qread(file.path(rdsi_dir, ("prep/random_forest/eez_2015_2017_reference_cells_0/access_mask_2015_2017_VNM.qs"))) 
test <- qs::qread(file.path(rdsi_dir, ("prep/random_forest/eez_2015_2017_reference_cells_0/access_mask_2015_2017_IND.qs"))) 
test <- qs::qread(file.path(rdsi_dir, ("prep/random_forest/eez_2015_2017_reference_cells_0/access_mask_2015_2017_IDN.qs")))
test <- qs::qread(file.path(rdsi_dir, ("prep/random_forest/eez_2015_2017_reference_cells_0/access_mask_2015_2017_THA.qs"))) 
test <- qs::qread(file.path(rdsi_dir, ("prep/random_forest/eez_2015_2017_reference_cells_0/access_mask_2015_2017_BGD.qs"))) 
# OK so THA and BGD might be too small??

# test <- qs::qread(file.path(rdsi_dir, ("prep/random_forest/eez_2015_2017_reference_cells_0/access_mask_2015_2017_YEM.qs")))  
# 
# test2 <- test %>%
#   distinct(lon, lat, keep) %>%
#   rast(., type = "xyz")

```


After fitting the models, we generated predictions for the full spatiotemporal prediction set. As a result, stage-1 predictions are in-sample, because the occurrence model was fit using all available data, whereas stage-2 predictions are effectively out-of-sample with respect to zero-effort cells, because the positive-effort model was fit only on cells with observed fishing effort. 

First we calculated maximum domestic spatial footprint number of cells we will allow for allocation (above code): 

N_ref was defined as the largest annual number of domestic cells retained in any year from 2015–2017, rather than the union of cells across all three years. This ensured that the reference footprint represented a plausible maximum annual domestic extent, while avoiding inflation of the footprint through aggregation of interannual shifts in fishing location across the reference period. This reference period was chosen because we assumed domestic fishing extent would be most fully represented in the most recent years, and because predictions from 2015 onward were considered the most reliable. To define the reference footprint, we retained cells with stage-1 predicted presence values of p > 0. 

In some countries, however, this raw reference footprint was implausibly small relative to the size of the domestic EEZ, likely reflecting sparse AIS coverage or limited training data rather than a truly narrow fishing footprint. To avoid propagating these likely observation-driven artifacts into the hindcast, we imposed a minimum reference footprint equal to 30% of domestic EEZ cells whenever the calculated N_ref fell below that threshold. This floor was intended as a conservative safeguard against unrealistically small domestic footprints in data-poor cases, particularly in some small or weakly observed countries in South and Southeast Asia. For example, when no floor was applied, the procedure yielded an N_ref of only 3 cells for Vietnam, representing about 4% of cells in its EEZ, which we considered implausibly low.

Ultimately, we chose N_ref for each flag from the year which had the most amount of cells present, and if all years have the same amount of cells, we chose the latest year (I think this stipulation is unneccessary since we aren't actually limiting to the extent, just taking the number of allowable cells). 

Then we make annual predictions for 1950-2017: 
For annual predictions from 1950–2017, only cells that had a non-zero stage-1 predicted presence probability (p > 0). For cells outside the domestic EEZ, we retained a stricter criterion and only allowed allocation where the stage-1 predicted presence probability was at least 0.5 (p ≥ 0.5).

Within the domestic mask, eligible cells were ranked from highest to lowest stage-1 predicted presence probability. We then calculated the total number of domestic cells allowed to receive allocation in a given year as:

N_(year,total) = min(N_ref, N_ref*((E_year/E_ref))^alpha)

$N_{year,total} = \min\left(N_{ref},\; N_{ref}\left(\frac{E_{year}}{E_{ref}}\right)^{\alpha}\right)$


where N_ref is the number of cells in the maximum domestic spatial footprint mask, E_year is the total effort for the flag state in the that year, E_ref is the reference effort for that flag state (defined as the maximum annual effort observed during 2015–2017), and α = 0.5, such that spatial extent expands sublinearly with total effort. 

To prevent implausibly small domestic fishing footprints in early years, we imposed a minimum extent equal to (i) the greater of 10% of domestic EEZ cells or (ii) the number of cells with depths less than or equal to 200 meters within the domestic EEZ. We used the 200 meter threshold as a heuristic proxy for shelf-associated fishing grounds that were more likely to remain accessible to domestic fleets in earlier years of the hindcast, when fishing distributions were likely more spatially constrained (Chuenpagdee et al., 2006; Watson & Tidd, 2018; Pauly et al., 2020). The 10% threshold serves as a fallback for EEZs with very narrow shelves, such as oceanic islands, where a strict shelf-based constraint would likely underestimate the spatial extent of historically adapted domestic fisheries. This allows historical domestic fishing footprints to contract relative to the modern reference period while maintaining plausible minimum extents. Although this is not a standard approach in fisheries modelling, where constraints are more often based on technological or operational factors such as fuel capacity or travel distance, it is conceptually analogous to bounded allocation approaches used in land-use allocation models (Rahman & Szabó, 2021).

Because predictions were generated separately by gear and vessel length class, the annual domestic footprint constraint was applied independently within each gear. For a given flag-year, we first calculated the total allowable annual footprint size, N_(year,total). This value was then used as the maximum number of domestic cells that could receive allocation within each gear. Within gear, the allowable number of cells was apportioned across vessel length classes according to each length class’s share of total effort in that gear and year, rounding up to the nearest whole number:

$N_{year,g,l} = \left\lceil N_{year,total}\left(\frac{E_{year,g,l}}{E_{year,g}}\right)\right\rceil$

where E_(year,g,l) is the effort for the gear g and length class l in that flag-year, and E_(year,g) is the total effort across all length calsses for gear g inthat flag-year. Within each gear-length group, eligible domestic cells were ranked by stage-1 predicted presence probability, and the highest-ranked cells were retained up to the group-specific limit N_(year,g,l), Effort was then allocated across the retained cells.We applied the footprint limit independently by gear because gears were modeled separately and may occupy overlapping but distinct spatial footprints within the same flag-year; using a single shared pool of cells across all gears would have imposed artificial competition among gear groups for space.


First create master datasets with N_ref, N_year,total, so we don't have to calculate it in the forloop each time for each country. 

```{r}
## Now lets create a master dataset that has N_ref, and N_year,total in it, so that we don't have to calculate it in the forloop below. 

# grab N_ref
path <- file.path(rdsi_dir, "prep/random_forest/eez_2015_2017_reference_cells_0")
files <- list.files(path, full.names = TRUE)[-148] # remove the "yearly" folder

all_n_ref_data <- lapply(files, qread) %>%
  bind_rows()

n_ref_group <- all_n_ref_data %>%
  group_by(flag_fin) %>%
  summarise(n_ref = n()) %>%
  ungroup() # should be 147

# now calculate N_year,total
## for this we need to grab the effort for the corresponding years

## these were the four countries that we had to use different predictions for. Lets grab their reference effort:
# "BHR", "YEM", "BLZ", "QAT"
years <- 2015:2017

flags_to_run <- c("BHR", "YEM", "BLZ", "QAT")

eff_ref_missing <- data.frame(flag_fin = NA, year = NA, eff_ref = NA)

for(flag in flags_to_run){

 for(yr in years) {
# flag = "YEM"
# yr = 2017

# get combination of categories we need to run through our models
model_data_flag <- hist_fish_data %>%
    filter(flag_fin == flag, year == yr) %>%
  dplyr::select(flag_fin, gear, length_category, year, total_nominal_fishing_hours)

if(nrow(model_data_flag) == 0){
  
  model_data_flag <- hist_fish_data %>%
      dplyr::select(flag_fin, gear, length_category, year, total_nominal_fishing_hours) %>%
      filter(flag_fin == flag)
  
   max_year <- max(model_data_flag$year)
   
     
  model_data_flag <- hist_fish_data %>%
      dplyr::select(flag_fin, gear, length_category, year, total_nominal_fishing_hours) %>%
      filter(flag_fin == flag, year == max_year) %>%
    group_by(flag_fin, year) %>%
    summarise(eff_ref = sum(total_nominal_fishing_hours, na.rm = TRUE)) %>%
    ungroup()
  
}

eff_ref_missing <- rbind(eff_ref_missing, model_data_flag)

 }
}

eff_ref_missing <- eff_ref_missing %>%
  filter(!is.na(flag_fin)) %>%
  distinct()


max_eff_ref <- hist_fish_data %>%
  filter(year %in% c(2015:2017)) %>%
      group_by(flag_fin, year) %>%
    summarise(nom_sum = sum(total_nominal_fishing_hours, na.rm = TRUE)) %>%
    ungroup() %>%
  group_by(flag_fin) %>%
  summarise(eff_ref = max(nom_sum, na.rm = TRUE)) %>%
  ungroup() %>%
    filter(!(flag_fin %in% c("YEM", "BLZ", "BHR", "QAT"))) %>%
  rbind(eff_ref_missing %>%
          dplyr::select(-year))

# now get total effort for every year
eff_ref_df <- hist_fish_data %>%
  dplyr::group_by(flag_fin, year) %>%
  summarise(eff_year = sum(total_nominal_fishing_hours, na.rm = TRUE)) %>%
  ungroup() 

eez_size_df <- env_data_loop %>%
  filter(eez_sovereign != "High seas") %>%
  distinct(pixel_id, lon, lat, eez_sovereign) %>%
  group_by(flag_fin = eez_sovereign) %>%
    summarise(n_cells_eez = n()) %>%
    ungroup()



# calculate floor based on limit of 20m of depth per EEZ

depth_floor_df <- env_data_fishing_areas %>%
  distinct(pixel_id, lon, lat, flag_fin = eez_sovereign, elevation_m) %>%
  filter(elevation_m >= -200,
         flag_fin != "High seas") %>%
  group_by(flag_fin) %>%
  summarise(n_cells_depth = n()) %>%
  ungroup() %>%
  right_join(eez_size_df) %>%
  mutate(n_cells_eez_10 = ceiling(0.1*n_cells_eez)) %>%
  mutate(n_cells_depth = ifelse(is.na(n_cells_depth), 0, n_cells_depth)) %>%
  mutate(n_cells_depth_constraint = pmax(n_cells_eez_10, n_cells_depth)) %>%
  dplyr::select(flag_fin, n_cells_depth_constraint)


# now calculate n_year_total

n_year_total_df <- eff_ref_df %>%
  left_join(max_eff_ref) %>%
  left_join(n_ref_group) %>%
  left_join(eez_size_df) %>%
  left_join(depth_floor_df) %>%
  filter(!is.na(n_ref)) %>% # these are the ones we regional model
  mutate(n_year_total_1 = ceiling(n_ref *((eff_year/eff_ref)^0.5))) %>%
  mutate(n_year_total_fin = pmin(n_ref, n_year_total_1)) %>% # take the rowwise minimum of n_ref and n_year_total to get the final value. We don't want n_year_total to exceed the maximum reference number of cells
  mutate(n_year_total_floor = pmax(ceiling(0.1*n_ref), n_year_total_fin),
         n_year_total_floor_eez = pmax(ceiling(0.1*n_cells_eez), n_year_total_fin),
         n_year_total_depth = pmax(ceiling(n_cells_depth_constraint), n_year_total_fin)) # apply a floor of 10% of n_ref or the number of cells with dpeth >200m. This way we won't allow n_ref to get all the way down to 1. 


```


Now we can rank the p>0 predictions and limit them n_year_total_fin by selecting the top n_year_total probability values. 

After that, we can join to stage 2 predictions and allocate national effort to the predicted cells. 

```{r}
stage_1_path <- file.path(rdsi_dir, "prep/random_forest/stage_1_predictions/yearly/")
stage_1_files <- list.files(stage_1_path, full.names = TRUE)
stage_1_flags <- unique(sub(".*model_preds_([A-Z]{3})_.*", "\\1", stage_1_files))

stage_2_path <- file.path(rdsi_dir, "prep/random_forest/stage_2_predictions/yearly/")
stage_2_files <- list.files(stage_2_path, full.names = TRUE)
stage_2_flags <- unique(sub(".*model_preds_([A-Z]{3})_.*", "\\1", stage_2_files))

setdiff(stage_1_flags, stage_2_flags) # 0
setdiff(stage_2_flags, stage_1_flags) # 0 

model_flags <- intersect(stage_1_flags, stage_2_flags) # 147; perfect! 

years <- c(1950:2017)

out_dir <- file.path(rdsi_dir, "prep/random_forest/predictions_industrial_flag_revised_floor_depth/yearly")

tasks <- CJ(flag = model_flags, year = years)
tasks[, out_file := file.path(out_dir, sprintf("model_preds_%s_%d.qs", flag, year))]
tasks <- tasks[!file.exists(out_file)]

data_to_run <- hist_fish_data %>%
  distinct(flag_fin, year) %>%
  left_join(tasks, by = c("year", "flag_fin" = "flag")) %>%
  filter(!is.na(out_file))

flags_to_run <- unique(data_to_run$flag_fin) 
years_to_run <- unique(data_to_run$year) 


cl <- makeCluster(38) 
registerDoParallel(cl)

 foreach(flag = model_flags, .packages = c("dplyr", "tidyverse", "randomForest", "qs", "glue", "ranger", "terra")) %dopar% {

 for(yr in years) {
# flag = "GAB"
# yr = 2016

# grab country level fishing effort data for flag and year
model_data_flag <- hist_fish_data %>%
  dplyr::select(flag_fin, gear, length_category, year, total_nominal_fishing_hours) %>%
  filter(flag_fin == flag, year == yr)

if(nrow(model_data_flag) == 0){
    cat("skipping", flag, yr, "check to make sure this is right")
  next()
}


# grab the number of cells we'll allow for allocation
n_year_total <- n_year_total_df %>%
  filter(flag_fin == flag,
         year == yr) %>%
  pull(n_year_total_depth)
 
# now grab stage 1 predictions
stage1_preds <- qread(file.path(rdsi_dir, glue("prep/random_forest/stage_1_predictions/yearly/model_preds_{flag}_{yr}.qs")))


# Run simulation where we only allow the top n_year_total cells to be allocated in for domestic fishing

stage1_preds_no_limit <- stage1_preds %>%
    mutate(pred_presence_rescale = case_when(
        flag_fin == eez_sovereign & pred_presence > 0 ~ 1,
        pred_presence >= 0.5 ~ 1,
        TRUE ~ 0
    ))

## now we need to figure out how many cells we will allocate to each gear and length class
gear_length_breakdown <- model_data_flag %>%
  group_by(gear) %>%
  mutate(total_effort_gear = sum(total_nominal_fishing_hours, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(n_cells_allowed = ceiling(n_year_total*(total_nominal_fishing_hours/total_effort_gear))) %>%
  dplyr::select(flag_fin, year, gear, length_category, n_cells_allowed)

stage1_preds_no_limit_choose_dom <- stage1_preds_no_limit %>%
  filter(flag_fin == eez_sovereign) %>%
  left_join(gear_length_breakdown) %>%
  group_by(flag_fin, year, gear, length_category) %>%
  arrange(desc(pred_presence), .by_group = TRUE) %>%
  mutate(
    pred_rank = row_number(),
    pred_presence_rescale = if_else(pred_rank <= dplyr::first(n_cells_allowed), 1, 0)
  ) %>%
  ungroup() 

# test <- stage1_preds_no_limit_choose_dom %>%
#   filter(pred_presence_rescale == 1) %>%
#   dplyr::select(lon, lat, pred_presence_rescale) %>%
#   rast(., type = "xyz")

stage_1_preds_no_limit_final <- stage1_preds_no_limit %>%
  filter(flag_fin != eez_sovereign) %>%
  rbind(., stage1_preds_no_limit_choose_dom %>% dplyr::select(colnames(stage1_preds_no_limit)))

stage2_preds <- qread(file.path(rdsi_dir, glue("prep/random_forest/stage_2_predictions/yearly/model_preds_{flag}_{yr}.qs")))

all_preds_no_limit <- stage2_preds %>%
  dplyr::select(lat, lon, gear, length_category, year, flag_fin, pred_prop) %>% # probably select lat, lon, gear, length, year, flag here?
  left_join(., stage_1_preds_no_limit_final) %>% 
    dplyr::select(lon, lat, flag_fin,gear, length_category, year, fao_id, eez_id, eez_sovereign, eez_country_name, pred_prop, pred_presence_rescale) %>% 
  mutate(pred_prop_final = pred_prop*pred_presence_rescale) %>%
  filter(pred_prop_final > 0) %>%
  filter(!is.na(pred_prop_final)) %>%
 group_by(year, flag_fin, gear, length_category) %>%
  mutate(prop_fishing_hours_cell_predict_rescaled = pred_prop_final / sum(pred_prop_final, na.rm = TRUE)) %>%
 ungroup() %>% # need to rescale the predictions to be between 0 and 1 here so that we can allocate effort
  left_join(hist_fish_data) %>%
  mutate(nom_active_fishing_hours = prop_fishing_hours_cell_predict_rescaled*total_nominal_fishing_hours,
         eff_active_fishing_hours = prop_fishing_hours_cell_predict_rescaled*total_effective_fishing_hours) %>%
  left_join(global_grid) %>%
  left_join(fao_lookup) %>%
  left_join(eez_lookup) %>%
    dplyr::select(pixel_id, lon, lat, year, flag_fin, gear, length_category, eez_sovereign, fao_id, fao_major_fishing_area, nom_active_fishing_hours, eff_active_fishing_hours) %>%
  filter(nom_active_fishing_hours > 0) %>%
  mutate(sector = "Industrial")

# test_preds_no_limit <- all_preds_no_limit %>%
#   # filter(gear == "Trawl_Bottom") %>%
#   dplyr::group_by(lon, lat) %>%
#   summarise(nom_active_fishing_hours = sum(nom_active_fishing_hours, na.rm = TRUE)) %>%
#   ungroup() %>%
#   rast(., type = "xyz")
# plot(test_preds_no_limit)



if(nrow(all_preds_no_limit) == 0){
    cat("skipping", flag, yr, "no predictions made...")
  next()
}

qs::qsave(all_preds_no_limit, file.path(rdsi_dir, glue("prep/random_forest/predictions_industrial_flag_revised_floor_depth/yearly/model_preds_{flag}_{yr}.qs"))) # qs is smaller

  }

}

stopCluster(cl)


```

Combine all years for flag and save 

 - only takes a minute
  - need to convert hours to days 
  - need to calculate nv
    -  `nv = nominal_effort / (days_at_sea*mean_engine_p)`
 
```{r}
library(strex)
flags <- unique(str_after_nth(str_before_last(list.files(file.path("/home/ubuntu/data_storage/prep/random_forest/predictions_industrial_flag_revised_floor_0/yearly")),"_"), "_", 2)) # 147 flags, cool.

days_to_hours_conversion <- qs::qread(here("data/int/hours_to_days_conversion.qs")) %>%
  rename(lon = x, lat = y) %>%
  left_join(global_grid)

nv_conversion <- qs::qread(here("data/model_features/rousseau_ind_effort_nv_conversion.qs"))

data_grid_area <- data_grid_area %>%
  st_drop_geometry()

cl <- makeCluster(30)  # could probably increase this? 
registerDoParallel(cl)

foreach(flag = flags, .packages = c("qs", "dplyr", "tidyverse", "glue", "countrycode")) %dopar% {
  # flag = "HKG"
  
  all_files_flag_no_limit <- list.files(file.path(rdsi_dir, "prep/random_forest/predictions_industrial_flag_revised_floor_0/yearly/"), pattern = flag, full.names = TRUE)
  
  
  # hours / hours/day = hours * day/hours = days

  all_data_flag_no_limit <- lapply(all_files_flag_no_limit, qread) %>%
    bind_rows() %>%
    left_join(eez_lookup) %>%
    mutate(flag_country_name = countrycode(flag_fin, origin = "iso3c", destination = "country.name")) %>%
    dplyr::select(pixel_id, lon, lat, year, flag_fin, flag_country_name, gear, length_category, eez_sovereign, eez_country_name, fao_id, fao_major_fishing_area, nom_active_fishing_hours, eff_active_fishing_hours, sector) %>%
    left_join(nv_conversion) %>%
          mutate(nv = nom_active_fishing_hours/(days_at_sea*mean_engine_p)) %>%
    left_join(., days_to_hours_conversion) %>%
    mutate(nom_active_fishing_days = nom_active_fishing_hours/mean,
           eff_active_fishing_days = eff_active_fishing_hours/mean) %>%
    dplyr::select(-mean, -days_at_sea, -mean_engine_p, -nom_active_days, -eff_active_days, -p, -gt) %>%
        rename(flag_country_iso3c = flag_fin, eez_sovereign_iso3c = eez_sovereign, eez_sovereign_name = eez_country_name, fao_fishing_id = fao_id) %>%
    left_join(data_grid_area) %>%
    mutate(pixel_area_km2 = pixel_area_m2/1000000) %>%
    mutate(eff_days_km2 = eff_active_fishing_days/pixel_area_km2,
           nom_days_km2 = nom_active_fishing_days/pixel_area_km2,
           eff_hours_km2 = eff_active_fishing_hours/pixel_area_km2,
           nom_hours_km2 = nom_active_fishing_hours/pixel_area_km2,
           nv_km2 = nv/pixel_area_km2) # calculate hours per km2 here too
  
  qs::qsave(all_data_flag_no_limit, file.path(rdsi_dir, glue("prep/random_forest/predictions_industrial_flag_revised_floor_0/model_preds_1950_2017_{flag}.qs")))
  
}

stopCluster(cl)


test_preds <- qs::qread("/home/ubuntu/data_storage/prep/random_forest/predictions_industrial_flag_revised_floor_depth/model_preds_1950_2017_IDN.qs") %>%
  filter(year == 1950) %>%
  filter(flag_country_iso3c == eez_sovereign_iso3c) %>%
  dplyr::group_by(lon, lat) %>%
  summarise(nom_active_fishing_hours = sum(nom_active_fishing_hours, na.rm = TRUE)) %>%
  ungroup() %>% 
  left_join(data_grid_area) %>%
  mutate(hours_km2 = nom_active_fishing_hours/(pixel_area_m2/1000000)) %>%
  dplyr::select(lon, lat, hours_km2) %>%
  rast(., type = "xyz")

plot(test_preds)

## earlier predictions might be TOO constricted now... But i think still better than before...

```

```{r}

codes <- fread("https://data.imas.utas.edu.au/attachments/1241a51d-c8c2-4432-aa68-3d2bae142794/SAUPtoCountry.csv")

ind_df <- fread("https://data.imas.utas.edu.au/attachments/1241a51d-c8c2-4432-aa68-3d2bae142794/effort_mapped/mapped_1950_I.csv")

ind_df_filt <- ind_df %>%
  filter(SAUP == 356)
  
test <- ind_df_filt %>%
  group_by(Lon, Lat) %>%
  summarise(NomActive = sum(NomActive, na.rm = TRUE)) %>%
  ungroup()

test2 <- test %>%
  rast(., type = "xyz") # yeah, Rousseau IND is pretty spread across the coastline for 1950. 
  
```

Make predictions for a specific example country (IDN perhaps?) to show why we need Nref for the supplemental 

```{r}

```



Other option for allocation:

Worst case/easiest: 
 - Create maximum accessibility/extent layer with 2015-2017 union
 - Run using p > min(0.5, 99th quantile)
 - limit those predictions to the accessibility layer
    - Problem is that most hindcast years will probably look similar to the maximum extent layer if allocating in many cells
 



