---
title: "Save data for Zenodo publishing"
output: html_document
date: "2026-04-01"
---

# Summary

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(terra)
library(foreach)
library(doParallel)
library(glue)
library(qs)
library(here)
library(countrycode)
library(strex)
library(data.table)
library(sf)

source(here("R/dir.R"))


  global_grid <- read.csv(here("data/model_features/deg_1_x_1/global_grid.csv"))
mollweide_projection <- "+proj=moll +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +units=m +no_defs"

data_grid_area <- global_grid %>%
    st_as_sf(wkt = "geometry_wkt",
           crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")%>% 
  st_wrap_dateline(options = c("WRAPDATELINE=YES", "DATELINEOFFSET=180"), quiet = TRUE) %>% 
  st_transform(mollweide_projection) %>%
  mutate(pixel_area_m2 = st_area(geometry_wkt)%>%
           units::drop_units()) 

```


Save csv files to upload to Zenodo and the app 

 - We need to grab any gapfilled flags that are not in the flag level modelling from the regional level modelling 
 
 We gap fill using regional models where 1) a flag country is not included at all in the AIS data (e.g., DOM, MYT, RAA, RAM), where a flag country has less than 100 observations across all years of data (e.g., GLP, COG, COG, JAM), or 3) where the flag-specific model was unable to predict any presence for a specific flag/gear/length combination (e.g., GTM bottom trawls 2015-2017). 


```{r}
# predictions_industrial_flag_revised_floor_0

  modelled_flags_original <- str_before_first(str_after_last(list.files(file.path(rdsi_dir, "prep/random_forest/predictions_industrial_flag_revised_floor_0/"), pattern = ".qs", full.names = TRUE), "_"), "\\.qs")

  modelled_flags_regional <- str_before_first(str_after_last(list.files(file.path(rdsi_dir, "prep/random_forest/predictions_regional/"), pattern = ".qs", full.names = TRUE), "_"), "\\.qs")
  
flags_to_gapfill <- setdiff(modelled_flags_regional, modelled_flags_original)


all_flags <- c(modelled_flags_original, flags_to_gapfill)

low_sample_flags <- readRDS(here("data/model_features/prepared_data_1deg.rds")) %>%
  group_by(flag_fin) %>%
  summarise(n_obs = n()) %>%
  ungroup() %>%
  filter(flag_fin %in% all_flags) %>%
  filter(n_obs < 100) %>%
  pull(flag_fin) %>%
  unique()
                

cl <- makeCluster(24)  # could probably increase this? 
registerDoParallel(cl)

foreach(flag = all_flags, .packages = c("qs", "dplyr", "tidyverse", "glue", "countrycode")) %dopar% {
  # flag = "IRQ"
  
  
  all_files_flag_original <- list.files(file.path(rdsi_dir, "prep/random_forest/predictions_industrial_flag_revised_floor_0/"), pattern = flag, full.names = TRUE)
  all_files_flag_regional_gf <- list.files(file.path(rdsi_dir, "prep/random_forest/predictions_regional/"), pattern = flag, full.names = TRUE)

  
  #   # Skip iteration if no files are found
  # if (length(all_files_flag) == 0) next
  

  all_data_flag_original <- lapply(all_files_flag_original, qread) %>%
    bind_rows() %>%
    mutate(data_type = "flag model")

    if(flag == "GAB"){
    
    all_data_flag_original <- all_data_flag_original %>%
      filter(year != 2017)
  } # we're going to use regional predictions for GAB 2017, because country predictions are horrible.
  
  
  all_data_flag_regional_gf <- lapply(all_files_flag_regional_gf, qread) %>%
    bind_rows()  %>%
    mutate(data_type = "regional model")
  
    if(nrow(all_data_flag_original) == 0|flag %in% c(low_sample_flags)){
    
    all_data_flag_original <- all_data_flag_regional_gf %>%
      mutate(nom_active_fishing_hours = 0, eff_active_fishing_hours = 0, nom_active_fishing_days = 0, eff_active_fishing_days = 0, eff_days_km2 = 0, nom_days_km2 = 0, eff_hours_km2 = 0, nom_hours_km2 = 0, nv = 0, nv_km2 = 0, data_type = NA)
    
    }
  
if(nrow(all_data_flag_regional_gf) == 0){
    
    all_data_flag_regional_gf <- all_data_flag_original %>%
      mutate(nom_active_fishing_hours = 0, eff_active_fishing_hours = 0, nom_active_fishing_days = 0, eff_active_fishing_days = 0, eff_days_km2 = 0, nom_days_km2 = 0, eff_hours_km2 = 0, nom_hours_km2 = 0, nv = 0, nv_km2 = 0, data_type = NA)
    
  }
  
  original_cats <- all_data_flag_original %>%
    filter(nom_active_fishing_hours > 0) %>%
    distinct(year, flag_country_iso3c, gear, length_category) 
  
  regional_cats <- all_data_flag_regional_gf %>%
    filter(nom_active_fishing_hours > 0) %>%
    distinct(year, flag_country_iso3c, gear, length_category) 
  
  missing_cats <- regional_cats %>%
   anti_join(original_cats, by = c("year", "flag_country_iso3c", "gear", "length_category"))

  data_to_add <- all_data_flag_regional_gf %>%
    semi_join(missing_cats, by = c("flag_country_iso3c", "year", "gear", "length_category"))
  

  all_data_fin <- all_data_flag_original %>%
    filter(nom_active_fishing_hours > 0) %>%
    rbind(data_to_add) %>%
    mutate(flag_country_name = case_when(
      flag_country_iso3c == "RAA" ~ "Azores", 
      flag_country_iso3c == "RAM" ~ "Madeira", 
      flag_country_iso3c == "COG" ~ "Republic of the Congo",
      flag_country_iso3c == "COD" ~ "Democratic Republic of the Congo",
      TRUE ~ flag_country_name
    )) %>%
    dplyr::select(-pixel_area_m2, -pixel_id, -eff_days_km2, -eff_hours_km2, -nom_days_km2, -nom_hours_km2, -nv_km2)
  

  
  write.csv(all_data_fin, file.path(rdsi_dir, glue("prep/random_forest/zenodo_data/mapped_by_flag_country_industrial/model_preds_1950_2017_{flag}.csv")), row.names = FALSE)
}

stopCluster(cl)


## save as one large csv

  all_files_flag <- list.files(file.path(rdsi_dir, "prep/random_forest/zenodo_data/mapped_by_flag_country_industrial/"), full.names = TRUE)

    all_data <- lapply(all_files_flag, fread) %>%
      bind_rows()
    
fwrite(all_data, file.path(rdsi_dir, "prep/random_forest/zenodo_data/mapped_industrial_effort_predictions_1950_2017.csv"), row.names = FALSE)

##### Save the large map files as zip files too - use terminal for this:

# cd /home/ubuntu/data_storage/prep/random_forest/zenodo_data
## first do all the maps together
# zip mapped_industrial_effort_predictions_1950_2017.zip mapped_industrial_effort_predictions_1950_2017.csv

## next do the folder of country maps
# zip -r mapped_by_flag_country_industrial.zip mapped_by_flag_country_industrial


# now save grouped by 
all_data_grouped <- all_data %>% 
  group_by(year, flag_country_iso3c, flag_country_name, gear, length_category, eez_sovereign_iso3c, eez_sovereign_name, fao_fishing_id, fao_major_fishing_area, sector, model_type = data_type) %>%
  summarise(eff_active_fishing_hours = sum(eff_active_fishing_hours, na.rm = TRUE), 
            eff_active_fishing_days = sum(eff_active_fishing_days, na.rm = TRUE), 
            nom_active_fishing_hours = sum(nom_active_fishing_hours, na.rm = TRUE), 
            nom_active_fishing_days = sum(nom_active_fishing_days, na.rm = TRUE),
            nv = sum(nv, na.rm = TRUE)) %>%
  ungroup()


fwrite(all_data_grouped, file.path(rdsi_dir, "prep/random_forest/zenodo_data/industrial_effort_predictions_by_flag_eez_fao_gear_length_1950_2017.csv"), row.names = FALSE)

```


Check to make sure we are not missing any effort 

```{r}
## check to make sure we are not missing any effort
  all_data_grouped <- fread(file.path(rdsi_dir, "prep/random_forest/zenodo_data/industrial_effort_predictions_by_flag_eez_fao_gear_length_1950_2017.csv"))
  
all_data_grouped_2 <- all_data_grouped %>% 
  group_by(year, flag_country_iso3c, flag_country_name, gear, length_category) %>%
    summarise(modeled_eff_active_fishing_hours = sum(eff_active_fishing_hours, na.rm = TRUE), 
            modeled_eff_active_fishing_days = sum(eff_active_fishing_days, na.rm = TRUE), 
            modeled_nom_active_fishing_hours = sum(nom_active_fishing_hours, na.rm = TRUE), 
            modeled_nom_active_fishing_days = sum(nom_active_fishing_days, na.rm = TRUE)) %>%
  ungroup() %>%
    mutate(flag_country_name = case_when(
          flag_country_iso3c == "RAM" ~ "Madeira",
          flag_country_iso3c == "RAA" ~ "Azores",     
          flag_country_iso3c == "COG" ~ "Republic of the Congo",
      flag_country_iso3c == "COD" ~ "Democratic Republic of the Congo",
          TRUE ~ flag_country_name
  ))

hist_fish_data <- qs::qread(here("data/int/rousseau_gear_fix.qs")) %>%
    ## remove artisanal
    filter(sector == "I") %>%
    group_by(year, flag_fin = country, gear = gear_new, length_category) %>%
    summarize(
      total_nominal_fishing_hours = sum(nom_active_hours, na.rm = TRUE),
      total_effective_fishing_hours = sum(eff_active_hours, na.rm = TRUE),
      total_nominal_fishing_days = sum(nom_active, na.rm = TRUE),
      total_effective_fishing_days = sum(eff_active, na.rm = TRUE)
    ) %>%
  ungroup() %>%
    dplyr::select(flag_country_iso3c = flag_fin, year, gear, length_category, total_nominal_fishing_hours, total_effective_fishing_hours, total_nominal_fishing_days, total_effective_fishing_days) %>%
        filter(total_nominal_fishing_hours > 0) %>%
  mutate(flag_country_name = countrycode(flag_country_iso3c, origin = "iso3c", destination = "country.name")) %>%
  mutate(flag_country_name = case_when(
          flag_country_iso3c == "RAM" ~ "Madeira",
          flag_country_iso3c == "RAA" ~ "Azores", 
                flag_country_iso3c == "COG" ~ "Republic of the Congo",
      flag_country_iso3c == "COD" ~ "Democratic Republic of the Congo",
          TRUE ~ flag_country_name
  )) %>% 
  left_join(all_data_grouped_2) %>%
  mutate(modeled_eff_active_fishing_hours = ifelse(is.na(modeled_eff_active_fishing_hours), 0, modeled_eff_active_fishing_hours),
         modeled_eff_active_fishing_days = ifelse(is.na(modeled_eff_active_fishing_days), 0, modeled_eff_active_fishing_days), 
         modeled_nom_active_fishing_hours = ifelse(is.na(modeled_nom_active_fishing_hours), 0, modeled_nom_active_fishing_hours),
         modeled_nom_active_fishing_days = ifelse(is.na(modeled_nom_active_fishing_days), 0, modeled_nom_active_fishing_days)) %>%
  mutate(proportion_hours_modeled = modeled_nom_active_fishing_hours/total_nominal_fishing_hours)

missing_data <- hist_fish_data %>%
  filter(proportion_hours_modeled != 1) %>%
  filter(proportion_hours_modeled < 0.9999) # should be 0 after our gapfilling. It is, so we actually do not need to save this file anymore. 
 
# fwrite(hist_fish_data, file.path(rdsi_dir, "prep/random_forest/zenodo_data/known_industrial_effort_rousseau.csv"), row.names = FALSE) 


# fwrite(missing_data, file.path(rdsi_dir, "prep/random_forest/zenodo_data/effort_not_modelled.csv"), row.names = FALSE) # similarly, we modelled all effort, so dont need this file anymore

```

How much was gapfilled 

```{r}
all_data_grouped <- fread(file.path(rdsi_dir, "prep/random_forest/zenodo_data/industrial_effort_predictions_by_flag_eez_fao_gear_length_1950_2017.csv"))

gapfill_df <- all_data_grouped %>%
  group_by(flag_country_name, year, model_type) %>%
  summarise(eff = sum(nom_active_fishing_days, na.rm = TRUE)) %>% 
              ungroup() %>%
  pivot_wider(names_from = "model_type", values_from = eff) %>%
  janitor::clean_names() %>%
mutate(across(everything(), ~ ifelse(is.na(.), 0, .)))

sum(gapfill_df$regional_model)/sum(all_data_grouped$nom_active_fishing_hours)

gapfill_df %>% filter(year == 2017) %>% pull(regional_model) %>% sum()/sum(all_data_grouped$nom_active_fishing_hours)


full_predictions <- gapfill_df %>%
  filter(flag_model == 0, regional_model > 0) %>%
  filter(!(flag_country_name %in% c("Azores", "Madeira", "Mayotte", "Dominican Republic")))
sum(full_predictions$regional_model)/sum(all_data_grouped$nom_active_fishing_hours)

low_sample_flags <- readRDS(here("data/model_features/prepared_data_1deg.rds")) %>%
  group_by(flag_fin) %>%
  summarise(n_obs = n()) %>%
  ungroup() %>%
  filter(flag_fin %in% unique(all_data_grouped$flag_country_iso3c)) %>%
  filter(n_obs < 100) %>%
  pull(flag_fin) %>%
  unique()
              

partial_preds <- gapfill_df %>%
  filter(flag_model >0, regional_model > 0)
sum(partial_preds$regional_model)/sum(all_data_grouped$nom_active_fishing_hours)

```



```{r}
test_aus <- fread(file.path(rdsi_dir, "prep/random_forest/zenodo_data/mapped_by_flag_country_industrial/model_preds_1950_2017_GAB.csv"))

test_pixel <- test_aus %>%
  distinct(lon, lat, pixel_area_km2)

test_2 <- test_aus %>%
 filter(year == 2017) %>%
  group_by(lon, lat) %>%
  summarise(eff = sum(nom_active_fishing_hours)) %>%
  ungroup() %>%
  full_join(data_grid_area %>% st_drop_geometry) %>%
  mutate(eff = ifelse(is.na(eff), 0, eff)) %>%
  mutate(eff_km2 = eff/(pixel_area_m2/1000000)) %>%
  dplyr::select(lon, lat, eff_km2) %>%
  rast(., type = "xyz")

plot(test_2)
plot(log(test_2+1))

```

```{r}
# check china. Looks like not enough effort in earlier years? 
test_chn <- fread(file.path(rdsi_dir, "prep/random_forest/zenodo_data/mapped_by_flag_country_industrial/model_preds_1950_2017_CHN.csv"))

 test_chn %>%
   filter(year == 1950) %>%
   pull(nom_active_fishing_hours) %>%
   sum() # 7141969
 
 /home/ubuntu/data_storage/prep/random_forest/predictions_flag_no_access
 
test_chn2 <- qread(file.path(rdsi_dir, "prep/random_forest/predictions_flag_no_access/model_preds_1950_2017_CHN.qs"))

 test_chn2 %>%
   filter(year == 1950) %>%
   pull(nom_active_fishing_hours) %>%
   sum() # 7141969
 # perfect
 
# Industrial notes:
 
# VNM too sparse now
# IDN going back in time
# IND going back in time? 
# Thailand in general 

 
 
```



