---
title: "Extract fishing effort data for case studies"
output: html_document
date: "2024-07-02"
---

# Setup

Load packages and directories

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# remotes::install_github("GlobalFishingWatch/gfwr") #install GFW api package

library(gfwr)
library(terra)
library(tidyverse)
library(here)
library(glue)
library(data.table)
library(janitor)
library(sf)
library(rnaturalearth)
library(countrycode)
library(fasterize)
library(raster)
library(mapview)

source(here("R/dir.R"))


```

# Summary

In this script we choose 3 case study examples of country with high coverage in the global fishing watch data, based on information from Paolo et al., 2024: https://www.nature.com/articles/s41586-023-06825-8#data-availability


# Data Sources 

## Paolo et al. 2024

**Reference**:
Paolo et al. [2024]. https://www.nature.com/articles/s41586-023-06825-8#data-availability

**Downloaded**: July 02, 2024


# Methods 

Identify countries with high coverage and choose a high, medium, and low income country to use for random forest analysis

```{r}

country_coverage <- read.csv(file.path(rdsi_raw_dir, "global_fishing_watch/paolo_et_al/bar_chart_by_country.csv"))

colnames(country_coverage) # perc_dark_fishing is the variable which described the % of fishing in an EEZ which is NOT tracked by GFW. So we will want to choose countries with a low %

good_coverage_iso <- country_coverage %>% 
  dplyr::select(eez_iso3, perc_dark_fishing) %>%
  arrange(perc_dark_fishing)
write.csv(good_coverage_iso, here("data/int/gfw_country_coverage.csv"), row.names = FALSE)

# %>%
#   filter(perc_dark_fishing < 50) %>% 
#   pull(eez_iso3)

## ISL, FRO, SGP, ARG, VUT, NOR, MUS, GGY, HRV, DNK, ESP, LVA, NLD, BEL, IRL, SJM, FRA, SWE, LTU, GBR, PRT, ASM, FLK, ITA, ZAF, URY, GUF, FJI, NCL, SVN, DEU all have perc_dark_fishing <50

## lets start from the top and high, middle, and low income countries. Will need to match to some sort of data describing country classifications; World bank income groups: https://ourworldindata.org/grapher/world-bank-income-groups#:~:text=Lower%2Dmiddle%2Dincome%20countries%20are,%2413%2C846%20or%20more%20in%202022.

income_groups <- read.csv(here("data/raw/world-bank-income-groups.csv")) %>%
  clean_names() %>%
  filter(year == 2020,
         code %in% good_coverage_iso) %>% # filter for max year in gfw and countries with good coverage
  left_join(., country_coverage %>% dplyr::select(code = eez_iso3, perc_dark_fishing)) %>%
  arrange(perc_dark_fishing)

## iceland, argentia, vanuatu? 


```



