---
title: "Create EEZ region ID rasters for GFW cells"
output: html_document
date: "2024-07-02"
---

# Setup

Load packages and directories

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# remotes::install_github("GlobalFishingWatch/gfwr") #install GFW api package

library(gfwr)
library(terra)
library(tidyverse)
library(here)
library(glue)
library(data.table)
library(janitor)
library(sf)
library(rnaturalearth)
library(countrycode)
library(fasterize)
library(raster)
library(mapview)
library(tidyterra)

source(here("R/dir.R"))


```

# Summary


# Data Sources 


# Methods 

```{r}

## read in 2020 data 

effort_2020 <- fread(file.path(rdsi_raw_dir, "global_fishing_watch/raw_mmsi_data/prepped_data/all_effort_2020.csv"))

extent_world <- c(-180, 180, -90, 90)

# Define the resolution
resolution <- 0.1

# Load EEZ shapefile
eez <- st_read(file.path(rdsi_raw_dir, "ohi/regions_mol.shp")) %>%
  st_transform(., crs = "+proj=longlat +elips=WGS84") ## reproject to lat lon to match the GFW data


# Create an empty raster
raster_world <- rast(extent = ext(eez), resolution = resolution, crs = crs(eez)) # match everything to eez shapefile

# Filter for EEZ regions

eez_sf <- eez %>%
  filter(rgn_typ == "eez") %>%
  mutate(rgn_id = as.numeric(rgn_id))

raster_world <- raster(raster_world)

# Rasterize EEZ data
eez_raster <- fasterize(eez_sf, raster_world, field = "rgn_id") # try with fasterize

# Plot the raster
plot(eez_raster)

writeRaster(eez_raster, here("int/eez_raster.tif"), overwrite = TRUE)

eez_rast <- rast(here("int/eez_raster.tif"))

rgn_id_key <- read.csv(here("raw/ohi_region_key.csv")) %>%
  filter(rgn_typ == "eez") %>%
  distinct(rgn_id, rgn_nam, rgn_key)

eez_raster_df <- terra::as.data.frame(eez_rast, xy = TRUE) %>%
  rename(rgn_id = eez_raster) %>%
  left_join(rgn_id_key) %>%
  mutate(x = round(x, 1),
         y = round(y, 1))

fwrite(eez_raster_df, here("data/int/eez_raster_df.csv"), row.names = FALSE) ## cool, this is the df that we need to join with gfw csvs to identify the eez area where the fishing is happening, irrespective of flag country

effort_2020_eez <- effort_2020 %>%
  left_join(eez_raster_df, by = c("cell_ll_lat" = "y", "cell_ll_lon" = "x")) ## i think i need a cell_id to join by instead... lat lon could be a bit off because of rounding.. but this will do for now i suppose

effort_2020_case_study <- effort_2020_eez %>%
  filter(rgn_key %in% c("ARG", "ISL", "VUT")) 

vut_test <- effort_2020_eez %>% 
  filter(rgn_key == "VUT") %>%
  st_as_sf(., coords = c("cell_ll_lon", "cell_ll_lat"), crs = st_crs(eez_sf))

# mapview(head(vut_test, 10000)) ## cool, looks like it worked

## save the case study data 

fwrite(effort_2020_case_study, file.path(rdsi_dir, "prep/random_forest/effort_2020_vut_arg_isl.csv"), row.names = FALSE)


```

```{r}
## lets save a land raster


extent_world <- c(-180, 180, -90, 90)

# Define the resolution
resolution <- 0.1

# Load EEZ shapefile
eez <- st_read(file.path(rdsi_raw_dir, "ohi/regions_mol.shp")) %>%
  st_transform(., crs = "+proj=longlat +elips=WGS84") ## reproject to lat lon to match the GFW data


# Create an empty raster
raster_world <- rast(extent = ext(eez), resolution = resolution, crs = crs(eez)) # match everything to eez shapefile

# Filter for EEZ regions

eez_sf <- eez %>%
  filter(rgn_typ %in% c( "land", "land-noeez")) %>%
  mutate(rgn_id = as.numeric(rgn_id))

raster_world <- raster(raster_world)

# Rasterize EEZ data
eez_raster <- fasterize(eez_sf, raster_world, field = "rgn_id") # try with fasterize

# Plot the raster
plot(eez_raster)

writeRaster(eez_raster, here("data/int/land_raster.tif"), overwrite = TRUE)

```


Look at higher res fishing effort data (see if its actually higher res...) and if it matches the other data 

```{r}

test_mmsi <- read.csv(file.path(rdsi_raw_dir, "global_fishing_watch/apparent_fishing_hours_mmsi/ARG_8466_2020_annual_effort_grid_highres.csv")) 

sum(test_mmsi$apparent_fishing_hours) # 607874.1

test_mmsi_rast <- test_mmsi %>%
  dplyr::select(x, y, apparent_fishing_hours) %>%
  rast(., type = "xyz") # it is higher res...so maybe use this data? need to see how this compares to the other data (distance to port, shore, etc) resolution

test <- read.csv(file.path(rdsi_dir, "prep/random_forest/effort_2020_vut_arg_isl.csv")) %>%
  filter(rgn_key == "ARG") %>%
  dplyr::select(x =  cell_ll_lon, y = cell_ll_lat , fishing_hours) %>%
  rast(., type = "xyz")

# sum(test$fishing_hours, na.rm = TRUE) # 624262.2 - ok its pretty close so thats good

plot(test)

```



Now lets prep the other explanatory variable rasters. We will need to make them all the same resolution (0.1) as the fishing effort data 

Distance to port
 - probably distance to closest port, so will want to take minimum when aggregating

```{r}

dtp <- rast(file.path(rdsi_raw_dir, "global_fishing_watch/distance_from_port/distance-from-port-v1.tiff")) # ok this is 0.01, so it is the same as the higher res data... maybe consider using that instead of the 0.1 res data? Can see how much it changes when I aggregate up to 0.1

dtp_0.1 <- aggregate(dtp, fact = 10, fun = "min") # would mean be approriate? Is the data minimum distance to port from the cell edge or center of cell? May need to resample beforehand? 

writeRaster(dtp_0.1, file.path(rdsi_raw_dir, "global_fishing_watch/distance_from_port/distance-from-port-v1_0.1.tif"), overwrite = TRUE)

par(mfrow = c(1, 2))
plot(dtp)
plot(dtp_0.1)

## looks pretty good i reckon.. would need to mask out the land areas
```


Distance to shore
 - distance to nearest coastline... so probably wanna take the min when aggregating up 
 
```{r}
dts <- rast(file.path(rdsi_raw_dir, "global_fishing_watch/distance_from_shore/distance-from-shore.tif")) # ok this is 0.01, so it is the same as the higher res data... maybe consider using that instead of the 0.1 res data? Can see how much it changes when I aggregate up to 0.1

dts_0.1 <- aggregate(dts, fact = 10, fun = "min") # would mean be approriate? Is the data minimum distance to port from the cell edge or center of cell? May need to resample beforehand? 

writeRaster(dts_0.1, file.path(rdsi_raw_dir, "global_fishing_watch/distance_from_shore/distance-from-shore_0.1.tif"), overwrite = TRUE)

par(mfrow = c(1, 2))
plot(dts)
plot(dts_0.1)

# also looks pretty good 


```


Bathymetry

```{r}
## would probably wanna take the mean when aggregating
bathymetry <- rast(file.path(rdsi_raw_dir, "global_fishing_watch/bathymetry/bathymetry.tif")) ## this is 0.0083 res.. so would need to resample regardless. Will need to mask out land as well

## lets mask out the land now 
land_rast <- rast(here("int/land_raster.tif"))
land_rast[land_rast > 0] <- 1
plot(land_rast)


# bathy_mask <- mask(bathymetry, land_rast %>% project(., bathymetry), inverse = TRUE)
# plot(bathy_mask)

# bathy_0.01 <- resample(bathymetry, dts, method = "bilinear")  # this gonna take forever.... should probably mask out land first 
# bathy_0.05_ag <- aggregate(bathymetry, fact = 6) # aggregate to get near 0.05 resolution.. then disaggregate down to 0.1
# bathy_0.01 <- disagg(bathymetry, fact = 5) # aggregate to get near 0.05 resolution.. then disaggregate down to 0.1


# for 0.1 res we will need to aggregate (fact = 10) and then resample

raster_world

bathy_0.1_ag <- aggregate(bathymetry, fact = 12, fun = "mean") # aggregate to get to 0.1 resolution; I guess terra rounds to 0.1??? This seemed to work.. I'll take the mean
plot(bathy_0.1_ag) # now need to mask out land areas

bathy_ocean <- mask(bathy_0.1_ag, land_rast %>% project(., bathy_0.1_ag), inverse = TRUE) # mask out ocean 

writeRaster(bathy_ocean, file.path(rdsi_dir, "prep/random_forest/bathymetry_0.1_res.tif"), overwrite = TRUE)

plot(bathy_ocean) 

test <- bathy_ocean
test[test > 0] <- NA
plot(test) # oh, ok.. there isn't really any above 0... good

```


SST

 - 0.25 degree Monthly mean SST from NOAA: https://psl.noaa.gov/data/gridded/data.noaa.oisst.v2.highres.html
 - resampled to 0.1 degree grid
 

```{r}
eez_rast <- rast(here("data/int/eez_raster.tif"))

sst <- rast("/home/ubuntu/data_storage/raw_data/noaa/sst/sst.mon.mean.nc")

nlyr(sst) # 514 layers... so 2024 - 1981 = 43*12 = 516.. ok that makes sense. Each layer is a month. What would the 2020 layers be? 

test_df <- sst[[1]] %>% 
  as.data.frame(., xy = TRUE)

# Extract the time attribute
time <- time(sst)

# Extract the year from the time
years <- format(time, "%Y")

indices_2020 <- which(years == "2020")

sst_2020 <- sst[[indices_2020]]

sst_2020_mean <- app(sst_2020, fun = "mean", na.rm = TRUE)

plot(sst_2020_mean)

## ok now we need to resample/disaggregate to 0.1 degrees....

sst_2020_resample <- disagg(sst_2020_mean, fact = 2, method = "bilinear") %>%
  resample(., eez_rast, method = "bilinear")  ## cool that worked! 

writeRaster(sst_2020_resample, file.path(rdsi_raw_dir, "noaa/sst/resample_0.1/sst_2020_0.1.tif"), overwrite = TRUE)

```


CHLa

https://zenodo.org/records/7092220

```{r}

chl_stack <- rast(list.files(file.path(rdsi_raw_dir, "yu_2022/"), full.names = TRUE, pattern = "2022"))

chl_rast_mean <- app(chl_stack, "mean", na.rm = TRUE)

chl_rast_agg <- aggregate(chl_rast_mean, fact = 2, fun = "mean") %>%
  resample(., eez_rast, method = "bilinear")

writeRaster(chl_rast_agg, file.path(rdsi_dir, "prep/random_forest/chla_2020_0.1.tif"), overwrite = TRUE)

```

