---
title: "Spatializing fishing effort predictions based on Global Fishing Watch observed data"
author: "Gage Clawson"
date: "2024-07-04"
output:
  html_document: default
  pdf_document: default
---

# Setup

Load packages and directories

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(terra)
library(tidyverse)
library(here)
library(glue)
library(data.table)
library(janitor)
library(sf)
library(mapview)
library(randomForest)
library(tidyterra)
library(PerformanceAnalytics) ## great correlation plots and other analysis tools
library(MASS) ## stepAIC, etc
library(modEvA) ## Dsquared
library(PerformanceAnalytics) ## great correlation plots and other analysis tools
library(mgcv) ## gams
library(randomForest)
library(gbm) ## brts
library(visreg)
library(raster)


source(here("R/dir.R"))


```

# Summary


# Data Sources 

GFW effort, bathymetry (GEBCO), DTS, DTP, SST, Chla


# Methods 

Fit models to test presence of fishing effort and run random forest to determine variable importance

Predict presence with best fit model

Create rules to spatialise effort into EEZ where industrial effort is mostly likely to be (by gear type and length?)

```{r}

bathy <- rast(file.path(rdsi_dir, "prep/random_forest/bathymetry_0.1_res.tif"))

dts <- rast(file.path(rdsi_raw_dir, "global_fishing_watch/distance_from_shore/distance-from-shore_0.1.tif"))

dtp <- rast(file.path(rdsi_raw_dir, "global_fishing_watch/distance_from_port/distance-from-port-v1_0.1.tif"))

sst <- rast(file.path(rdsi_raw_dir, "noaa/sst/resample_0.1/sst_2020_0.1.tif"))

chla <- rast(file.path(rdsi_dir, "prep/random_forest/chla_2020_0.1.tif"))

effort_subset <- read.csv(file.path(rdsi_dir, "prep/random_forest/effort_2020_vut_arg_isl.csv")) %>%
  rename(x = cell_ll_lon, y = cell_ll_lat)

eez_rast <- rast(here("int/eez_raster.tif"))

eez_rast_df <- fread(here("int/eez_raster_df.csv")) 

rgn_keys <- read.csv(here("raw/ohi_region_key.csv"))

```


Start with Argentina model

Prep all rasters to Argentina's EEZ and have a value for each explanatory variable in each cell 
 
```{r}
eez_arg_df <- eez_rast_df %>%
  filter(rgn_id == 172) %>%
  dplyr::select(x, y) %>%
  mutate(eez = 1)

eez_arg_rast <- rast(eez_arg_df, crs = crs(eez_rast), type = "xyz")


# depth
bathy_arg <- crop(bathy, eez_arg_rast) %>% # crop to correct region
  project(., eez_arg_rast) %>% # make sure extents match
  mask(., eez_arg_rast) # now mask out any areas not in EEZ

## distance to short
dts_arg <- crop(dts, eez_arg_rast) %>% # crop to correct region
  project(., eez_arg_rast) %>% # make sure extents match
  mask(., eez_arg_rast) # now mask out any areas not in EEZ

# distance to port
dtp_arg <- crop(dtp, eez_arg_rast) %>% # crop to correct region
  project(., eez_arg_rast) %>% # make sure extents match
  mask(., eez_arg_rast) # now mask out any areas not in EEZ

# sst
sst_arg <- crop(sst, eez_arg_rast) %>% # crop to correct region
  project(., eez_arg_rast) %>% # make sure extents match
  mask(., eez_arg_rast) # now mask out any areas not in EEZ

chl_arg <- crop(chla, eez_arg_rast) %>% # crop to correct region
  project(., eez_arg_rast) %>% # make sure extents match
  mask(., eez_arg_rast) %>% # now mask out any areas not in EEZ
  rename(chl = mean)


# effort hours
effort_arg <- effort_subset %>% 
  group_by(x, y) %>%
  summarise(fishing_hours = sum(fishing_hours, na.rm = TRUE)) %>%
  ungroup() %>%
  rast(., type = "xyz", crs = crs(eez_arg_rast)) %>%
  crop(., eez_arg_rast) %>%
  project(., eez_arg_rast) %>%
  tidyterra::mutate(fishing_hours = ifelse(is.na(fishing_hours), 0, fishing_hours)) %>%
  mask(., eez_arg_rast) # ok so as.data.frame removes NAs? 


# now stack all rasters

all_rasts <- c(effort_arg, dtp_arg, dts_arg, bathy_arg, sst_arg, chl_arg) %>%
  as.data.frame(., xy = TRUE, na.rm = NA) %>%
  mutate(fis_pres = as.factor(ifelse(fishing_hours > 0, 1, 0))) %>%
  #dplyr::select(-fishing_hours) %>%
  filter(!is.na(bathymetry),
         !is.na(mean),
         !is.na(chl)) %>% # need to filter out NAs for RF to work 
  rename(dtp = 4, dts = 5, sst = mean) 

all_env_stack <- c(dtp_arg, dts_arg, bathy_arg, sst_arg, chl_arg) %>%
    rename(dtp = 1, dts = 2, sst = mean) 

```

Start with random forest model on presence of fishing effort

```{r, include=FALSE}

## first lets look at correlations to see if we should remove any variables
chart.Correlation(all_rasts[,4:8])

## dts and dtp are really highly correlated.. makes a lot of sense - will want to remove one of those. My guess is distance to port is more important for Argentina, considering they have a lot of industrial fishing
all_rasts_2 <- all_rasts %>% dplyr::select(-dts)

chart.Correlation(all_rasts_2[,4:7]) # ok, everything looks good now


## Subdivide into test and training sets
set.seed(123)
sub <- sample(2,nrow(all_rasts),replace=TRUE,prob=c(0.7,0.3))
all_rasts_train <- all_rasts[sub==1,]
all_rasts_test <- all_rasts[sub==2,]



fit <- randomForest(fis_pres~dtp  + bathymetry + sst + chl, data=all_rasts_train, type = "classification")
print(fit)

# Call:
#  randomForest(formula = fis_pres ~ dtp + bathymetry + sst + chl,      data = all_rasts_train, type = "classification") 
#                Type of random forest: classification
#                      Number of trees: 500
# No. of variables tried at each split: 2
# 
#         OOB estimate of  error rate: 10.21%
# Confusion matrix:
#      0    1 class.error
# 0 2997  552  0.15553677
# 1  359 5011  0.06685289

plot(fit)

varImpPlot(fit) # depth, sst, chl, dtp in that order

## Partial dependence plots
opar <- par(mfrow=c(3,2))
partialPlot(fit, all_rasts_train, y)
partialPlot(fit, all_rasts_train, bathymetry)
partialPlot(fit, all_rasts_train, sst)
partialPlot(fit, all_rasts_train, x)
partialPlot(fit, all_rasts_train, dtp)
partialPlot(fit, all_rasts_train, dts)
par(opar)


pr <- predict(fit, newdata=all_rasts_test, type="response")
table(all_rasts_test$fis_pres, pr)

  #  pr
  #      0    1
  # 0 1272  220
  # 1  165 2117

## while this is kind of interesting, we have actual magnitude of fishing hours which is spatialized. A more appropriate method would be to use models to estimate locations/magnitudes of effort in EEZs, based on the variables we have

```

Make models with magnitude of hours 

```{r}

all_rasts <- all_rasts %>%
  filter(fishing_hours > 0)

# all_rasts_stack <- rast(all_rasts, type = "xyz")

## lets check the hours of effort against predictor variables
par(mfrow=c(3,2))
for(i in 4:8){
  plot(all_rasts[,1]~all_rasts[,i], ylab="Effort hours", xlab=names(all_rasts)[i], main=paste0("Effort vs ",names(all_rasts)[i]))
  l <- loess(all_rasts[,1]~all_rasts[,i])
  j <- order(all_rasts[,i]) ## values need to be ordered for the red line to run along and not zigzag
  lines(all_rasts[j,i],l$fitted[j], col="red")
}

## ok so depth is the most obvious polynomial. Potentially chl and sst. Let's just rock with depth to start

## Rounding will introduce only little error, so we'll round the values to be able to use poisson and negative binomial models, and compare each method with each other (e.g. poisson and gaussian model)
all_rasts$fishing_hours <- ceiling(all_rasts$fishing_hours)

##
effort.formula <- formula("fishing_hours ~ bathymetry+sst+chl+I(bathymetry^2)+dtp")
## we also setup the formula for the lognormal model, which uses the log-transformed response:
effort.formula.ln <- formula("log(fishing_hours) ~ bathymetry+sst+chl+I(bathymetry^2)+dtp")

## GLMs

## fitting the model and stepwise dropping variables using stepAIC, and the option "trace=FALSE" to not display each step
fit.n <- stepAIC(glm(effort.formula, data=all_rasts, family=gaussian(link="identity")), trace=FALSE)
## check the summary and diagnostics
summary(fit.n)
par(mfrow=c(2,2))
plot(fit.n)
## AIC is realllllly big. This is a horrible model. 

## lognormal
fit.ln <- stepAIC(glm("log(ceiling(fishing_hours)) ~ bathymetry+sst+chl+I(bathymetry^2)+dtp", data=all_rasts, family=gaussian(link="identity")), trace=FALSE)
summary(fit.ln)
par(mfrow=c(2,2))
plot(fit.ln)

mu <- fitted(fit.ln)
sigma <- summary(fit.ln)$df.residual/nrow(all_rasts)*sqrt(summary(fit.ln)$dispersion)
-2*sum(log(dlnorm(all_rasts$fishing_hours,mu,sigma)))+2*(length(coef(fit.ln))+1)

## this one is a bit better, but still not great. AIC is lower and plots look not as skewed

## poisson

fit.p <- stepAIC(glm(effort.formula, data=all_rasts, family=poisson()), trace=FALSE)
summary(fit.p)
par(mfrow=c(2,2))
plot(fit.p)

## yikes...  by far the worst

fit.rf <- randomForest(effort.formula, data=all_rasts, ntree=500)
fit.rf
par(mfrow=c(2,1))
plot(fit.rf)
varImpPlot(fit.rf) # interesting, dtp is most important according to this. Looks like depth isn't so important anymore (this would probably change if split by gear type, species, etc)

## GAMs - add smooth spatial term

fit.gam <- gam(ceiling(fishing_hours) ~ bathymetry+sst+I(bathymetry^2)+chl+ dtp+ s(x,y), data=all_rasts, family=nb())
summary(fit.gam)
par(mfrow=c(2,2))
plot(fit.gam)

## R2 not great... but not a huge deal. Not really sure how to interpret the plot... 


fit.gam2 <- gam(ceiling(fishing_hours) ~ s(bathymetry)+s(sst)+s(chl)+s(x, y), data=all_rasts, family=nb())
summary(fit.gam2)
par(mfrow=c(2,2))
plot(fit.gam2)

## seems a bit better? Based on R2 and deviance explained... however everythign is significant, so maybe not


```

Make predictions

```{r}
#extract both mean and se of predictions
predfun <- function(model, data) {
  v <- predict(model, data, type="response", se.fit=TRUE)
  cbind(p=as.vector(v$fit), se=as.vector(v$se.fit))
}

## for the lognormal:
predfun.ln <- function(model, data) {
  v <- predict(model, data, type="response", se.fit=TRUE)
  cbind(p=as.vector(exp(v$fit)), se=as.vector(exp(v$fit+v$se.fit)-exp(v$fit)))
}

#for the gams, we need to create a raster of lons and lats to predict on
ra.lon <- raster(all_env_stack$bathymetry)
ra.lon[] <- coordinates(stack(all_env_stack)$bathymetry)[,1]
ra.lat <- raster(all_env_stack$bathymetry)
ra.lat[] <- coordinates(stack(all_env_stack)$bathymetry)[,2]
ra.lon <- rast(ra.lon)
ra.lat <- rast(ra.lat)
all_env_stack$x <- ra.lon
all_env_stack$y <- ra.lat


#generate prediction maps
pred.n <- terra::predict(all_env_stack, fit.n, fun=predfun, index=1:2)
pred.ln <- predict(all_env_stack, fit.ln, fun=predfun.ln, index=1:2)
pred.p <-    predict(all_env_stack, fit.p, fun=predfun, index=1:2)
# pred.nb <-   predict(all_rasts_stack, fit.nb, fun=predfun, index=1:2)
pred.gam <-  predict(all_env_stack, fit.gam, fun=predfun, index=1:2)
pred.gam2 <-  predict(all_env_stack, fit.gam2, fun=predfun, index=1:2)
names(pred.ln) <- names(pred.p) <- names(pred.gam) <- names(pred.gam2) <- c("mean prediction", "std error")

## brt prediction is a bit different here
all_env_stack$bathy2 <- all_env_stack$bathymetry^2

## rf prediction
pred.rf <-   predict(all_env_stack, fit.rf, type="response")

#some plotting parameters:
pt.cex <- c(sqrt(10/pi)/2,sqrt(50/pi)/2,sqrt(100/pi)/2)
cex <- sqrt(all_rasts$fishing_hours/pi)/2
l <- c(10,50,100)
```

Now plot the predictions! 

```{r}
fish_eff_rast <- all_rasts %>% 
  dplyr::select(x, y, fishing_hours) %>%
  rast(., type = "xyz")

## normal
## Plot spatial predictions and overlay observed abundance  
par(mfrow=c(1,3))
## mean
plot(pred.n,1, main="normal - mean")
## standard error
plot(pred.n,2, main="normal - se")
plot(fish_eff_rast, main = "Observed hours")

## lognormal
par(mfrow=c(1,3))
## mean
plot(pred.ln,1, main="lognormal - mean")
## standard error
plot(pred.ln,2, main="lognormal - se")
plot(fish_eff_rast, main = "Observed hours")


## poisson
par(mfrow=c(1,3))
## mean
plot(pred.p,1, main="poisson - mean")
## standard error
plot(pred.p,2, main="poisson - se")
plot(fish_eff_rast, main = "Observed hours")


## gam
par(mfrow=c(1,3))
## mean
plot(pred.gam,1, main="GAM - mean")
## standard error
plot(pred.gam,2, main="GAM - se")
plot(fish_eff_rast, main = "Observed hours")


## gam with smoothing
par(mfrow=c(1,3))
## mean
plot(pred.gam2,1, main="GAM (smoothed) - mean")
## standard error
plot(pred.gam2,2, main="GAM (smoothed) - se")
plot(fish_eff_rast, main = "Observed hours")

# i think this looks the best 


## random forest
par(mfrow=c(1,2))
## mean
plot(pred.rf,1, main="Random Forest - mean")
plot(fish_eff_rast, main = "Observed hours") 

## ok this is the best by far

## is the RF overfit? 
# Is fishing hugging the EEZ? 

```

Do same analysis but with Iceland since they have the best coverage

```{r iceland setup}
eez_isl_df <- eez_rast_df %>%
  filter(rgn_id == 143) %>%
  dplyr::select(x, y) %>%
  mutate(eez = 1)

eez_isl_rast <- rast(eez_isl_df, crs = crs(eez_rast), type = "xyz")


# depth
bathy_isl <- crop(bathy, eez_isl_rast) %>% # crop to correct region
  project(., eez_isl_rast) %>% # make sure extents match
  mask(., eez_isl_rast) # now mask out any areas not in EEZ

## distance to short
dts_isl <- crop(dts, eez_isl_rast) %>% # crop to correct region
  project(., eez_isl_rast) %>% # make sure extents match
  mask(., eez_isl_rast) # now mask out any areas not in EEZ

# distance to port
dtp_isl <- crop(dtp, eez_isl_rast) %>% # crop to correct region
  project(., eez_isl_rast) %>% # make sure extents match
  mask(., eez_isl_rast) # now mask out any areas not in EEZ

# sst
sst_isl <- crop(sst, eez_isl_rast) %>% # crop to correct region
  project(., eez_isl_rast) %>% # make sure extents match
  mask(., eez_isl_rast) # now mask out any areas not in EEZ

chl_isl <- crop(chla, eez_isl_rast) %>% # crop to correct region
  project(., eez_isl_rast) %>% # make sure extents match
  mask(., eez_isl_rast) %>% # now mask out any areas not in EEZ
  rename(chl = mean)


# effort hours
effort_isl <- effort_subset %>% 
  group_by(x, y) %>%
  summarise(fishing_hours = sum(fishing_hours, na.rm = TRUE)) %>%
  ungroup() %>%
  rast(., type = "xyz", crs = crs(eez_isl_rast)) %>%
  crop(., eez_isl_rast) %>%
  project(., eez_isl_rast) %>%
  tidyterra::mutate(fishing_hours = ifelse(is.na(fishing_hours), 0, fishing_hours)) %>%
  mask(., eez_isl_rast) # ok so as.data.frame removes NAs? 


# now stack all rasters

all_rasts <- c(effort_isl, dtp_isl, dts_isl, bathy_isl, sst_isl, chl_isl) %>%
  as.data.frame(., xy = TRUE, na.rm = NA) %>%
  mutate(fis_pres = as.factor(ifelse(fishing_hours > 0, 1, 0))) %>%
  #dplyr::select(-fishing_hours) %>%
  filter(!is.na(bathymetry),
         !is.na(mean),
         !is.na(chl)) %>% # need to filter out NAs for RF to work 
  rename(dtp = 4, dts = 5, sst = mean) 

all_env_stack <- c(dtp_isl, dts_isl, bathy_isl, sst_isl, chl_isl) %>%
    rename(dtp = 1, dts = 2, sst = mean) 

```

```{r corr charts}

## first lets look at correlations to see if we should remove any variables
chart.Correlation(all_rasts[,4:8])

## dts and dtp are really highly correlated.. makes a lot of sense - will want to remove one of those. My guess is distance to port is more important for Argentina, considering they have a lot of industrial fishing
all_rasts_2 <- all_rasts %>% dplyr::select(-dts)

chart.Correlation(all_rasts_2[,4:7]) # ok, now depth and dtp are correlated.. For this simple analysis, I'll keep depth

chart.Correlation(all_rasts_2[,5:7]) # look ok now

```

Make models with magnitude of hours 

```{r modelling}

all_rasts <- all_rasts %>%
  filter(fishing_hours > 0)

## lets check the hours of effort against predictor variables
par(mfrow=c(3,2))
for(i in 4:8){
  plot(all_rasts[,1]~all_rasts[,i], ylab="Effort hours", xlab=names(all_rasts)[i], main=paste0("Effort vs ",names(all_rasts)[i]))
  l <- loess(all_rasts[,1]~all_rasts[,i])
  j <- order(all_rasts[,i]) ## values need to be ordered for the red line to run along and not zigzag
  lines(all_rasts[j,i],l$fitted[j], col="red")
}

## ok so sst is the most obvious polynomial.

## Rounding will introduce only little error, so we'll round the values to be able to use poisson and negative binomial models, and compare each method with each other (e.g. poisson and gaussian model)
all_rasts$fishing_hours <- ceiling(all_rasts$fishing_hours)

##
effort.formula <- formula("fishing_hours ~ bathymetry+sst+chl+I(sst^2)")
## we also setup the formula for the lognormal model, which uses the log-transformed response:
effort.formula.ln <- formula("log(fishing_hours) ~ bathymetry+sst+chl+I(sst^2)")

## GLMs

## fitting the model and stepwise dropping variables using stepAIC, and the option "trace=FALSE" to not display each step
fit.n <- stepAIC(glm(effort.formula, data=all_rasts, family=gaussian(link="identity")), trace=FALSE)
## check the summary and diagnostics
summary(fit.n)
par(mfrow=c(2,2))
plot(fit.n)
## AIC is realllllly big. This is a horrible model. 

## lognormal
fit.ln <- stepAIC(glm("log(ceiling(fishing_hours)) ~ bathymetry+sst+chl+I(sst^2)", data=all_rasts, family=gaussian(link="identity")), trace=FALSE)
summary(fit.ln)
par(mfrow=c(2,2))
plot(fit.ln)

mu <- fitted(fit.ln)
sigma <- summary(fit.ln)$df.residual/nrow(all_rasts)*sqrt(summary(fit.ln)$dispersion)
-2*sum(log(dlnorm(all_rasts$fishing_hours,mu,sigma)))+2*(length(coef(fit.ln))+1)

## this one is a bit better, but still not great. AIC is lower and plots look not as skewed

## poisson

fit.p <- stepAIC(glm(effort.formula, data=all_rasts, family=poisson()), trace=FALSE)
summary(fit.p)
par(mfrow=c(2,2))
plot(fit.p)

## yikes...  still bad


## Negative binomial

fit.nb <- stepAIC(glm.nb(effort.formula, data=all_rasts), trace=FALSE)
summary(fit.nb)
par(mfrow=c(2,2))
plot(fit.nb)
# overfit? 

fit.rf <- randomForest(effort.formula, data=all_rasts, ntree=500)
fit.rf
par(mfrow=c(2,1))
plot(fit.rf) # depth by far most important

## GAMs - add smooth spatial term

fit.gam <- gam(ceiling(fishing_hours) ~ bathymetry+sst+I(sst^2)+chl+ s(x,y), data=all_rasts, family=nb())
summary(fit.gam)
par(mfrow=c(2,2))
plot(fit.gam)

## R2 not great... but not a huge deal. Not really sure how to interpret the plot... 


fit.gam2 <- gam(ceiling(fishing_hours) ~ s(bathymetry)+s(sst)+s(chl)+s(x, y), data=all_rasts, family=nb())
summary(fit.gam2)
par(mfrow=c(2,2))
plot(fit.gam2)

## seems a bit better? Based on R2 and deviance explained... however everythign is significant, so maybe not


```

Make predictions

```{r predictions}
#extract both mean and se of predictions
predfun <- function(model, data) {
  v <- predict(model, data, type="response", se.fit=TRUE)
  cbind(p=as.vector(v$fit), se=as.vector(v$se.fit))
}

## for the lognormal:
predfun.ln <- function(model, data) {
  v <- predict(model, data, type="response", se.fit=TRUE)
  cbind(p=as.vector(exp(v$fit)), se=as.vector(exp(v$fit+v$se.fit)-exp(v$fit)))
}

#for the gams, we need to create a raster of lons and lats to predict on
ra.lon <- raster(all_env_stack$bathymetry)
ra.lon[] <- coordinates(stack(all_env_stack)$bathymetry)[,1]
ra.lat <- raster(all_env_stack$bathymetry)
ra.lat[] <- coordinates(stack(all_env_stack)$bathymetry)[,2]
ra.lon <- rast(ra.lon)
ra.lat <- rast(ra.lat)
all_env_stack$x <- ra.lon
all_env_stack$y <- ra.lat


#generate prediction maps
pred.n <- terra::predict(all_env_stack, fit.n, fun=predfun, index=1:2)
pred.ln <- predict(all_env_stack, fit.ln, fun=predfun.ln, index=1:2)
pred.p <-    predict(all_env_stack, fit.p, fun=predfun, index=1:2)
pred.nb <-   predict(all_env_stack, fit.nb, fun=predfun, index=1:2)
pred.gam <-  predict(all_env_stack, fit.gam, fun=predfun, index=1:2)
pred.gam2 <-  predict(all_env_stack, fit.gam2, fun=predfun, index=1:2)
names(pred.ln) <- names(pred.p) <- names(pred.gam) <- names(pred.gam2) <- c("mean prediction", "std error")

## brt prediction is a bit different here
all_env_stack$bathy2 <- all_env_stack$bathymetry^2

## rf prediction
pred.rf <-   predict(all_env_stack, fit.rf, type="response")

#some plotting parameters:
pt.cex <- c(sqrt(10/pi)/2,sqrt(50/pi)/2,sqrt(100/pi)/2)
cex <- sqrt(all_rasts$fishing_hours/pi)/2
l <- c(10,50,100)
```

Now plot the predictions! 

```{r plotting}
fish_eff_rast <- all_rasts %>% 
  dplyr::select(x, y, fishing_hours) %>%
  rast(., type = "xyz")

## normal
## Plot spatial predictions and overlay observed abundance  
par(mfrow=c(1,3))
## mean
plot(pred.n,1, main="normal - mean")
## standard error
plot(pred.n,2, main="normal - se")
plot(fish_eff_rast, main = "Observed hours")

## lognormal
par(mfrow=c(1,3))
## mean
plot(pred.ln,1, main="lognormal - mean")
## standard error
plot(pred.ln,2, main="lognormal - se")
plot(fish_eff_rast, main = "Observed hours")
# yikes

## poisson
par(mfrow=c(1,3))
## mean
plot(pred.p,1, main="poisson - mean")
## standard error
plot(pred.p,2, main="poisson - se")
plot(fish_eff_rast, main = "Observed hours")


## gam
par(mfrow=c(1,3))
## mean
plot(pred.gam,1, main="GAM - mean")
## standard error
plot(pred.gam,2, main="GAM - se")
plot(fish_eff_rast, main = "Observed hours")


## gam with smoothing
par(mfrow=c(1,3))
## mean
plot(pred.gam2,1, main="GAM (smoothed) - mean")
## standard error
plot(pred.gam2,2, main="GAM (smoothed) - se")
plot(fish_eff_rast, main = "Observed hours")

# i think this looks the best 


## random forest
par(mfrow=c(1,2))
## mean
plot(pred.rf,1, main="Random Forest - mean predictions")
plot(fish_eff_rast, main = "Observed hours") 

## ok this is the best by far but is it overfit? 

```



