---
title: "Match gear and length types between effort data"
output: html_document
date: "2024-07-02"
---

# Setup

Load packages and directories

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(gfwr)
library(terra)
library(tidyverse)
library(here)
library(glue)
library(data.table)
library(janitor)
library(sf)
library(countrycode)
library(raster)

source(here("R/dir.R"))


```

# Summary

In this script we match gear types (aside from trawling gears, which we do in 03 script) in the global fishing watch data and the Rousseau et al., 2024 data. 

# Data Sources 

## Global Fishing Watch

**Reference**: Global Datasets of AIS-based Fishing Effort and Vessel Presence. Version 3.0. 

**Downloaded**: October 10, 2024

**Native resolution**: 0.01 and 0.1 degree 

## IMAS effort data

**Reference**: Rousseau, Y., Blanchard, J.L., Novaglio, C. et al. A database of mapped global fishing activity 1950–2017. Sci Data 11, 48 (2024). https://doi.org/10.1038/s41597-023-02824-6

**Native resolution**: 0.5 degrees, country level 

# Methods 

Load GFW and Rousseau data 

```{r}

# codes used in rousseau et al
country_codes <- read.csv("https://data.imas.utas.edu.au/attachments/1241a51d-c8c2-4432-aa68-3d2bae142794/SAUPtoCountry.csv")

# country level effort data
rousseau_eff <- read.csv("https://data.imas.utas.edu.au/attachments/1241a51d-c8c2-4432-aa68-3d2bae142794/TotalEffortby_FishingCountry_LengthBoat_Gear_Sector.csv") %>%
  filter(!is.na(Length_Category)) %>%
  dplyr::select(-X) %>%
  clean_names()



```


We need to reconcile the gfw data categories to those of the Rousseau categories. 
 - Some of the gears have different names, we will combine them as best we can. 
 - We need to split the "trawlers" category from GFW based on proportion of bottom vs midwater trawls reported in Rousseau et al. 
 - GFW reports length as a numeric but Rousseau is category

Get country level trawling proportions from Rousseau et al. We want to know the proportion of a flag countries' operations which are bottom vs mid water trawling. We'll use this information later on the allocate the GFW "trawlers" category to bottom vs midwater. 

```{r}

## filter rousseau data for trawlers 

unique(rousseau_eff$gear)

trawling_eff <- rousseau_eff %>% 
  filter(gear %in% c("Trawl_Bottom", "Trawl_Midwater_or_Unsp")) %>%
  pivot_wider(names_from = "gear", values_from = "nv") %>%
  group_by(saup, year) %>%
  summarise(Trawl_Bottom = sum(Trawl_Bottom, na.rm = TRUE), 
            Trawl_Midwater_or_Unsp = sum(Trawl_Midwater_or_Unsp, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(prop_bottom_trawls = Trawl_Bottom/(Trawl_Bottom+Trawl_Midwater_or_Unsp),
         prop_mid_trawls = Trawl_Midwater_or_Unsp/(Trawl_Bottom+Trawl_Midwater_or_Unsp)) %>% 
  dplyr::select(saup, year, prop_bottom_trawls, prop_mid_trawls)

qs::qsave(trawling_eff, here("data/int/prop_trawling_year_country.qs")) # save 

```


Group gfw data by cell, year, flag, gear (after resolving gear issues i.e., grouping into other, don't worry about trawling yet), length category 
   - sum of engine power?, gt?, hours, and fishing hours

```{r}


vessel_info <- read.csv(file.path(rdsi_raw_dir, "global_fishing_watch/fishing-vessels-v2.csv")) # unfortunately have to use v2 vessel info right now, as I can't find v3 on the API

## fix rousseau gear categories. Put ones that don't match into an "Other" category 
## "Falling_Gear", "Lift_Nets", "Lines_Unspecified", "Others_Multiple_Gears", "Others_Others", "Others_Support", "Others_Unknown" ~ "Other"

rousseau_gear_fix <- rousseau_eff %>%
  mutate(gear_new = ifelse(gear %in% c("Falling_Gear", "Lift_Nets", "Lines_Unspecified", "Others_Multiple_Gears", "Others_Others", "Others_Support", "Others_Unknown"), "Other", gear)) %>%
  group_by(saup, year, sector, gear_new, length_category, country) %>% 
  summarise(nom_active = sum(nom_active, na.rm = TRUE),
            eff_active = sum(eff_active, na.rm = TRUE),
            nv = sum(nv, na.rm = TRUE), 
            p = sum(p, na.rm = TRUE), 
            gt = sum(gt, na.rm = TRUE), 
            nom_active_hours = sum(nom_active_hours, na.rm = TRUE), 
            eff_active_hours = sum(eff_active_hours, na.rm = TRUE)) %>%
  ungroup()

qsave(rousseau_gear_fix, here("data/int/rousseau_gear_fix.qs"))

sort(unique(rousseau_gear_fix$gear_new)) # look at gears in rousseau

#  [1] "Dredges"                   "Gillnets"                  "Lines_Handlines_and_poles" "Lines_Longlines"           "Other"                     "Pots_and_Traps"           
#  [7] "Seine_Danish_and_Other"    "Seine_Purse_Seine"         "Trawl_Bottom"              "Trawl_Midwater_or_Unsp"   


# GFW Geartypes:
# - fishing: a combination of vessels of unknown fishing gear
#  - drifting_longlines: drifting longlines
#  - seiners: vessels using seine nets, including potential purse seine vessels
#    targeting tuna and other species, as well as danish and other seines
#      - purse_seines: purse seines, both pelagic and demersal
#         - tuna_purse_seines: large purse seines primarily fishing for tuna.
#         - other_purse_seines: purse seiners fishing for mackerel, anchovies, etc, often smaller and operating nearer the coast than tuna purse seines.
#     - other_seines: danish seines and other seiners not using purse seines.
#  - trawlers: trawlers, all types
#  - pole_and_line: vessel from which people fish with pole and line.
#  - trollers: vessel that tows multiple fishing lines.
#  - fixed_gear: a category that includes potential set longlines, set gillnets,  and pots and traps
#      - pots_and_traps: vessel that deploys pots (small, portable traps) or traps to
#        catch fish
#      - set_longlines: vessel that fishes by setting longlines anchored to the
#        seafloor. These lines have shorter hooked, typically baited, lines hanging
#        from them
#      - set_gillnets: vessel that fishes by setting gillnets anchored to the seafloor.
#  - dredge_fishing: vessel that tows a dredge the scrapes up edible bottom. Fall under "Hooks and Lines" per ISSCFG
#    dwellers such as scallops or oysters.
#  - squid_jigger: squid jiggers, mostly large industrial pelagic operating vessels

## ok so based on the gear types, we're going to have to do some combining... both for GFW and IMAS data

sort(unique(vessel_info$vessel_class_gfw))

gfw_fix <- vessel_info %>%
  mutate(length_category = case_when(
    length_m_gfw < 6  ~ "Less than 6m",
    length_m_gfw >= 6 & length_m_gfw < 12 ~ "6-12m",
    length_m_gfw >= 12 & length_m_gfw < 24 ~ "12-24m",
    length_m_gfw >= 24 & length_m_gfw <= 50 ~ "24-50m",
    length_m_gfw > 50 ~ "Over 50m",
    TRUE ~ NA
  )) %>%
  mutate(gear = case_when(
    vessel_class_gfw %in% c("dredge_fishing") ~ "Dredges",
    vessel_class_gfw %in% c("set_gillnets") ~ "Gillnets",
    vessel_class_gfw %in% c("pole_and_line", "squid_jigger", "trollers") ~ "Lines_Handlines_and_poles",
    vessel_class_gfw %in% c("drifting_longlines", "set_longlines") ~ "Lines_Longlines",
    vessel_class_gfw %in% c("fishing", "fixed_gear") ~ "Other",
    vessel_class_gfw %in% c("pots_and_traps") ~ "Pots_and_Traps",
    vessel_class_gfw %in% c("other_seines") ~ "Seine_Danish_and_Other",
    vessel_class_gfw %in% c("other_purse_seines", "purse_seines", "tuna_purse_seines", "seiners") ~ "Seine_Purse_Seine",
    TRUE ~ vessel_class_gfw 
   ))

unique(gfw_fix$length_category) # [1] "24-50m"       "Over 50m"     "12-24m"       "6-12m"        NA             "Less than 6m"


## lets fix the na lengths by filling them in with the most common length in that country
test_lengths <- gfw_fix %>%
  group_by(flag_gfw, length_category) %>%
  summarise(count = n()) %>%
    group_by(flag_gfw) %>%
  filter(count == max(count)) %>%
  ungroup() %>%
  dplyr::select(flag_gfw, most_common_length = length_category)


missing_lengths <- gfw_fix %>%
  filter(is.na(length_category)) %>%
  left_join(test_lengths) %>%
  mutate(length_category = ifelse(is.na(length_category), most_common_length, length_category)) %>%
  dplyr::select(-most_common_length)


gfw_fix_fin <- gfw_fix %>%
  filter(!is.na(length_category)) %>%
  rbind(missing_lengths)
  
  
qs::qsave(gfw_fix_fin, here("data/int/gfw_vessel_registry_fix.qs")) # now save this file so we can read back in later and join to gfw data
  
```


