---
title: "Re-save global fishing watch data with new gear types and proportions in each cell"
output: html_document
date: "2024-07-02"
---

# Setup

Load packages and directories

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(gfwr)
library(terra)
library(tidyverse)
library(here)
library(glue)
library(data.table)
library(janitor)
library(sf)
library(countrycode)
library(raster)
library(qs)
library(doParallel)
library(foreach)

source(here("R/dir.R"))

```

# Summary

We take GFW data and calculate for each gear type, vessel length, flag country, the proportion of catch occurring in that cell out of the total for those categories. 

We also split the trawling category by proportionally allocating GFW trawling data into bottom and mid-water trawls. 


# Data Sources 

## Global Fishing Watch

**Reference**: Global Datasets of AIS-based Fishing Effort and Vessel Presence. Version 2.0 and 3.0. 

**Downloaded**: October 10, 2024

**Native resolution**: 0.1 degree 

Global Fishing Watch. 2025. Global Apparent Fishing Effort Dataset, Version 3.0. doi:10.5281/zenodo.14982712


## IMAS effort data

**Reference**: Rousseau, Y., Blanchard, J.L., Novaglio, C. et al. A database of mapped global fishing activity 1950–2017. Sci Data 11, 48 (2024). https://doi.org/10.1038/s41597-023-02824-6

**Native resolution**: 0.5 degrees, country level 


# Methods 

Load GFW vessel data with new gear types and load Rousseau trawling proportions

```{r}

gfw_vessel_registry <- qread(here("data/int/gfw_vessel_registry_fix.qs")) %>%
  dplyr::distinct(mmsi, flag_gfw, length_category, gear, tonnage_gt_gfw, engine_power_kw_gfw)

saup_table <- read.csv("https://data.imas.utas.edu.au/attachments/1241a51d-c8c2-4432-aa68-3d2bae142794/SAUPtoCountry.csv") %>% clean_names()

rousseau_trawl_props <- qs::qread(here("data/int/prop_trawling_year_country.qs")) %>%
  left_join(saup_table) %>%
  dplyr::select(-x, -saup, -region)

rousseau_gear_fix <- qs::qread(here("data/int/rousseau_gear_fix.qs"))

eez_lookup <- read.csv(here("data/raw/spatial/eez_lookup.csv")) %>%
  clean_names() # made in /R/spatial_files.R

```

Re-save global fishing watch yearly data without vessel MMSI identifiers

Group gfw data by cell, year, flag, gear, length category 
   - sum of engine power, gt, hours, and fishing hours

```{r}

most_common_lengths <- gfw_vessel_registry %>%
  group_by(flag_gfw, length_category) %>%
  summarise(count = n()) %>%
    group_by(flag_gfw) %>%
  filter(count == max(count)) %>%
  ungroup() %>%
  dplyr::select(flag_fin = flag_gfw, most_common_length = length_category)

gear_lengths_iso3c <- gfw_vessel_registry %>%
  group_by(flag_gfw, gear, length_category) %>%
  summarise(count = n()) %>%
    group_by(flag_gfw, length_category) %>%
  filter(count == max(count)) %>%
  ungroup() %>%
  dplyr::select(flag_fin = flag_gfw, length_category, most_common_gear_l_flag = gear)

gear_iso3c <- gfw_vessel_registry %>%
  group_by(flag_gfw, gear) %>%
  summarise(count = n()) %>%
    group_by(flag_gfw) %>%
  filter(count == max(count)) %>%
  ungroup() %>%
  dplyr::select(flag_fin = flag_gfw, most_common_gear_flag = gear)
  
cl <- makeCluster(3)
registerDoParallel(cl)

foreach(yr = 2013:2024, .packages = c(
  "qs", "here", "dplyr", "tidyr", "glue", "terra", "countrycode"
)) %dopar% {

 #  yr = 2012

eez_rgns <- qread(here("data/raw/spatial/eez_rast_csv_0.1.qs")) # made in spatial.R script

gfw_year <- qread(file.path(rdsi_raw_dir, glue("global_fishing_watch/apparent_fishing_hours_mmsi/v3_aggregated/all_effort_{yr}.qs"))) %>%
  dplyr::select(x = cell_ll_lon, y = cell_ll_lat,  mmsi, hours, fishing_hours) %>%
  left_join(gfw_vessel_registry) %>%
  rename(flag_fin = flag_gfw)
  

# test <- gfw_year %>%
#   group_by(x, y) %>%
#   summarise(hours = sum(fishing_hours, na.rm = TRUE)) %>%
#   ungroup()
# 
# test_rast <- rast(test, type = "xyz")
 
# test <- gfw_year_new_gears %>%
#   filter(is.na(gear))

  ## lets fix the NA lengths by filling them in with the most common length in that country
missing_lengths <- gfw_year %>%
  filter(is.na(length_category)) %>%
  left_join(most_common_lengths) %>%
  mutate(length_category = ifelse(is.na(length_category), most_common_length, length_category)) %>%
  dplyr::select(-most_common_length)

## now join back together and group by new gear and length categories (after making sure there are no NAs)
gfw_length_fix <- gfw_year %>%
  filter(!is.na(length_category)) %>%
  rbind(missing_lengths) 

  ## lets fix the missing gears by filling them in with the most common gear type within the country and length category

missing_gears <- gfw_length_fix %>%
  filter(is.na(gear)) %>%
  left_join(gear_lengths_iso3c) %>%
  left_join(gear_iso3c) %>%
    mutate(gear = ifelse(is.na(gear), most_common_gear_l_flag, gear)) %>%
   mutate(gear = ifelse(is.na(gear), most_common_gear_flag, gear))  %>%
  dplyr::select(-most_common_gear_l_flag, -most_common_gear_flag)

## now join back together and group by new gear and length categories (after making sure there are no NAs)
gfw_gear_fix <- gfw_length_fix %>%
  filter(!is.na(gear)) %>%
  rbind(missing_gears)  %>%
  mutate(year = yr)


gfw_fix <- gfw_gear_fix %>%
  dplyr::filter(gear != "trawlers" | is.na(gear)) %>% # filter out any remaining missing gears and trawlers (we fix trawlers later)
    group_by(year, flag_fin) %>%
  mutate(tonnage_gt_mean = mean(tonnage_gt_gfw, na.rm = TRUE),
         engine_power_mean = mean(engine_power_kw_gfw, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(engine_power_kw_gfw = ifelse(is.na(engine_power_kw_gfw), engine_power_mean, engine_power_kw_gfw),
         tonnage_gt_gfw = ifelse(is.na(tonnage_gt_gfw), tonnage_gt_mean, tonnage_gt_gfw)) %>% # fill in any missing engine or tonnage if possible
  group_by(x, y, year, flag_fin, gear, length_category) %>%
  summarise(
            fishing_hours = sum(fishing_hours, na.rm = TRUE), 
            engine_power_kw_gfw = sum(engine_power_kw_gfw, na.rm = TRUE),
            tonnage_gt_gfw = sum(tonnage_gt_gfw),
            nv = n_distinct(mmsi)) %>% ## just sum all of these categories for now
  ungroup() %>%
  filter(fishing_hours > 0)

# test <- filter(gfw_fix, is.na(flag_fin)) # not very many left for 2012, which is good! 
      
## ok there are some missing flags... maybe just assume it is the country where it is being fished (the eez)?

# we are using the flag country proportions of bottom trawls vs mid trawls in their fleets, calculated in 01_fix_categories.Rmd

gfw_fix_trawl <- gfw_gear_fix %>% 
  filter(gear == "trawlers") %>% 
  filter(!is.na(length_category)) %>%
  rbind(missing_lengths %>% filter(gear == "trawlers")) %>%
    group_by(x, y, year, flag_fin, gear, length_category) %>%
  summarise(
            fishing_hours = sum(fishing_hours, na.rm = TRUE), 
            engine_power_kw_gfw = sum(engine_power_kw_gfw, na.rm = TRUE),
            tonnage_gt_gfw = sum(tonnage_gt_gfw),
            nv = n_distinct(mmsi)) %>% ## just sum all of these categories for now
  ungroup() %>%
  left_join(rousseau_trawl_props, by = c("year" = "year", "flag_fin" = "country")) %>%
  mutate(
         bt_fishing_hours = fishing_hours*prop_bottom_trawls,
         bt_engine_power = engine_power_kw_gfw*prop_bottom_trawls,
         bt_tonnage_gt = tonnage_gt_gfw*prop_bottom_trawls,
         bt_nv = nv*prop_bottom_trawls,
         mt_fishing_hours = fishing_hours*prop_mid_trawls,
         mt_engine_power = engine_power_kw_gfw*prop_mid_trawls,
         mt_tonnage_gt = tonnage_gt_gfw*prop_mid_trawls,
         mt_nv = nv*prop_mid_trawls)
  
## filter for bottom trawl
bottom_trawl_fix <- gfw_fix_trawl %>%
  dplyr::select(x, y, year, flag_fin, length_category, fishing_hours = bt_fishing_hours, engine_power_kw_gfw = bt_engine_power, tonnage_gt_gfw = bt_tonnage_gt, nv = bt_nv) %>%
  mutate(gear = "Trawl_Bottom")

# filter for midwater trawl
mid_trawl_fix <- gfw_fix_trawl %>%
  dplyr::select(x, y, year, flag_fin, length_category, fishing_hours = mt_fishing_hours, engine_power_kw_gfw = mt_engine_power, tonnage_gt_gfw = mt_tonnage_gt, nv = mt_nv) %>%
  mutate(gear = "Trawl_Midwater_or_Unsp")


gfw_fix_props <- gfw_fix %>%
  rbind(., bottom_trawl_fix) %>%
  rbind(., mid_trawl_fix) %>%
  group_by(year, flag_fin, gear, length_category) %>%
  mutate(
            fishing_hours_global = sum(fishing_hours, na.rm = TRUE), 
            engine_power_kw_gfw_global = sum(engine_power_kw_gfw, na.rm = TRUE),
            tonnage_gt_gfw_global = sum(tonnage_gt_gfw),
            nv_global = sum(nv)) %>%
  ungroup() %>%
  mutate(
         prop_fishing_hours_cell = fishing_hours/fishing_hours_global,
         prop_engine_power_kw_cell = engine_power_kw_gfw/engine_power_kw_gfw_global,
         prop_tonnage_gt_cell = tonnage_gt_gfw/tonnage_gt_gfw_global,
         prop_nv_cell = nv/nv_global) %>%
  filter(!is.na(prop_fishing_hours_cell)) %>% # filter out any NA because there is no effort in those cells (these are the ones where there was either no bottom or mid trawls based on props)
  left_join(eez_rgns) %>% 
  left_join(eez_lookup) %>%
    mutate(flag_fin = ifelse(flag_fin == ""|flag_fin == "UNK" | is.na(flag_fin), iso_sov1, flag_fin)) %>% # fill in flag with where the fishing occurs, if it is unknown.  
  mutate(flag_name = countrycode(sourcevar = flag_fin, origin = "iso3c", destination = "country.name")) %>%
  mutate(flag_fin = ifelse(flag_fin %in% c("HSX", "LND"), "UNK", flag_fin),
         flag_name = ifelse(flag_fin %in% c("UNK"), "Unknown", flag_name)) %>% # %>% # ok these cells are all either on land or in the high seas and don't report the flag country... so we just assume they are unknown 
  # filter(fishing_hours > 0) # if you filter out 0 fishing hours, there are no NAs. I'll keep 0's in for now
  dplyr::select(-mrgid_sov1, -iso_ter1, -geoname, -fishing_hours_global, -nv_global, -engine_power_kw_gfw_global, -tonnage_gt_gfw_global) %>%
  dplyr::select(x, y, everything()) %>%
  filter(fishing_hours > 0)

# test_na <- gfw_fix_props %>%
#     filter(is.na(flag_name)) # none
# 
# test_0 <- gfw_fix_props %>%
#   filter(fishing_hours == 0)
# 
# test_na <- gfw_fix_props %>%
#     filter(is.na(mrgid) & is.na(flag_name)) # none

qs::qsave(gfw_fix_props, file.path(rdsi_dir, glue("prep/gfw_props/deg_10/all_effort_gear_length_props_{yr}.qs")))

  }

stopCluster(cl)

# quick data check

test <- qs::qread(file.path(rdsi_dir, "prep/gfw_props/deg_10/all_effort_gear_length_props_2012.qs"))
sum(test$fishing_hours)
test_na <- test %>%
  filter(is.na(flag_fin))

test2 <- test %>%
  filter(fishing_hours > 0) %>%
  distinct(x, y) %>%
  mutate(new = 1) %>%
  rast(., type = "xyz")

test3 <- gfw_year %>%
  filter(fishing_hours > 0) %>%
  distinct(x,y) %>%
    mutate(new = 1) %>%
  rast(., type = "xyz")

plot(test3)

```


