---
title: "Aggregate GFW data to 0.5 degrees cells"
output: html_document
date: "2024-12-12"
---

# Setup

Load packages and directories

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(gfwr)
library(terra)
library(tidyverse)
library(here)
library(glue)
library(data.table)
library(janitor)
library(sf)
library(countrycode)
library(raster)
library(qs)

source(here("R/dir.R"))


```

# Summary

We aggregate the gfw data from 0.1 to 0.5 degree cells. To do this, we have to run really similar code to that of 03 script, because we aggregate first and then change the gear types. 

# Data Sources 

## Global Fishing Watch

**Reference**: Global Datasets of AIS-based Fishing Effort and Vessel Presence. Version 2.0 and 3.0. 

**Downloaded**: October 10, 2024

**Native resolution**: 0.01 and 0.1 degree 

## IMAS effort data

**Reference**: Rousseau, Y., Blanchard, J.L., Novaglio, C. et al. A database of mapped global fishing activity 1950–2017. Sci Data 11, 48 (2024). https://doi.org/10.1038/s41597-023-02824-6

**Native resolution**: 0.5 degrees, country level 


# Methods 

Load GFW vessel data with new gear types and load Rousseau trawling proportions

```{r}

gfw_vessel_registry <- qread(here("data/int/gfw_vessel_registry_fix.qs")) %>%
  dplyr::select(mmsi, flag_gfw, vessel_class_gfw, length_m_gfw, length_category, gear, tonnage_gt_gfw, engine_power_kw_gfw)

saup_table <- read.csv("https://data.imas.utas.edu.au/attachments/1241a51d-c8c2-4432-aa68-3d2bae142794/SAUPtoCountry.csv") %>% clean_names()

rousseau_trawl_props <- qs::qread(here("data/int/prop_trawling_year_country.qs")) %>%
  left_join(saup_table) %>%
  dplyr::select(-x, -saup, -region)

rousseau_gear_fix <- qs::qread(here("data/int/rousseau_gear_fix.qs"))

eez_lookup <- read.csv(here("data/raw/spatial/eez_lookup.csv")) %>%
  clean_names() # made in /R/spatial_files.R

```

Re-save global fishing watch yearly data without vessel MMSI identifiers

Group gfw data by cell, year, flag, gear (after resolving gear issues i.e., grouping into other, don't worry about trawling yet), length category 
   - sum of engine power, gt, hours, and fishing hours

```{r}

test_lengths <- gfw_vessel_registry %>%
  group_by(flag_gfw, length_category) %>%
  summarise(count = n()) %>%
    group_by(flag_gfw) %>%
  filter(count == max(count)) %>%
  ungroup() %>%
  dplyr::select(flag_gfw, most_common_length = length_category)

  
for(y in 2012:2017){

 #  y = 2012
    
eez_rgns <- qread(here("data/raw/spatial/eez_rast_csv_0.1.qs"))

gfw_year <- read.csv(file.path(rdsi_raw_dir, glue("global_fishing_watch/apparent_fishing_hours_mmsi/v2_aggregated/all_effort_{y}.csv"))) %>%
  rename(flag = flag_gfw) %>%
    rename(x = cell_ll_lon, y = cell_ll_lat) %>%
  left_join(gfw_vessel_registry) %>%
  mutate(flag_fin = ifelse(is.na(flag), flag_gfw, flag)) %>% # fix missing flags... use API flag, and if there is no API flag, use flag from vessel registry... if it is still missing, we'll fix it later on based on where the fishing is actually happening
  group_by(x, y, flag_fin, vessel_class_gfw, length_m_gfw) %>%
  summarise(fishing_hours = sum(fishing_hours, na.rm = TRUE),
            hours = sum(hours, na.rm = TRUE), 
            engine_power_kw_gfw = sum(engine_power_kw_gfw, na.rm = TRUE),
            tonnage_gt_gfw = sum(tonnage_gt_gfw, na.rm = TRUE),
            nv = n_distinct(mmsi)) %>% 
  ungroup()

# Function to round to nearest 0.5
round_to_half <- function(x) {
  round(x * 2) / 2
}

# Aggregate from 0.1 to 0.5 degree cells
gfw_year_aggregated <- gfw_year %>%
  # Round coordinates to nearest 0.5 degree
  mutate(
    x_05 = round_to_half(x),
    y_05 = round_to_half(y)
  ) %>%
  # Group by rounded coordinates and existing grouping variables
  group_by(x_05, y_05, flag_fin, vessel_class_gfw, length_m_gfw) %>%
  # Aggregate numeric columns
  summarise(
    fishing_hours = sum(fishing_hours, na.rm = TRUE),
    hours = sum(hours, na.rm = TRUE),
    engine_power_kw_gfw = sum(engine_power_kw_gfw, na.rm = TRUE),
    tonnage_gt_gfw = sum(tonnage_gt_gfw, na.rm = TRUE), # not sure if it is appropraite to sum these? Idk if we even need them...
    nv = sum(nv) # Sum the vessel counts
  ) %>%
  ungroup() %>%
  # Rename coordinates back to original names if needed
  rename(x = x_05, y = y_05)
    

# test <- gfw_year_aggregated %>%
#   group_by(x, y) %>%
#   summarise(fishing_hours = sum(fishing_hours, na.rm = TRUE)) %>%
#   ungroup() %>%
#   rast(., type ="xyz") # cool that worked! 

gfw_year_new_gears <-  gfw_year_aggregated %>% 
  mutate(length_category = case_when(
    length_m_gfw < 6  ~ "Less than 6m",
    length_m_gfw >= 6 & length_m_gfw < 12 ~ "6-12m",
    length_m_gfw >= 12 & length_m_gfw < 24 ~ "12-24m",
    length_m_gfw >= 24 & length_m_gfw <= 50 ~ "24-50m",
    length_m_gfw > 50 ~ "Over 50m",
    TRUE ~ NA
  )) %>%
  mutate(gear = case_when(
    vessel_class_gfw %in% c("dredge_fishing") ~ "Dredges",
    vessel_class_gfw %in% c("set_gillnets") ~ "Gillnets",
    vessel_class_gfw %in% c("pole_and_line", "squid_jigger", "trollers") ~ "Lines_Handlines_and_poles",
    vessel_class_gfw %in% c("drifting_longlines", "set_longlines") ~ "Lines_Longlines",
    vessel_class_gfw %in% c("fishing", "fixed_gear") ~ "Other",
    vessel_class_gfw %in% c("pots_and_traps") ~ "Pots_and_Traps",
    vessel_class_gfw %in% c("other_seines") ~ "Seine_Danish_and_Other",
    vessel_class_gfw %in% c("other_purse_seines", "purse_seines", "tuna_purse_seines", "seiners") ~ "Seine_Purse_Seine",
    TRUE ~ vessel_class_gfw
   ))

  ## lets fix the NA lengths by filling them in with the most common length in that country


missing_lengths <- gfw_year_new_gears %>%
  filter(is.na(length_category)) %>%
  left_join(test_lengths, by = c("flag_fin" = "flag_gfw")) %>%
  mutate(length_category = ifelse(is.na(length_category), most_common_length, length_category)) %>%
  dplyr::select(-most_common_length)

## now join back together and group by new gear and length categories (after making sure there are no NAs)
gfw_length_fix <- gfw_year_new_gears %>%
  filter(!is.na(length_category)) %>%
  rbind(missing_lengths) 

  ## lets fix the missing gears by filling them in with the most common gear type within the country and length category

gear_lengths_iso3c <- gfw_vessel_registry %>%
  group_by(flag_gfw, gear, length_category) %>%
  summarise(count = n()) %>%
    group_by(flag_gfw, length_category) %>%
  filter(count == max(count)) %>%
  ungroup() %>%
  dplyr::select(flag_fin = flag_gfw, length_category, most_common_gear_l_flag = gear)

gear_iso3c <- gfw_vessel_registry %>%
  group_by(flag_gfw, gear) %>%
  summarise(count = n()) %>%
    group_by(flag_gfw) %>%
  filter(count == max(count)) %>%
  ungroup() %>%
  dplyr::select(flag_fin = flag_gfw, most_common_gear_flag = gear)


missing_gears <- gfw_length_fix %>%
  filter(is.na(gear)) %>%
  left_join(gear_lengths_iso3c) %>%
  left_join(gear_iso3c) %>%
    mutate(gear = ifelse(is.na(gear), most_common_gear_l_flag, gear)) %>%
   mutate(gear = ifelse(is.na(gear), most_common_gear_flag, gear))  %>%
  dplyr::select(-most_common_gear_l_flag, -most_common_gear_flag)

## now join back together and group by new gear and length categories (after making sure there are no NAs)
gfw_gear_fix <- gfw_length_fix %>%
  filter(!is.na(gear)) %>%
  rbind(missing_gears) 


gfw_fix <- gfw_gear_fix %>% ## ok still have ~400 observations missing gear for 2012 all HTI or BTN.
  dplyr::filter(gear != "trawlers" | is.na(gear)) %>%
    group_by(flag_fin) %>%
  mutate(tonnage_gt_mean = mean(tonnage_gt_gfw, na.rm = TRUE),
         engine_power_mean = mean(engine_power_kw_gfw, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(engine_power_kw_gfw = ifelse(is.na(engine_power_kw_gfw), engine_power_mean, engine_power_kw_gfw),
         tonnage_gt_gfw = ifelse(is.na(tonnage_gt_gfw), tonnage_gt_mean, tonnage_gt_gfw)) %>% # fill in any missing engine or tonnage if possible
  group_by(x, y, flag_fin, gear, length_category) %>%
  summarise(
            fishing_hours = sum(fishing_hours, na.rm = TRUE), 
            engine_power_kw_gfw = sum(engine_power_kw_gfw, na.rm = TRUE),
            tonnage_gt_gfw = sum(tonnage_gt_gfw, na.rm = TRUE),
            nv = sum(nv, na.rm = TRUE)) %>% ## just sum all of these categories for now
  ungroup() 

      
## ok there are some missing flags... maybe just assume it is the country where it is being fished (the eez)?

# we are using the flag country proportions of bottom trawls vs mid trawls in their fleets, calculated in 01_fix_categories.Rmd

gfw_fix_trawl <- gfw_gear_fix %>% 
  filter(gear == "trawlers") %>% 
  filter(!is.na(length_category)) %>%
  rbind(missing_lengths %>% filter(gear == "trawlers")) %>%
    group_by(x, y, flag_fin, gear, length_category) %>%
  summarise(
            fishing_hours = sum(fishing_hours, na.rm = TRUE), 
            engine_power_kw_gfw = sum(engine_power_kw_gfw, na.rm = TRUE),
            tonnage_gt_gfw = sum(tonnage_gt_gfw, na.rm = TRUE),
            nv = sum(nv, na.rm = TRUE)) %>% ## just sum all of these categories for now
  ungroup() %>%
  left_join(rousseau_trawl_props %>% filter(year == y), by = c("flag_fin" = "country")) %>%
  mutate(
    # bottom trawls
         bt_fishing_hours = fishing_hours*prop_bottom_trawls,
         bt_engine_power = engine_power_kw_gfw*prop_bottom_trawls,
         bt_tonnage_gt = tonnage_gt_gfw*prop_bottom_trawls,
         bt_nv = nv*prop_bottom_trawls,
   # midwater trawls
         mt_fishing_hours = fishing_hours*prop_mid_trawls,
         mt_engine_power = engine_power_kw_gfw*prop_mid_trawls,
         mt_tonnage_gt = tonnage_gt_gfw*prop_mid_trawls,
         mt_nv = nv*prop_mid_trawls)
  
## filter for bottom trawl
bottom_trawl_fix <- gfw_fix_trawl %>%
  dplyr::select(x, y, flag_fin, length_category, fishing_hours = bt_fishing_hours, engine_power_kw_gfw = bt_engine_power, tonnage_gt_gfw = bt_tonnage_gt, nv = bt_nv) %>%
  mutate(gear = "Trawl_Bottom")

# filter for midwater trawl
mid_trawl_fix <- gfw_fix_trawl %>%
  dplyr::select(x, y, flag_fin, length_category, fishing_hours = mt_fishing_hours, engine_power_kw_gfw = mt_engine_power, tonnage_gt_gfw = mt_tonnage_gt, nv = mt_nv) %>%
  mutate(gear = "Trawl_Midwater_or_Unsp")


gfw_fix_props <- gfw_fix %>%
  rbind(., bottom_trawl_fix) %>%
  rbind(., mid_trawl_fix) %>%
  group_by(flag_fin, gear, length_category) %>%
  mutate(
            fishing_hours_global = sum(fishing_hours, na.rm = TRUE), 
            engine_power_kw_gfw_global = sum(engine_power_kw_gfw, na.rm = TRUE),
            tonnage_gt_gfw_global = sum(tonnage_gt_gfw, na.rm = TRUE),
            nv_global = sum(nv, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(
         prop_fishing_hours_cell = fishing_hours/fishing_hours_global,
         prop_engine_power_kw_cell = engine_power_kw_gfw/engine_power_kw_gfw_global,
         prop_tonnage_gt_cell = tonnage_gt_gfw/tonnage_gt_gfw_global,
         prop_nv_cell = nv/nv_global) %>%
  filter(!is.na(prop_fishing_hours_cell)) %>% # filter out any NA because there is no effort in those cells (these are the ones where there was either no bottom or mid trawls based on props)
  left_join(eez_rgns) %>% # this join really fucks speed up...
  left_join(eez_lookup) %>%
    mutate(flag_fin = ifelse(flag_fin == ""|flag_fin == "UNK" | is.na(flag_fin), iso_sov1, flag_fin)) %>% # fill in flag with where the fishing occurs, if it is missing. 
  mutate(flag_name = countrycode(sourcevar = flag_fin, origin = "iso3c", destination = "country.name")) %>%
  mutate(flag_fin = ifelse(flag_fin %in% c("HSX", "LND"), "UNK", flag_fin),
         flag_name = ifelse(flag_fin %in% c("UNK"), "Unknown", flag_name)) %>% # %>% # ok these cells are all either on land or in the high seas and don't report the flag country... so we just assume they are unknown unfortunately. 
  # filter(fishing_hours > 0) # if you filter out 0 fishing hours, there are no NAs. I'll keep 0's in for now
  dplyr::select(-mrgid_sov1, -iso_ter1, -geoname, -fishing_hours_global, -nv_global, -engine_power_kw_gfw_global, -tonnage_gt_gfw_global) %>%
  dplyr::select(x, y, everything()) %>%
  mutate(year = y)

# sum(gfw_year$fishing_hours) - sum(gfw_fix_props$fishing_hours) # 2012 is off by ~80k hours??


qs::qsave(gfw_fix_props, file.path(rdsi_dir, glue("prep/gfw_props/deg_half/all_effort_gear_length_props_{y}.qs")))

  }



test_5 <- qread( file.path(rdsi_dir, glue("prep/gfw_props/deg_half/all_effort_gear_length_props_2012.qs")))

test_10 <-  qread(file.path(rdsi_dir, glue("prep/gfw_props/deg_10/all_effort_gear_length_props_2012.qs")))

test_100 <-  qread(file.path(rdsi_dir, glue("prep/gfw_props/deg_100/all_effort_gear_length_props_2012.qs")))


sum(test_5$fishing_hours) # 8386403

sum(test_10$fishing_hours) # 8386403

sum(test_100$fishing_hours) # 6461973

## well, half degree and 0.1 degree match so that's good! We don't use 0.01 degree for our analysis anyways. 

```


